---
title: "The Price of Risk Drives the Business Cycle"
subtitle: "Graduate Student Summer Seminar Series"
title-note: "^1^The views expressed here are the authors’ and are not necessarily representative of the views of the International Monetary Fund, its Management, or its Executive Directors."
author:
  - name: "Tobias Adrian"
    affiliation: "International Monetary Fund^1^"
  - name: "Matthew DeHaven"
    affiliation: Brown University
  - name: "Fernando Duarte"
    affiliation: Brown University
date: "2024-06-27"
callout-appearance: minimal
editor:
  render-on-save: true
format:
  clean-revealjs:
    self-contained: true
---

```{r}
#| include: false
library(data.table)
library(tidyfast)
library(vars)
library(svars)
library(fevdid)
library(ggplot2)
library(patchwork)
library(vfciBCHelpers)
```

```{r}
#| include: false

theme_pres <-
    theme_bw(base_size = 16) +
    theme(
        plot.title = element_text(hjust = 0.5),
        axis.text = element_text(size = 12, margin = margin(0, 0, 0, 0)),
        axis.title = element_text(size = 14),
        plot.caption = element_text(size = 11),
        panel.grid = element_blank(),
        plot.background = element_blank(),
        panel.background = element_blank(),
        strip.background = element_blank(),
        legend.key=element_rect(fill = alpha("white", 0)),
        legend.position = "top",
        legend.background = element_blank(),
        legend.title = element_text(size = 14),
        legend.margin = margin(0, 0, 0, 0),
        legend.text = element_text(size = 10),
        plot.margin = margin(10, 10, 0, 10, "pt")
        )

var_order <- c(
    VFCI = "vfci",
    Unemployment = "unemployment",
    `Fed Funds` = "interest",
    Inflation = "inflation",
    Output = "output",
    Investment = "investment",
    Consumption = "consumption",
    `Hours Worked` = "hours_worked",
    `Labor Share` = "labor_share",
    TFP = "TFP",
    `Labor Prod.` = "productivity"
)

crossplot_order <- c("Output", "Investment", "Consumption", "Hours Worked", "Fed Funds", "Labor Prod.", "Inflation", "TFP", "VFCI")


## Colors
bg_line_color <- "gray50"
recession_color <- "gray80"

hetreg_color <- "steelblue"
max_share_color <- "darkorange"
```

```{r var_setup}
lags <- 2
bc_freqs <- c( 2 * pi / 32, 2 * pi / 6)

vfci_data <- get_var_data(cols = c(vfci = "vfci_fgr1gdpc1", vfci_10 = "vfci_fgr10gdpc1", mu = "mu_fgr1gdpc1"))

data <- get_var_data(vfci = "vfci_fgr1gdpc1")
data[, t := .I - lags]

v <- fit_var(data[, -c("t", "vfci")], lags = lags)
v_vfci <- fit_var(data[, -"t"], lags = lags)

mv <- id_fevdfd(v, "unemployment", bc_freqs)
hv <- id_linear_het_reg(v, "unemployment")
cv <- fevdid::id_ordered_chol(reorder_var(v_vfci, "vfci"))
```

```{r load_recession_dates}
rec_dates <- get_recession_dt()
```

```{r}
pr_data <-
  hv$het_reg$dt |>
  merge(data[, .(t, date)], by = "t") |>
  _[variable == "unemployment"]  
```

## Co-movement over the Business Cycle

Empirically, we observe co-movement across macro variables over the business cycle.

. . .

**Key Question in Macroeconomics:** What drives this co-movement?

- A single shock
- A common propagation mechanism

. . .

We show that...

- the "price of risk" is a common propagation mechanism that can be used to discern between models of the business cycle.

## Outline

0. Result preview

. . .

1. What is the price of risk?

. . .

2. Vector Auto Regression (VAR) setup
    - Identification problem
    - Instrumenting with the VFCI

. . .

3. Estimate the price of risk internal to the VAR
    - Looks like the business cycle

. . .

4. Interpretation and next steps

## Result Preview: *quacks* like a business cycle

```{r}
#| echo: false
#| 

hd_data <- hd(hv)$hd |>
  as.data.table() |>
  merge(data[, .(t, date)], by = "t") |>
  _[impulse == "Chol_1"] |>
  _[response == "unemployment"]
  
preview_fig <-
  ggplot() +
  geom_rect(
    data = rec_dates[start %between% as.Date(c("1965-01-01", "2017-01-01"))],
    aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf),
    fill = recession_color
  ) +
  geom_hline(yintercept = 0, color = bg_line_color) + 
  geom_line(
    data = hd_data,
    aes(
      x = date,
      y = hd
    )
  ) +
  labs(
    x = NULL,
    y = "Contributions to Unemployment",
    title = "Historical Contributions to Unemployment Rate from Identifed VAR Shock",
    caption = "NBER recession dates shaded.\nVAR shock identified using the linear heteroskedasticity present in unemployment residuals."
  ) +
  scale_x_date(
    breaks = seq(as.Date("1960-01-01"), as.Date("2020-01-01"), by = "10 years"),
    labels = seq(1960, 2020, 10),
    limits = as.Date(c("1960-01-01", "2021-01-01")),
    expand = c(0,0)
  ) +
  theme_pres
preview_fig
```

## What is the Price of Risk?

- A measure of the amount of compensation required to hold an additional unit of risk (volatility).

. . .

The SDF (stochastic discount factor) can be decomposed...

$$
\begin{align}
\text{SDF}_{t+1} = \mathbb{E}_t[\text{SDF}_{t+1}] + \text{Vol}_t[{\text{SDF}_{t+1}}] \epsilon_{t+1}
\end{align}
$$

- $\mathbb{E}_t[\text{SDF}_{t+1}]$ is the expected mean,
- $\text{Vol}_t[{\text{SDF}_{t+1}}]$ is the expected volatility (the price of risk),
- $\epsilon_{t+1}$ is an unforseen innovation (structural shocks).


## Estimating the Price of Risk

In most macro models, the Euler equation links the SDF to marginal utility:

$$
\text{SDF}_{t+1} = \beta \left(\frac{C_{t+1}}{C_t}\right)^{-\gamma}
$$

. . .

Leveraging this, we can estimate the price of risk using changes in the volatility of consumption growth (or a similar measure).

## VFCI is an Estimate of the Market Price of Risk

Volatility Financial Conditions Index (VFCI) is **a** measure of the price of risk.

. . .

- Introduced in Adrian, Duarte, and Iyer (WP)

. . .

The VFCI is estimated as the predictable 

- changes in the volatility of GDP growth,
- using financial returns.

## VFCI

```{r}
ggplot() +
  geom_rect(
    data = rec_dates[start %between% as.Date(c("1965-01-01", "2017-01-01"))],
    aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf),
    fill = recession_color
  ) +
  geom_hline(yintercept = 0, color = bg_line_color) + 
  geom_line(data = vfci_data, aes(x = date, y = scale(vfci))) +  
  scale_x_date(
    breaks = seq(as.Date("1960-01-01"), as.Date("2020-01-01"), by = "10 years"),
    labels = seq(1960, 2020, 10),
    limits = as.Date(c("1960-01-01", "2021-01-01")),
    expand = c(0,0)
  ) +
  labs(
    x = NULL,
    y = "Coditional Volatility (VFCI)",
    title = "Volatility Financial Conditions Index (VFCI)"
  ) +
  theme_pres
```

##  Mean - Vol Relationship

A key empirical result is that the price of risk (measured by the VFCI), is linearly linked to the conditional mean.



```{r}
#| fig-align: center
vfci_data |>
  ggplot(aes(
    x = vfci,
    y = mu
  )) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_point() +
  theme_pres + 
  theme(aspect.ratio = 1) +
  labs(
    y = "Conditional Mean",
    x = "Condtional Volatility\n(VFCI)",
    title = "Conditionally Linear Mean - Vol"
  )
```

. . .

We will leverage this relationship to identify the business cycle in our VAR.



## VAR Setup

A strucutral VAR with $p$ lags and $k$ variables in vector $y_t$,
$$
B_0 y_t = B_1 y_{t-1} + . . . + B_p y_{t-p} + w_t
$$

the shocks $w_t$ are assumed to be orthogonal and to have unit variance.

. . .

Empirically, only the following $A_i$ matrices and reduced form residuals, $u_t$, are observed,

$$
 y_t = \underbrace{B_0^{-1}B_1}_{A_1} y_{t-1} + \dots + \underbrace{B_0^{-1}B_p}_{A_p} y_{t-p} + \underbrace{B_0^{-1}w_t}_{u_t}
$$





## Identification Problem

Need to identify a column of the impact matrix $B_0^{-1}$ to back out a structural shock:

$$
u_t = B_0^{-1} w_t
$$

. . .

We will start by using the VFCI as an instrument,

$$
\begin{align}
(1)\ \ \ \ \ \mathbf{E}[\text{VFCI}_t w_{1,t}^\prime] &= \alpha \neq 0
\\
(2)\ \ \ \mathbf{E}[\text{VFCI}_t w_{2:k,t}^\prime] &=  0
\end{align}
$$

Where (1) is a relavence condition and (2) is an exogeneity condition.

## Externally Instrumented Price of Risk
```{r}
#| echo: false
#| 

hd_data <- hd(cv)$hd |>
  as.data.table() |>
  merge(data[, .(t, date)], by = "t") |>
  _[impulse == "Chol_1"] |>
  _[response == "unemployment"]
  
ggplot() +
  geom_rect(
    data = rec_dates[start %between% as.Date(c("1965-01-01", "2017-01-01"))],
    aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf),
    fill = recession_color
  ) +
  geom_hline(yintercept = 0, color = bg_line_color) + 
  geom_line(
    data = hd_data,
    aes(
      x = date,
      y = hd
    )
  ) +
  labs(
    x = NULL,
    y = "Contributions to Unemployment",
    title = "Historical Contributions to Unemployment Rate from Identifed VAR Shock",
    caption = "NBER recession dates shaded.\n VAR shock identified using the VFCI as an external instrument."
  ) +
  scale_x_date(
    breaks = seq(as.Date("1960-01-01"), as.Date("2020-01-01"), by = "10 years"),
    labels = seq(1960, 2020, 10),
    limits = as.Date(c("1960-01-01", "2021-01-01")),
    expand = c(0,0)
  ) +
  theme_pres
```


## Mean-Vol Relationship in the VAR

Taking our VAR with $p$ lags and $k$ variables in vector $y_t$,

$$
y_t = A_1 y_{t-1} + \text{ ... } + A_p y_{t-p} + u_t
$$

The log variance of the residual is then, $\sigma_t = \log\left(\widehat{u}_t^2\right)$.

Then estimate,

$$
\sigma_t = \alpha y_{t} + \varepsilon_t
$$

The mean-vol relationship is between $\widehat{y}_t$ and $\widehat{\sigma}_t$.

:::{.callout-note}
There are $k$ mean-vol relationships in the VAR, one for each variable.
:::


## Mean-Vol Relationship in the VAR {#mean-vol-unemployment}

Using unemployment, we can see a strong linear relationship between the conditional mean and the conditional volatility.

```{r}
var_means <-
  fitted(v) |>
  as.data.table() |> 
  _[, t := .I] |>
  tidyfast::dt_pivot_longer(-t, names_to = "variable", values_to = "var_fitted")

mean_vol_data <- 
  hv$het_reg$dt |> 
  merge(var_means, by = c("t", "variable"))

mean_vol_data |>
  _[variable == "unemployment"] |>
  ggplot(aes(y = var_fitted, x = scale(log_var_fitted) )) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_point() +
  theme_pres +
  labs(
    x = "Conditional Volatility",
    y = "Conditional Mean",
    title = "VAR Implied Mean-Vol for Unemployment"
  ) +
  theme(aspect.ratio = 1)
```


## VAR Implied Price of Risk

```{r}
ggplot() +
  geom_rect(
    data = rec_dates[start %between% as.Date(c("1965-01-01", "2017-01-01"))],
    aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf),
    fill = recession_color
  ) +
  geom_hline(yintercept = 0, color = bg_line_color) + 
  geom_line(
    data = pr_data,
    aes(
      x = date,
      y = scale(log_var_fitted)
    )
  ) +
  scale_x_date(
    breaks = seq(as.Date("1960-01-01"), as.Date("2020-01-01"), by = "10 years"),
    labels = seq(1960, 2020, 10),
    limits = as.Date(c("1960-01-01", "2021-01-01")),
    expand = c(0,0)
  ) +
  labs(
    title = "Price of Risk Implied by VAR Residuals of Unemployment",
    x = NULL,
    y = "Conditional Volatility"
  ) +
  theme_pres

```

## Result review: *quacks* like a business cycle

```{r}
#| echo: false
#| 
preview_fig
```


## Matches the "Business Cycle Anatomy" Shock

Angeletos et al. (2020) identify a "Main Business Cycle (MBC) Shock" which drives most of the unforecasted variation in unemployment.

. . .

Our internal VAR shock has a correlation with the MBC shock of $0.75$.


## Matches the "Business Cycle Anatomy" Shock

```{r}
#| echo: false
#| 

hd_hv_data <- hd(hv)$hd |>
  as.data.table() |>
  merge(data[, .(t, date)], by = "t") |>
  _[impulse == "Chol_1"] |>
  _[response == "unemployment"]
  
hd_mv_data <- hd(mv)$hd |>
  as.data.table() |>
  merge(data[, .(t, date)], by = "t") |>
  _[impulse == "Main"] |>
  _[response == "unemployment"]

hd_data <- rbind(
  hd_hv_data[, method := "hetreg"],
  hd_mv_data[, method := "max_share"]
)

ggplot() +
  geom_rect(
    data = rec_dates[start %between% as.Date(c("1965-01-01", "2017-01-01"))],
    aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf),
    fill = recession_color
  ) +
  geom_hline(yintercept = 0, color = bg_line_color) + 
  geom_line(
    data = hd_data,
    aes(
      x = date,
      y = hd,
      color = method
    )
  ) +
  labs(
    x = NULL,
    y = "Contributions to Unemployment",
    title = "Historical Contributions to Unemployment Rate from Identifed VAR Shocks",
    caption = "NBER recession dates shaded."
  ) +
  scale_x_date(
    breaks = seq(as.Date("1960-01-01"), as.Date("2020-01-01"), by = "10 years"),
    labels = seq(1960, 2020, 10),
    limits = as.Date(c("1960-01-01", "2021-01-01")),
    expand = c(0,0)
  ) +
  scale_color_manual(
    values = c(hetreg_color, max_share_color),
    labels = c(hetreg = "Internal VAR\nPrice of Risk", max_share = "Max Share"),
    name = "Method"
  ) +
  theme_pres
```

## A New Business Cycle Fact

We show a new key feature of business cycles:

- the price of risk comoves with the mean of macro varaibles during the cycle.

. . .

This is not a feature in Real Business Cycle (RBC) or New Keynesian models.

. . .

You need a model that links second moments (volatilities) to first moments.


## Financial Conditions Propagate the Business Cycle

There exists a general equilibrium structural model that is consistent with

- the existence of the mean-vol relationship in the empirical data,

- and that the price of risk comoves with the business cycle.

. . .

This model has financial conditions aggregate, amplify, and propagate disparate shocks to the rest of the economy.

- Driven by a financial friction: the value-at-risk constraint.

## Conclusion

- Can use the VFCI or an internal VAR implied price of risk to identify a shock.
    - Looks like a business cycle.
    - Matches the "Main Business Cycle" shock.

- Adds a new empirical fact for distinguishing models of the business cycle.

- Consistent with a structural model where financial frictions create a common propagation mechanism through the price of risk.


# Appendix {.appendix visibility="uncounted"}

## Comparing Internal and External Price of Risk {visibility="uncounted"}

```{r}
comp_data <- 
  pr_data |>
  merge(data[, .(date, vfci)], by = "date")
  
ggplot() +
  geom_rect(
    data = rec_dates[start %between% as.Date(c("1965-01-01", "2017-01-01"))],
    aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf),
    fill = recession_color
  ) +
  geom_line(data = comp_data, aes(x = date, y = scale(log_var_fitted), color = "int_vfic")) + 
  geom_line(data = comp_data, aes(x = date, y = scale(vfci), color = "ext_vfci")) + 
  scale_x_date(
    breaks = seq(as.Date("1960-01-01"), as.Date("2020-01-01"), by = "10 years"),
    labels = seq(1960, 2020, 10),
    limits = as.Date(c("1960-01-01", "2021-01-01")),
    expand = c(0,0)
  ) +
  labs(
  ) +
  theme_pres

```

## Appendix: IRF of HVAR {visibility="uncounted"}

```{r}
irf(hv, n.ahead = 40)$irf |>
  as.data.table() |>
  _[impulse == "Chol_1"] |>
  ggplot(aes(
    x = h,
    y = irf
  )) +
  geom_hline(yintercept = 0, color = bg_line_color) +
  geom_line() + 
  facet_wrap(
    vars(response),
    scales = "free_y"
  ) +
  theme_pres
```
