---
title: "Financial Conditions and the Business Cycle"
subtitle: "Brown Macro Breakfast"
title-note: "^1^The views expressed here are the authors’ and are not necessarily representative of the views of the International Monetary Fund, its Management, or its Executive Directors."
author:
  - name: "Tobias Adrian"
    affiliation: "International Monetary Fund^1^"
  - name: "Matthew DeHaven"
    affiliation: Brown University
  - name: "Fernando Duarte"
    affiliation: Brown University
date: "2024-02-21"
callout-appearance: minimal
editor:
  render-on-save: true
format:
  clean-revealjs:
    self-contained: true
---

```{r}
#| include: false
require(dplyr)
require(data.table)
require(tidyfast)
require(vars)
require(svars)
require(fevdid)
require(ggplot2)
require(patchwork)
require(gt)
```

```{r}
#| include: false

theme_pres <-
    theme_bw(base_size = 16) +
    theme(
        plot.title = element_text(hjust = 0.5),
        axis.text = element_text(size = 12, margin = margin(0, 0, 0, 0)),
        axis.title = element_text(size = 14),
        plot.caption = element_text(size = 11),
        panel.grid = element_blank(),
        plot.background = element_blank(),
        panel.background = element_blank(),
        strip.background = element_blank(),
        legend.key=element_rect(fill = alpha("white", 0)),
        legend.position = "top",
        legend.background = element_blank(),
        legend.title = element_text(size = 14),
        legend.margin = margin(0, 0, 0, 0),
        legend.text = element_text(size = 10),
        plot.margin = margin(10, 10, 0, 10, "pt")
        )

var_order <- c(
    VFCI = "vfci",
    Unemployment = "unemployment",
    `Fed Funds` = "interest",
    Inflation = "inflation",
    Output = "output",
    Investment = "investment",
    Consumption = "consumption",
    `Hours Worked` = "hours_worked",
    `Labor Share` = "labor_share",
    TFP = "TFP",
    `Labor Prod.` = "productivity"
)

crossplot_order <- c("Output", "Investment", "Consumption", "Hours Worked", "Fed Funds", "Labor Prod.", "Inflation", "TFP", "VFCI")


## Colors
vfci_color <- "steelblue"
u_color <- "darkorange"
vfci_chol_color <- "goldenrod"
```

```{r}
#| include: false
vardata <- fread(here::here("./data/vfciBC_data.csv"))
```

```{r}
#| include: false

## Get Recession data
require(quantmod)
getSymbols("USREC", src = "FRED")

start <- zoo::index(USREC[which(diff(USREC$USREC)==1)])
end   <- zoo::index(USREC[which(diff(USREC$USREC)==-1)-1])

rec <- data.table(start = start, end = end[-1])
```

```{r}
#| include: false

vfciBCdata <- copy(vardata)[, .(date, output, investment, consumption, hours_worked,
    unemployment, labor_share, interest, inflation, productivity, TFP, vfci = vfci_fgr10gdpc1)] |>
    _[date <= as.Date("2017-01-01")]

## Target the BC frequency and umemployment variable
bc_freqs <- c(2 * pi / 32, 2 * pi / 6)
lags <- 2

## Fit the VAR
v <- VAR(vfciBCdata[, -"date"], p = lags, type = "const")

mv_vfci <- id_fevdfd(v, "vfci", bc_freqs, sign = "neg")
mv_u <- id_fevdfd(v, "unemployment", bc_freqs)

##### IRF

irf_vfci <- irf(mv_vfci, n.ahead = 40, impulse = "Main")$irf
irf_u <- irf(mv_u, n.ahead = 40, impulse = "Main")$irf

irf_vfci$model <- "vfci"
irf_u$model <- "u"

irf_df <- rbind(irf_vfci, irf_u)

setDT(irf_df)

##### HD

hd_vfci <- hd(mv_vfci)$hd
hd_u <- hd(mv_u)$hd

hd_vfci$model <- "vfci"
hd_u$model <- "u"

mv_hd_df <- rbind(hd_vfci, hd_u)

setDT(mv_hd_df)

mv_hd_df <- mv_hd_df[impulse == "Main"]
mv_hd_df[, date := rep(vfciBCdata[-(1:lags),"date"][[1]], 22)]

## Plotting
irf_df[, response := factor(response, levels = var_order, labels = names(var_order), ordered = TRUE)]
```

```{r}
targets <- names(vfciBCdata[, -1])

mv_l <- lapply(seq_along(targets), function(i) {
  id_fevdfd(v, targets[[i]], bc_freqs, sign = ifelse(targets[[i]] %in% c("unemployment", "labor_share"), "pos", "neg"))
}) 

hd_df <- rbindlist(lapply(seq_along(targets), function(i) {
  hd <- hd(mv_l[[i]])$hd |> setDT()
  hd <- hd[impulse == "Main"]
  hd$target <- targets[[i]]
  hd$date <- rep(vfciBCdata[-c(1:lags), "date"][[1]], length(targets))
  hd
}))

hs_df <- rbindlist(lapply(seq_along(targets), function(i) {
  hs <- hs(mv_l[[i]])$hs |> setDT()
  hs <- hs[impulse == "Main"]
  hs$target <- targets[[i]]
  hs$date <- vfciBCdata[-c(1:lags), "date"][[1]]
  hs
}))

hs_cross <- merge(
  hs_df[, .(date, target.x = target, hs.x = hs)],
  hs_df[, .(date, target.y = target, hs.y = hs)],
  by = "date",
  all = TRUE,
  allow.cartesian = TRUE
)


hs_cross <- hs_cross |>
  filter(target.x == "unemployment")
hs_cross[, target.y := factor(target.y, levels = var_order, labels = names(var_order), ordered = TRUE)]

corrs <- hs_cross[, .(
  corr = cor(hs.x, hs.y),
  min_x = min(hs.x) * 0.95,
  max_x = max(hs.x) * 0.95,
  max_y = max(hs.y) * 0.975,
  min_y = min(hs.y) * 0.975
  ), by = .(target.x, target.y)] |>
  dplyr::filter(target.x == "unemployment") 
corrs[, loc_y := ifelse(corr < 0, min_y, max_y)]
corrs[, loc_x := ifelse(corr < 0, min_x, max_x)]

```

```{r}
meanvol_data <- copy(vfciBCdata)
meanvol_data[, t := seq_len(nrow(vfciBCdata)) - lags]

meanvol_data <- meanvol_data |>
  dt_pivot_longer(-c(date, t), names_to = "variable", values_to = "original")

resid_data <-
  residuals(v) |>
  as.data.table() |>
  _[, t := .I] |>
  dt_pivot_longer(-t, names_to = "variable", values_to = "residuals")

fitted_data <-
  fitted(v) |>
  as.data.table() |>
  _[, t := .I] |>
  dt_pivot_longer(-t, names_to = "variable", values_to = "fitted")

meanvol_data <- meanvol_data |>
  merge(resid_data, by = c("t", "variable")) |>
  merge(fitted_data, by = c("t", "variable"))

meanvol_data[, log_var := log(residuals^2)]

reg_data <- meanvol_data |>
  merge(
    dt_pivot_wider(meanvol_data[,.(t, variable, original)], names_from = variable, values_from = original),
    by = c("t"),
    allow.cartesian = TRUE
  )

fitted_log_var_data <-
  rbindlist(lapply(unique(reg_data$variable), function(n) {
    het_model <- lm(formula = "log_var ~ 
    lag(output) + lag(investment) + lag(consumption) + lag(hours_worked) + lag(unemployment) + 
    lag(labor_share) + lag(interest) + lag(inflation) + lag(productivity) + lag(TFP) + lag(vfci) +
    lag(output, 2) + lag(investment, 2) + lag(consumption, 2) + lag(hours_worked, 2) + lag(unemployment, 2) + 
    lag(labor_share, 2) + lag(interest, 2) + lag(inflation, 2) + lag(productivity, 2) + lag(TFP, 2) + lag(vfci, 2)
    ", data = reg_data[variable == n])
    fitted(het_model) |>
      as.data.table() |>
      setnames("V1", "fitted_log_var") |>
      _[, variable := n] |>
      _[, t := .I] |>
      _[, resid_log_var := residuals(het_model)]
  }))

meanvol_data <- meanvol_data |>
  merge(fitted_log_var_data, by = c("t", "variable"))

diff_vars <- c("output", "consumption", "investment", "productivity", "TFP")
meanvol_data[, fitted_adj := fitted]
meanvol_data[variable %in% diff_vars, fitted_adj := fitted - shift(fitted, n = 1, type = "lag"), by = "variable"]
```

## Goal for Presentation Today

Focus on the "story" of the project.

Last presentation before I start drafting the paper (3rd year paper).

- Only minor changes to analysis before that draft

- Looking for comments about...
   - structure
   - framing
   - how convincing/persuasive the story is

But feel free to give comments on everything too!

## Co-movement over the Business Cycle

Empirically, we observe co-movement across macro variables over the business cycle.

. . .

**Key Question in Macroeconomics:** What drives this co-movement?

- A single shock
- A common propagation mechanism

. . .

We argue that...

- financial conditions are the common propagation mechanism

## The Business Cycle {#bussiness-cycle-def}

Defined to be a cycle with a period of 6 to 32 quarters.

. . .

- Roughly the range of peak-to-peak NBER recession dates.

. . .

- Usually business cycles are on the longer end of this range


[[Business Cycle Lengths]{.button}](#app-bussiness-cycle-lengths)

## Co-movement in Macro Variables

Angeletos, Collard, and Dellas (2020) run a VAR with 10 macro variables:

. . .

- Output, Unemployment, Investment, Hours Worked, Consumption
- Inflation, Interest Rate, Labor Share, Productivity, TFP

. . .

... and identify a single shock using the max-share identification method.


## Max Share Identification Method

Choose the single shock that explains the largest share of the forecast error variance (FEV) for 

**(1) unemployment** at the 

**(2) business cycle frequency**.

. . .

:::{.callout-note}
Both (1) and (2) can be changed to target different variables or frequencies.
:::

. . .

They call this shock the "main business cycle shock".

## The Main Business Cycle Shock

The shock looks like a usual VAR innovation.

```{r}
#| echo: false
#| fig-align: center
#| fig-height: 3.5

p1_df <- hs_df[target == "unemployment"]

p1 <- 
  ggplot() +
  geom_hline(yintercept = 0, color = "gray50") +
  geom_rect(
    data = rec[start >= as.Date("1962-01-01")],
    aes(
      xmin = start,
      xmax = end,
      ymin = -Inf,
      ymax = +Inf
    ),
    fill = "gray70",
    color = NA,
    alpha = 0.5
  ) +
  geom_line(
    data = p1_df,
    aes(
    x = date,
    y = hs
  )) + 
  labs(
    x = "",
    y = "Business Cycle Shock"
  ) +
  scale_x_date(
    expand = c(0.02, 0),
    breaks = seq(as.Date("1960-01-01"), as.Date("2020-01-01"), by = "10 years"),
    labels = seq(1960, 2020, by = 10)
  ) +
  theme_pres

p2 <- 
  hs_df[target == "unemployment"] |>
  ggplot(aes(
    x = hs
  )) +
  geom_vline(xintercept = 0, color = "black") +
  geom_histogram(fill = "gray50", color = "white") +
  scale_y_continuous(expand = c(0,0), limits = c(0, 30)) + 
  labs(
    x = "Business Cycle Shock",
    y = "Count"
  ) +
  theme_pres
p1 | p2 + plot_layout(widths = c(2,1))
```


## The Main Business Cycle Shock

But the structure of the VAR means the contribution of the shock to macro variables looks like the business cycle.

```{r}
#| echo: false
#| fig-align: center
#| fig-height: 3.5
p3_df <- hd_df |>
  _[target == "unemployment"] |>
  _[response == "unemployment"]

p3 <- 
  ggplot() +
  geom_hline(yintercept = 0, color = "gray50") +
  geom_rect(
    data = rec[start >= as.Date("1962-01-01")],
    aes(
      xmin = start,
      xmax = end,
      ymin = -Inf,
      ymax = +Inf
    ),
    fill = "gray70",
    color = NA,
    alpha = 0.5
  ) +
  geom_line(
    data = p3_df,
    aes(
    x = date,
    y = hd
  )) +
  labs(
    x = NULL,
    y = "Business Cycle Shock\nContribution to Unemployment"
  ) +
  scale_x_date(
    expand = c(0.02, 0),
    breaks = seq(as.Date("1960-01-01"), as.Date("2020-01-01"), by = "10 years"),
    labels = seq(1960, 2020, by = 10)
  ) +
  theme_pres

p3
```

## The Main Business Cycle Shock

This business cycle shock also explains a large amount of the forecast error variance (FEV) in:

- Output, Investment, Hours Worked, and Consumption

. . .

```{r}
fevdid::fevdfd(mv_u)$fevdfd |>
 as.data.table() |>
 _[impulse == "Main"] |>
 _[response %in% c("unemployment", "output", "hours_worked", "consumption", "investment")] |>
 _[f %between% bc_freqs,.(mean_fevdfd = mean(fevdfd)), by = .(response)] |>
 dt_pivot_longer(-response) |>
 _[, response := factor(response, levels = var_order, labels = names(var_order))] |>
 _[, value := round(value, 4)] |>
 dt_pivot_wider(names_from = response, values_from = value) |>
 _[, name := "Explained FEV"] |>
 setnames("name", " ") |>
 gt() |>
 fmt_markdown(columns = 1) |>
 fmt_percent() |>
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_body(columns = 1)
  ) |>
  tab_options(table.font.size = 18)
```


## The Main Business Cycle Shock

If the shock was constructed targeting either

- Output, Investment, Hours Worked, or Consumption

The same "main business cycle shock" would be identified.

```{r}
#| echo: false
#| fig-align: center
#| fig-height: 4

p <- hs_cross |>
  _[target.y %in% c("Output", "Investment", "Hours Worked", "Consumption")] |>
  ggplot(aes(
    x = hs.y,
    y = hs.x
  )) +
  geom_hline(yintercept = 0, color = "gray50") +
  geom_vline(xintercept = 0, color = "gray50") +
  geom_point(
    size = 3,
    alpha = 0.5
  ) +
  geom_text(
    data = corrs[target.y %in% c("Output", "Investment", "Hours Worked", "Consumption")],
    color = "firebrick",
    size = 5,
    hjust = 0,
    aes(
      x = min_y,
      y = max_x,
      label = round(corr, 2)
    )
  ) +
  facet_wrap(
    vars(target.y),
    nrow = 1,
    scales = "free_x"
    ) +
  labs(
    x = "Other Targeted Shock",
    y = "Targeted Unemployment",
    title = "Comparing VAR Shocks",
    caption = "Correlations shown in red numbers."
  ) +
  theme_pres

p
```


## The Main Business Cycle Shock

This is not the case for either

- Inflation or

- TFP

. . .

Thus the authors conclude that the empirical business cycle does not align well with

- RBC productivity shocks or
- New Keynesian demand shocks.


## What Drives the Co-movement?

There are two possibilities:

. . .

1. A single shock

. . .

2. A common propagation mechanism

. . .

Angeletos, Collard, and Dellas (2020) suggest it could be driven by a single shock to "higher order beliefs".

- A type of confidence shock

. . .

We will instead show that financial conditions are a common propagation mechanism.


## Introducing the VFCI

Our preferred measure is the **volatility financial conditions index** (VFCI) constructed in Adrian, Duarte, and Iyer (2023).

. . .

The VFCI...

- is a measure of the *price of risk*
  - can be interpreted as the level of risk aversion of the aggregate housheold

```{r}
#| echo: false
#| fig-align: center
#| fig-height: 3

p_data <- vardata |>
  _[,.(date, vfci = vfci_fgr10gdpc1)] |>
  dt_pivot_longer(-date) |>
  _[!is.na(value)] |>
  _[, value := scale(value)]

p <- ggplot() +
  geom_rect(
    data = rec[start >= as.Date("1962-01-01")],
    aes(
      xmin = start,
      xmax = end,
      ymin = -Inf,
      ymax = +Inf
    ),
    fill = "gray70",
    color = NA,
    alpha = 0.5
    ) +
  geom_hline(yintercept = 0) +
  geom_line(
    data = p_data,
  aes(
    x = date,
    y = value
  )
  ) +
  scale_x_date(
    expand = c(0.02, 0),
    breaks = seq(as.Date("1960-01-01"), as.Date("2020-01-01"), by = "10 years"),
    labels = seq(1960, 2020, by = 10)
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = "VFCI",
    caption = "Constructed using forward gdp growth 10 quarters ahead.\nNBER recessions in gray."
  ) +
  theme_pres

p
```

## Theory Behind the VFCI

In a NK model with financial intermediaries with a value-at-risk constraint...

. . .

- the market price of risk is a natural measure of financial conditions

. . .

- the constraint is tighter in bad times, a nonlinearity in the model.

. . .

The value-at-risk constraint links the mean and volatility of intermediaries' net worth.

. . .

- Which, in equilibrium, links the mean and volatility of output.

. . .


**Key:** The value-at-risk constraint aggregates and propagates diverse shocks into one channel.

. . .

- Any shock changes the volatility of net worth and thus of output.



## Empirical Construction of the VFCI

Estimating the VFCI from forward GDP growth, $\Delta \text{gdp}_{t+h}$, and asset returns, $R_t$,

$$
\Delta \text{gdp}_{t+h} = \theta R_t + \epsilon_t
$$

Then the variance of the residual can be estimated as,

$$
\log(\widehat{\epsilon}_t^2) = \delta R_t + e_t
$$

Then the VFCI is defined as,

$$
\text{VFCI}_t = \widehat{\delta} R_t
$$

and normalized to be mean 0 and variance 1.

## Targeting VFCI Identifies Same Shock

Including VFCI in the VAR and targeting it for the max-share method returns the same business cycle shock.

```{r}
#| echo: false
#| fig-align: center
#| fig-width: 5

p <- hs_cross |>
  _[target.y == "VFCI"] |>
  ggplot(aes(
    x = hs.y,
    y = hs.x
  )) +
  geom_hline(yintercept = 0, color = "gray50") +
  geom_vline(xintercept = 0, color = "gray50") +
  geom_point(
    size = 3,
    alpha = 0.5
  ) +
  geom_text(
    data = corrs[target.y == "VFCI"],
    color = "firebrick",
    size = 5,
    hjust = 0,
    aes(
      x = min_y,
      y = max_x,
      label = paste0("Corr: ", round(corr, 2))
    )
  ) +
  labs(
    x = "Targeted VFCI",
    y = "Targeted Unemployment",
    title = "Comparing VAR Shocks"
  ) +
  theme_pres

p
```

## Targeting VFCI Identifies Same Shock

The impulse response functions (IRFs) are also the same.

```{r}
#| echo: false
#| fig-align: center

p <-
  irf_df |>
  ggplot(aes(
    x = h,
    y = irf,
    color = model
  )) +
  geom_hline(yintercept = 0) +
  geom_line() +
  facet_wrap(
    vars(response),
    nrow = 3,
    scales = "free_y"
  ) +
  labs(
    y = "Impulse Response Function (IRF)",
    x = "Horizon"
  ) +
  scale_color_manual(
    name = "Targeted Variable",
    values = c(
      vfci = vfci_color,
      u = u_color
    ),
    labels = c(
      vfci = "VFCI",
      u = "Unemployment"
    )
  ) +
  theme_pres +
  theme(
    legend.position = c(0.875, 0.15)
    )

p
```

## Targeting VFCI Identifies Same Shock

And the contribution of each shock to unemployment are also similar.

```{r}
#| echo: false
#| fig-align: center
#| fig-height: 3.5
p_df <- mv_hd_df |>
  _[response == "unemployment"]

p <- 
  ggplot() +
  geom_hline(yintercept = 0, color = "gray50") +
  geom_rect(
    data = rec[start >= as.Date("1962-01-01")],
    aes(
      xmin = start,
      xmax = end,
      ymin = -Inf,
      ymax = +Inf
    ),
    fill = "gray70",
    color = NA,
    alpha = 0.5
  ) +
  geom_line(
    data = p_df,
    aes(
    x = date,
    y = hd,
    color = model
  )) +
  labs(
    x = NULL,
    y = "Shock Contribution to Unemployment"
  ) +
  scale_x_date(
    expand = c(0.02, 0),
    breaks = seq(as.Date("1960-01-01"), as.Date("2020-01-01"), by = "10 years"),
    labels = seq(1960, 2020, by = 10)
  ) +
  scale_color_manual(
    name = "Targeted Variable",
    values = c(
      vfci = vfci_color,
      u = u_color
    ),
    labels = c(
      vfci = "VFCI",
      u = "Unemployment"
    )
  ) +
  theme_pres

p
```


## VFCI Co-moves with the Business Cycle

We conclude that financial conditions co-move with the business cycle.

. . .

Why is a financial propagation more compelling than a single belief shock?

. . .

We also see evidence in the VAR of a strong mean-volatility relationship.

- This is present in the model with a value-at-risk constraint,
- but is not present in a model of higher order belief shocks (needs to be proven formally).


## Mean-Vol Relationship in the VAR

Taking our VAR with $p$ lags and $k$ variables in vector $y_t$,

$$
y_t = A_1 y_{t-1} + \text{ ... } + A_p y_{t-p} + u_t
$$

The log variance of the residual is then, $V_t = \log\left(\widehat{u}_t^2\right)$.

Then estimate,

$$
V_t = \alpha_1 y_{t-1} + \text{ ... } + \alpha_p y_{t-p} + \epsilon_t
$$

The mean-vol relationship is between $\widehat{y}_t$ and $\widehat{V}_t$.

:::{.callout-note}
There are $k$ mean-vol relationships in the VAR, one for each variable.
:::


## Mean-Vol Relationship in the VAR {#mean-vol-unemployment}

Using unemployment, we can see a strong linear relationship between the conditional mean and the conditional volatility.

```{r}
#| echo: false
#| fig-align: center
#| fig-height: 4
#| fig-width: 4.25

meanvol_data |>
  _[variable == "unemployment"] |>
  ggplot(aes(
    x = scale(fitted_log_var),
    y = fitted_adj
  )) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "firebrick") +
  labs(
    y = "Condtional Mean",
    x = "Conditional Volatility",
    title = "Mean-Vol: Unemployment"
  ) +
  theme_pres
```

[[Mean-Vol for Output]{.button}](#app-mean-vol-output)


## Financial Conditions Propagating the Business Cycle

There exists a general equilibrium structural model that is consistent with

- the existence of the mean-vol relationship in the empirical data,

- and that financial conditions move strongly with the business cycle,

implying financial conditions are a likely propagation mechanism.

. . .

If this is true, we would also expect a shock to financial conditions to generate the same business cycle dynamics.



## A Shock to VFCI

If we assume a Choleskey idenfication with VFCI ordered first,
then a shock to VFCI generates the same business cycle dynamics.

```{r}
#| echo: false
#| fig-align: center


vfciBCdata <- vardata[, .(date, vfci = vfci_fgr10gdpc1, output, investment, consumption,hours_worked, unemployment, labor_share, interest, inflation, productivity, TFP)] |>
 _[date <= as.Date("2017-01-01")]

## Target the BC frequency and umemployment variable
bc_freqs <- c(2 * pi / 32, 2 * pi / 6)
p <- 2

## Fit the VAR
v <- VAR(vfciBCdata[, -"date"], p = p, type = "const")

mv_vfci <- id_fevdfd(v, "vfci", bc_freqs, sign = "neg")
sv_vfci_chol <- id.chol(v) 

## Flip shock sign to align with FEVD shock direction
sv_vfci_chol$B <- sv_vfci_chol$B * -1


## IRF
irf_vfci <- irf(mv_vfci, n.ahead = 40, impulse = "Main")$irf
irf_vfci_chol <- irf(sv_vfci_chol, n.ahead = 40)$irf |>
      tidyr::pivot_longer(-V1) |>
      mutate(
        response = stringr::str_extract(name, "(?<=% ).*$"),
        impulse = stringr::str_extract(name, "(?<=\\[ ).*(?= \\])"),
        name = NULL
      ) |>
      rename(c(
        h = "V1",
        irf = "value"
      )) |>
      dplyr::filter(impulse == "vfci") |>
      dplyr::select(h, impulse, response, irf)

irf_vfci$model <- "fevdfd"
irf_vfci_chol$model <- "vfci_chol"

irf_df <- rbind(irf_vfci, irf_vfci_chol)

setDT(irf_df)

## Plotting
irf_df[, response := factor(response, levels = var_order, labels = names(var_order), ordered = TRUE)]

plot <-
  irf_df |>
  ggplot(aes(
    x = h,
    y = irf,
    color = model
  )) +
  geom_hline(yintercept = 0) +
  geom_line() +
  facet_wrap(
    vars(response),
    nrow = 3,
    scales = "free_y"
  ) +
  labs(
    y = "Impulse Response Function (IRF)",
    x = "Horizon"
  ) +
  scale_color_manual(
    name = "Identification Method",
    values = c(
      fevdfd = vfci_color,
      vfci_chol = vfci_chol_color
    ),
    labels = c(
      fevdfd = "Max Share",
      vfci_chol = "Cholesky"
    )
  ) +
  theme_pres +
  theme(
    legend.position = c(0.875, 0.15)
    )

plot
```


## Conclusion

Empirical facts of the business cycle are better explained by a common propagation mechanism through the financial sector.

- financial conditions move with the business cycle

- model with value-at-risk contstraint generates the empirically observed mean-volatility relationship

- financial condition shocks generate the same busines cycle


# Appendix {.appendix visibility="uncounted"}

## Appendix: Max-Share Method Details {visibility="uncounted"}

A SVAR(p) model with $p$ lags, for a vector of variables, $x_t$,

$$
B_0 x_t = B_1 x_{t-1} + . . . + B_p x_{t-p} + \epsilon_t
$$

Empirically, only the following $A_i$ matrices and reduced form residuals, $\nu_t$, are observed,

$$
 x_t = \underbrace{B_0^{-1}B_1}_{A_1} x_{t-1} + \dots + \underbrace{B_0^{-1}B_p}_{A_p} x_{t-p} + \underbrace{B_0^{-1}\epsilon_t}_{\upsilon_t}
$$

The identification problem is determining $B_0$.

$$
 \nu_t = B_0^{-1} \epsilon_t
$$

## Appendix: Max-Share Method Details {visibility="uncounted"}

Compute the forecast error for one target variable (i.e. $u$) for target horizon, $h$

$$
  F_{t+h} = x^{(u)}_{t+h} - x^{(u)}_{t+h \mid t} = \sum_{i=0}^{h-1} \underbrace{\Gamma_i}_{IRF} B_0^{(u)^{-1}} \epsilon_{t+h+i}
$$

Choose vector $B_0^{(u)}$ to maximize the variance of $F_{t+h}$,

$$
  \max_{B_0^{(u)}} \ \ \ \text{Var}[F_{t+h}]
$$

This will identify **one** shock, up to a change of sign.

$$
  \epsilon_t^{u} = B_0^{(u)} \hat{\upsilon}_t
$$

For business cycle shock, calculate forecast errors over a frequency range, 6 to 32 quarters.


## Appendix: Business Cycle Lengths {#app-bussiness-cycle-lengths visibility="uncounted"}

The shaded area shows the defined business cycle range of 6 to 32 quarters.

The box plot shows the actual distribution of business cycle lengths since 1950.

```{r}
#| echo: false
#| fig-align: center
#| fig-height: 3

rec |>
 _[,peak_to_peak := lubridate::interval(shift(start, 1, "lag"), start) %/% months(1) / 3] |>
 _[start >= as.Date("1950-01-01")] |>
 ggplot(aes(
  x = peak_to_peak
 )) +
 geom_rect(xmin = 6, xmax = 32, ymin = -Inf, ymax = +Inf, fill = "lightgray")+
 geom_boxplot() +
 scale_y_continuous(expand = c(0.2,0.2), breaks = NULL) +
 scale_x_continuous(limits = c(0,50), breaks = seq(0,50,10), expand = c(0,0)) +
 labs(
  x = "Peak-to-Peak Business Cycle Length",
  title = "NBER Cycle Lengths since 1950"
 ) +
 theme_pres
```

[[Back]{.button}](#bussiness-cycle-def)

## Appendix: Mean-Vol Relationship in VAR {#app-mean-vol-output visibility="uncounted"}

Using output instead.

```{r}
#| echo: false
#| fig-align: center
#| fig-height: 4
#| fig-width: 4.25
meanvol_data |>
  _[variable == "output"] |>
  ggplot(aes(
    x = scale(fitted_log_var),
    y = fitted_adj
  )) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "firebrick") +
  labs(
    y = "Condtional Mean\n(in log-differences)",
    x = "Conditional Volatility",
    title = "Mean-Vol: Output"
  ) +
  theme_pres
```

[[Back]{.button}](#mean-vol-unemployment)