---
title: "Mean-Vol Heteroskedasticity Charts with direct VAR"
format: pdf
project:
  execute-dir: project
---

VAR is estimated with direct horizon forecasts

- `mean_fitted` are the fitted values from the VAR model
- `mean_diff_detrend` take first differences (only for nonstationary variables)
- `mean_lm_detrend` are the residuals from a linear trend
- `mean_hp_detrend` are the residuals from a HP filter
- `mean_bk_detrend` are the residuals from a BK filter (bandpass)
- `mean_hm_detrend` are the residuals from a Hamiltonian filter

```{r}
#| include: false
library(data.table)
library(purrr)
library(vars)
library(svars)
require(fevdid)
require(ggplot2)
library(mFilter)
library(neverhpfilter)
library(vfciBCHelpers)
```

```{r, echo = F, warning = F}
lags <- 2

data <- get_var_data()
x <- data[, -"vfci"] |> copy()
v <- fit_var(x, lags = lags)

hetreg_list <- purrr::map(1:30, ~ fit_het_reg_from_var(fit_var(x, lags = lags, y_lead = .x), hetreg_horizon = 1))

diff_vars <- c("output", "investment", "consumption", "hours_worked", "productivity", "TFP")

hetreg_dt <- purrr::map(1:30, ~ {dt <- hetreg_list[[.x]]$dt; dt$horizon = .x; dt}) |>
  purrr::list_rbind()

## Use mean calculated from each VAR
hetreg_dt <- hetreg_dt |>
  setnames("fitted", "mean_fitted") |>
  _[, mean_diff_detrend := mean_fitted] |>
  _[grepl(paste0(diff_vars, collapse = "|"), variable), mean_diff_detrend := mean_fitted - shift(mean_fitted, 1, type = "lag"), by = .(variable, horizon)]  
  
hetreg_dt[, lm_trend := fitted(lm(data = .SD, mean_fitted ~ t)), by = .(variable, horizon)]
hetreg_dt[, hp_trend := mFilter::hpfilter(mean_fitted, type = "lambda", freq = 1600)$trend, by = .(variable, horizon)]
hetreg_dt[, bk_trend := mFilter::bkfilter(mean_fitted, pl = 6, pu = 32, nfix = 1
, type = "variable")$trend, by = .(variable, horizon)]
hetreg_dt[, hm_trend := coredata(yth_filter(as.xts(ts(mean_fitted, frequency = 4))))[, "y.trend"], by = .(variable, horizon)]

hetreg_dt[, mean_lm_detrend := mean_fitted - lm_trend]
hetreg_dt[, mean_hp_detrend := mean_fitted - hp_trend]
hetreg_dt[, mean_bk_detrend := mean_fitted - bk_trend]
hetreg_dt[, mean_hm_detrend := mean_fitted - hm_trend]

data <- hetreg_dt

data <- data[, .(t, horizon, variable, log_var_fitted, mean_fitted, mean_diff_detrend, mean_lm_detrend, mean_hp_detrend, mean_bk_detrend, mean_hm_detrend)] |>
  tidyfast::dt_pivot_longer(-c(t, horizon, variable, log_var_fitted))
data[, name := factor(name, levels = c("mean_fitted", "mean_diff_detrend", "mean_lm_detrend", "mean_hp_detrend", "mean_bk_detrend", "mean_hm_detrend"), ordered = TRUE)]
```


```{r}
#| output: asis
#| echo: false
for (h in 1:29) {
res <- map_chr(colnames(v$y), \(v) {
  knitr::knit_child(text = c(
    "## Horizon: `r h + 1`, Variable: `r v`",
    "```{r, echo = F, warning = F, fig.height = 3, out.width = '100%'}",
    "data[horizon == h] |>
    _[grepl(v, variable)] |>
    ggplot(aes(y = value, x = log_var_fitted)) +
    geom_point() +
    facet_wrap(vars(name), scales = 'free', ncol = 3) +
    theme_bw(base_size = 8)",
    "```",
    ""), envir = environment(), quiet = TRUE)
})
cat(res, sep = '\n')
}
```
