---
title: Financial Conditions and the Business Cycle
format:
  poster-typst: 
    size: "36x24"
    poster-authors: " "
    departments: " "
    footer-text: "Brown University Poster Session"
    footer-url: "www.matthewdehaven.com"
    footer-emails: "December 8, 2023"
    footer-color: "e6e6e6"
---

```{r}
#| include: false
require(data.table)
require(dplyr)
require(vars)
require(svars)
require(fevdid)
require(ggplot2)
```

```{r}
#| include: false

theme_pres <-
    theme_bw(base_size = 20) +
    theme(
        axis.text = element_text(size = 16, margin = margin(0, 0, 0, 0)),
        axis.title = element_text(size = 18),
        plot.caption = element_text(size = 14),
        panel.grid = element_blank(),
        plot.background = element_blank(),
        panel.background = element_blank(),
        strip.background = element_blank(),
        legend.key=element_rect(fill = alpha("white", 0)),
        legend.position = "top",
        legend.background = element_blank(),
        legend.title = element_text(size = 18),
        legend.margin = margin(0, 0, 0, 0),
        legend.text = element_text(size = 16),
        plot.margin = margin(2, 2, 2, 2, "pt")
        )

var_order <- c(
    VFCI = "vfci",
    Unemployment = "unemployment",
    `Fed Funds` = "interest",
    Inflation = "inflation",
    Output = "output",
    Investment = "investment",
    Consumption = "consumption",
    `Hours Worked` = "hours_worked",
    `Labor Share` = "labor_share",
    TFP = "TFP",
    `Labor Prod.` = "productivity"
)

crossplot_order <- c("Output", "Investment", "Consumption", "Hours Worked", "Fed Funds", "Labor Prod.", "Inflation", "TFP", "VFCI")


## Colors
vfci_color <- "steelblue"
u_color <- "darkorange"
vfci_chol_color <- "goldenrod"
```

```{r}
#| include: false
vardata <- fread(here::here("./data/vfciBC_data.csv"))
```

# Motivation
- Business cycles remarkably similar even when seemingly instigated by different shocks.

- Implies a common shock or a common propogation mechanism.

- "Business-Cycle Anatomy" by Angeletos, Collard, and Dellas (2020) use new max forecast error variance method to empirically identify a "business cycle" shock from a large VAR.

We show financial conditions are a compelling candidate for a common propogation mechanism.

# Volatility Financial Conditions Index (VFCI)
Introduced in "The Market Price of Risk and Macro-Financial Dynamics" by Adrian, Duarte, and Iyer (WP)

- Can be interpreted as the *price of risk*.

- Better than other financial condition indexes at explaining equity and bond risk premia.

- Causal evidence that a tightening of the VFCI leads to a decline in macroeconomic conditions, easing of monetary policy, but little impact on inflation.

- Constructed using (1) asset returns and (2) 10 quarter forward consumption growth

```{r}
#| include: false

## Get Recession data
require(quantmod)
getSymbols("USREC", src = "FRED")

start <- zoo::index(USREC[which(diff(USREC$USREC)==1)])
end   <- zoo::index(USREC[which(diff(USREC$USREC)==-1)-1])

rec <- data.frame(start = start, end = end[-1])
rec <- subset(rec, start >= as.Date("1962-01-01"))
rec <- subset(rec, end <= as.Date("2022-01-01"))
```

```{r}
#| include: false
df <- fread(here::here("./data/vfciBC_data.csv"))
setnames(df, "vfci", "VFCI")

p_data <- df |>
  tidyr::pivot_longer(-"date") |>
  filter(name %in% c("VFCI")) 

p <- ggplot() +
  geom_rect(
    data = rec,
    aes(
      xmin = start,
      xmax = end,
      ymin = -Inf,
      ymax = +Inf
    ),
    fill = "gray70",
    color = NA,
    alpha = 0.5
    ) +
  geom_hline(yintercept = 0) +
  geom_line(
    data = p_data,
  aes(
    x = date,
    y = value
  ),
    linewidth = 1.5
  ) +
  annotate(
    "text",
    x = as.Date("2022-01-01"),
    y = -0.925,
    size = 5,
    hjust = 1,
    label = "NBER recessions in gray."
  ) +
  scale_x_date(
    expand = c(0, 0),
    breaks = seq(as.Date("1960-01-01"), as.Date("2020-01-01"), by = "10 years"),
    labels = seq(1960, 2020, by = 10)
  ) +
  labs(
    x = NULL,
    y = NULL,
  ) +
  theme_pres

ggsave("./figs/vfci.svg", p, width = 10, height = 2.5, units = "in")
```

::: {.block fill="luma(240)" inset="15pt" radius="10pt" width="100%"}

![VFCI](./figs/vfci.svg){#fig-raw-vfci}

:::

<!-- As shown in @fig-raw-vfci, the VFCI peaks before each recession, as well as during some periods of financial stress, such as the savings and loans crisis during the 1980s, the Black Monday stock market crash in October of 1989, and the collapse of Long-Term Capital Management hedge fund in September of 1998. -->

# Max Forecast Error Variance (FEV) ID

A SVAR(p) model with $p$ lags, for a vector of variables, $x_t$,

$$
B_0 x_t = B_1 x_{t-1} + . . . + B_p x_{t-p} + \epsilon_t
$$

Empirically, only the following $A_i$ matrices and reduced form residuals, $\nu_t$, are observed,

$$
 x_t = \underbrace{B_0^{-1}B_1}_{A_1} x_{t-1} + \dots + \underbrace{B_0^{-1}B_p}_{A_p} x_{t-p} + \underbrace{B_0^{-1}\epsilon_t}_{\upsilon_t}
$$

The identification problem is determining $B_0$.

$$
 \nu_t = B_0^{-1} \epsilon_t
$$

Compute the forecast error for one target variable (i.e. $u$) for target horizon, $h$

$$
  F_{t+h} = x^{(u)}_{t+h} - x^{(u)}_{t+h \mid t} = \sum_{i=0}^{h-1} \underbrace{\Gamma_i}_{IRF} B_0^{(u)^{-1}} \epsilon_{t+h+i}
$$

Choose vector $B_0^(u)$ to maximize the variance of $F_{t+h}$,

$$
  \max_{B_0^{(u)}} \ \ \ \text{Var}[F_{t+h}]
$$

This will identify **one** shock, up to a change of sign.

$$
  \epsilon_t^{u} = B_0^{(u)} \hat{\upsilon}_t
$$

For business cycle shock, calculate forecast errors over a frequency range, 6 to 32 quarters.

```{=typst} 
#colbreak()
```

# Using VFCI to Identify the Business Cycle

We use the max FEV identification method to identify two shocks:

- one targeting unemployment---the "Business Cycle" shock,

- one targeting VFCI.

The dynamics of the IRFs are remarkably similar.

```{r}
#| include: false

vfciBCdata <- fread(here::here("./data/vfciBC_data.csv")) |>
    filter(date <= as.Date("2017-01-01"))

vfciBCdata <- vfciBCdata[, .(date, output, investment, consumption, hours_worked,
    unemployment, labor_share, interest, inflation, productivity, TFP, vfci = vfci_fgr10gdpc1)]

## Target the BC frequency and umemployment variable
bc_freqs <- c(2 * pi / 32, 2 * pi / 6)
p <- 2

## Fit the VAR
v <- VAR(vfciBCdata[, -"date"], p = p, type = "const")

mv_vfci <- id_fevdfd(v, "vfci", bc_freqs, sign = "neg")
mv_u <- id_fevdfd(v, "unemployment", bc_freqs)

##### IRF

irf_vfci <- irf(mv_vfci, n.ahead = 40, impulse = "Main")$irf
irf_u <- irf(mv_u, n.ahead = 40, impulse = "Main")$irf

irf_vfci$model <- "vfci"
irf_u$model <- "u"

irf_df <- rbind(irf_vfci, irf_u)

setDT(irf_df)

## Plotting
irf_df[, response := factor(response, levels = var_order, labels = names(var_order), ordered = TRUE)]

plot <-
  irf_df |>
  ggplot(aes(
    x = h,
    y = irf,
    color = model
  )) +
  geom_hline(yintercept = 0) +
  geom_line(linewidth = 1.5) +
  facet_wrap(
    vars(response),
    nrow = 3,
    scales = "free_y"
  ) +
  labs(
    y = "Impulse Response Function (IRF)",
    x = "Horizon"
  ) +
  scale_color_manual(
    name = "Targeted Variable",
    values = c(
      vfci = vfci_color,
      u = u_color
    ),
    labels = c(
      vfci = "VFCI",
      u = "Unemployment"
    )
  ) +
  theme_pres +
  theme(
    legend.position = c(0.875, 0.15)
    )

## Save to disk
ggsave(
  "./figs/irf_plot.svg",
  plot,
  width = 10,
  height = 7,
  units = "in"
  )
```

::: {.block fill=luma(240) inset="15pt" radius="10pt" width="100%"}

![IRFs for Max FEV Shocks Targeting Unemployment or VFCI](./figs/irf_plot.svg){#fig-irf-u-vfci}

:::

- Target each variable with max FEV method and compare with business cycle shock.

VFCI is clearly a part of the "business cycle" block. 
Inflation, TFP, productivity are not.

```{r}
#| include: false
x <- fread(here::here("./data/vfciBC_data.csv")) |>
    filter(date <= as.Date("2017-01-01"))

x <- x[, .(date, output, investment, consumption, hours_worked,
    unemployment, labor_share, interest, inflation, productivity, TFP, vfci = vfci_fgr10gdpc1)]

## Target the BC frequency and umemployment variable
bc_freqs <- c(2 * pi / 32, 2 * pi / 6)
targets <- names(x[, -1])

p <- 2
v <- vars::VAR(x[, -"date"], p = p, type = "const")

## ID shock
mv_l <- lapply(seq_along(targets), function(i) {
  id_fevdfd(v, targets[[i]], bc_freqs, sign = ifelse(targets[[i]] %in% c("unemployment", "labor_share"), "pos", "neg"))
}) 

hs_df <- rbindlist(lapply(seq_along(targets), function(i) {
  hs <- hs(mv_l[[i]])$hs |> setDT()
  hs <- hs[impulse == "Main"]
  hs$target <- targets[[i]]
  hs$date <- x[-c(1:p), "date"][[1]]
  hs
}))

hs_cross <- merge(
  hs_df[, .(date, target.x = target, hs.x = hs)],
  hs_df[, .(date, target.y = target, hs.y = hs)],
  by = "date",
  all = TRUE,
  allow.cartesian = TRUE
)



hs_cross <- hs_cross |>
  filter(target.x == "unemployment")
hs_cross[, target.y := factor(target.y, levels = var_order, labels = names(var_order), ordered = TRUE)]

corrs <- hs_cross[, .(
  corr = cor(hs.x, hs.y),
  min_x = min(hs.x) * 0.95,
  max_y = max(hs.y) * 0.85,
  min_y = min(hs.y) * 0.85
  ), by = .(target.x, target.y)] |>
  dplyr::filter(target.x == "unemployment") 
corrs[, loc_y := ifelse(corr < 0, min_y, max_y)]

plot_cross_vfci <- ggplot() +
  geom_hline(yintercept = 0, color = "gray50") +
  geom_vline(xintercept = 0, color = "gray50") +
  geom_point(
    data = hs_cross,
    size = 2,
    alpha = 0.5,
    aes(
      x = hs.x,
      y = hs.y
    )) +
  geom_text(
    data = corrs,
    color = "firebrick",
    size = 6,
    hjust = 0,
    aes(
      x = min_x,
      y = loc_y,
      label = round(corr, 3)
    )
  ) +
  facet_wrap(
    vars(target.y),
    scales = "free"
  ) +
  theme_pres +
  labs(
    x = "Unemployment Targeted Shock",
    y = "Other Targeted Shock",
    tag = "Correlations for each\nscatter shown in red."
  ) +
  theme(
    plot.margin = margin(2, 2, 2, 2, "pt"),
    plot.tag.position = c(.9, .18),
    plot.tag = element_text(size = 14, color = "firebrick")
    )

## Save to disk
ggsave(
  "./figs/correlation_scatterplots.svg",
  plot_cross_vfci,
  width = 10,
  height = 7,
  units = "in"
)

```

::: {.block fill="luma(240)" inset="15pt" radius="10pt" width="100%"}

![Scatterplot of Max FEV VAR Shocks](./figs/correlation_scatterplots.svg){#fig-u-other-scatter}

:::


# VFCI Shock Generates Same Dynamics

Assume a recursive identifcation scheme (i.e. Choleskey). Set VFCI as the first variable.

$$
  B_0^(vfci) = \begin{bmatrix} b_{0,1}^(vfci) & 0 & 0 & ... & 0 \end{bmatrix}
$$

- Implies that innovations to all other variables do not have a contemperoneous impact on the VFCI.

This can be justified by:

- VFCI is the only financial variable,

- VFCI reacts to any new shock before the slowly moving macro variables.

```{r}
#| include: false

vfciBCdata <- fread(here::here("./data/vfciBC_data.csv")) |>
    filter(date <= as.Date("2017-01-01"))

vfciBCdata <- vfciBCdata[, .(date, vfci = vfci_fgr10gdpc1, output, investment, consumption,hours_worked, unemployment, labor_share, interest, inflation, productivity, TFP)]

## Target the BC frequency and umemployment variable
bc_freqs <- c(2 * pi / 32, 2 * pi / 6)
p <- 2

## Fit the VAR
v <- VAR(vfciBCdata[, -"date"], p = p, type = "const")

mv_vfci <- id_fevdfd(v, "vfci", bc_freqs, sign = "neg")
sv_vfci_chol <- id.chol(v) 

## Flip shock sign to align with FEVD shock direction
sv_vfci_chol$B <- sv_vfci_chol$B * -1


## IRF
irf_vfci <- irf(mv_vfci, n.ahead = 40, impulse = "Main")$irf
irf_vfci_chol <- irf(sv_vfci_chol, n.ahead = 40)$irf |>
      tidyr::pivot_longer(-V1) |>
      mutate(
        response = stringr::str_extract(name, "(?<=% ).*$"),
        impulse = stringr::str_extract(name, "(?<=\\[ ).*(?= \\])"),
        name = NULL
      ) |>
      rename(c(
        h = "V1",
        irf = "value"
      )) |>
      dplyr::filter(impulse == "vfci") |>
      dplyr::select(h, impulse, response, irf)

irf_vfci$model <- "fevdfd"
irf_vfci_chol$model <- "vfci_chol"

irf_df <- rbind(irf_vfci, irf_vfci_chol)

setDT(irf_df)

## Plotting
irf_df[, response := factor(response, levels = var_order, labels = names(var_order), ordered = TRUE)]

plot <-
  irf_df |>
  ggplot(aes(
    x = h,
    y = irf,
    color = model
  )) +
  geom_hline(yintercept = 0) +
  geom_line(linewidth = 1.5) +
  facet_wrap(
    vars(response),
    nrow = 3,
    scales = "free_y"
  ) +
  labs(
    y = "Impulse Response Function (IRF)",
    x = "Horizon"
  ) +
  scale_color_manual(
    name = "Identification Method",
    values = c(
      fevdfd = vfci_color,
      vfci_chol = vfci_chol_color
    ),
    labels = c(
      fevdfd = "Max FEV",
      vfci_chol = "Choleskey"
    )
  ) +
  theme_pres +
  theme(
    legend.position = c(0.875, 0.15)
    )

## Save to disk
ggsave(
  "./figs/irf_choleskey_plot.svg",
  plot,
  width = 10,
  height = 7,
  units = "in"
  )
```

::: {.block fill="luma(240)" inset="15pt" radius="10pt" width="100%"}

![IRFs for VFCI Shock Identified with Max FEV or Choleskey](./figs/irf_choleskey_plot.svg){#fig-irf-maxfev-chol}

:::

A shock to VFCI causes the same dynamics as the identified business cycle shock.

- This is evidence that financial conditions could act as the common propogation mechanism of shocks to the economy.

# Conclusion

- We first showed that VFCI has the same business cycle properties as unemloyment, output, investment, consumption, and hours worked.

- Then we showed that shocks to the VFCI generate the exact dynamics seen in the business cycle.

The implication is that financial conditions are not just a reflection of macroeconomic events, but are either a source of shocks or a common transmission mechanism of shocks from elswhere in the economy.  

```{=typst}
#v(20pt)
```

This makes financial conditions extremely relevant for policymakers and an area that should be given a more prominent focus in macroeconomic research and modeling about the business cycle.
