---
title: "Financial Shocks (Ottonello & Song, WP)"
format: pdf
project:
  execute-dir: project
execute:
  freeze: auto
---

```{r packages}
#| include: false

library(data.table)
library(purrr)
library(vars)
library(svars)
require(fevdid)
require(ggplot2)
library(vfciBCHelpers)
library(tidyquant)
library(sovereign)

fred_key <- "7fdf94c38c6355269067736a82bf7874"
fredr::fredr_set_key(fred_key)
```

```{r data}
#| include: false

fin_shocks <- readxl::read_excel("./data-raw/finshocks_public.xlsx", sheet = "daily-level") |> setDT()
fin_shocks[, date := as.IDate(date)]

ebp_dt <- fread("./data-raw/ebp_csv.csv")

vix_data <- tq_get("^VIX", get = "stock.prices", from = "1990-01-01") |> setDT() |> _[,.(date, vix = log(close))]
u_rate <- fredr::fredr("UNRATE", frequency = "m") |> setDT() |> _[, .(date = as.IDate(date), u_rate = value)]
ip_rate <- fredr::fredr("INDPRO", frequency = "m") |> setDT() |> _[, .(date = as.IDate(date), ip = log(value))]
aaa10ym <- fredr::fredr("AAA10YM", frequency = "m") |> setDT() |> _[, .(date = as.IDate(date), aaa10ym = value)]
baa10ym <- fredr::fredr("BAA10YM", frequency = "m") |> setDT() |> _[, .(date = as.IDate(date), baa10ym = value)]
nfci <- fredr::fredr("NFCI", frequency = "m") |> setDT() |> _[, .(date = as.IDate(date), nfci = value)]
cpi <- fredr::fredr("CORESTICKM159SFRBATL", frequency = "m") |> setDT() |> _[, .(date = as.IDate(date), cpiex = value)]
ff <- fredr::fredr("FEDFUNDS", frequency = "m") |> setDT() |> _[, .(date = as.IDate(date), ff = value)]

dt <- merge(fin_shocks, vix_data, by = "date", all = TRUE)
dt <- dt[, .(
  vf = sum(vf, na.rm = TRUE),
  vf_outside = sum(vf_outside, na.rm = TRUE),
  vf_purged = sum(vf_purged, na.rm = TRUE),
  vix = mean(vix, na.rm = TRUE)
  ),
  by = .(date = as.IDate(format(date, "%Y-%m-01")))]

dt <- merge(dt, ebp_dt, all = TRUE, by = "date") |>
  merge(u_rate, by = "date", all = TRUE) |>
  merge(ip_rate, by = "date", all = TRUE) |>
  merge(aaa10ym, by = "date", all = TRUE) |>
  merge(baa10ym, by = "date", all = TRUE) |>
  merge(nfci, by = "date", all = TRUE) |>
  merge(cpi, by = "date", all = TRUE) |>
  merge(ff, by = "date", all = TRUE)
```


First, we replicate Figure 2 from Ottonello & Song (WP) using their financial shock.

```{r ottonello-song-fig2}
#| echo: false
#| fig-align: center
#| fig-width: 6
#| fig-height: 5

vdt <- dt[, .(date = as.Date(date), vf_purged, ebp, vix, u_rate, ip, aaa10ym, baa10ym)]
vdt <- vdt[complete.cases(vdt)]
iv <- sovereign::VAR(vdt, horizon = 1, p = 1, structure = "IV", instrument = "vf_purged", instrumented = "ebp")
iv_irfs <- sovereign::var_irf(iv, horizon = 24)
iv_irfs |> sovereign::plot_irf()
```


Next, use the same instrument with to identify a shock to the NFCI in a VAR with unemployment, industrial production, inflation, and the federal funds rate.

```{r iv-nfci}
#| echo: false
#| fig-align: center
#| fig-width: 6
#| fig-height: 5

vdt <- dt[, .(date = as.Date(date), vf_purged, nfci, u_rate, ip, cpiex, ff)]
vdt <- vdt[complete.cases(vdt)]
iv <- sovereign::VAR(vdt, horizon = 1, p = 1, structure = "IV", instrument = "vf_purged", instrumented = "nfci")
iv_irfs <- sovereign::var_irf(iv, horizon = 24)
iv_irfs |> sovereign::plot_irf()
```


The same thing, but with the VIX instead of the NFCI.

```{r iv-vix}
#| echo: false
#| fig-align: center
#| fig-width: 6
#| fig-height: 5

vdt <- dt[, .(date = as.Date(date), vf_purged, vix, u_rate, ip, cpiex, ff)]
vdt <- vdt[complete.cases(vdt)]
iv <- sovereign::VAR(vdt, horizon = 1, p = 1, structure = "IV", instrument = "vf_purged", instrumented = "vix")
iv_irfs <- sovereign::var_irf(iv, horizon = 24)
iv_irfs |> sovereign::plot_irf()
```


## Quarterly VFCI

Use the same shock, but aggregated to a quarterly frequency, to identify a VFCI shock in small BCA VAR.

```{r iv-vfci-small}
#| echo: false
#| fig-align: center
#| fig-width: 6
#| fig-height: 4

dt <- get_var_data(
    make_stationary = F,
    cols = c("vfci", "vf_purged", "unemployment", "output", "inflation")
  )  |>
  _[, date := as.Date(date)]

iv <- sovereign::VAR(dt, horizon = 1, p = 1, structure = "IV", instrument = "vf_purged", instrumented = "vfci")
iv_irf <- iv |> sovereign::var_irf()
iv_irf |> sovereign::plot_irf()
```

And finally, use the same instrument, but aggregated to a quarterly frequency, to identify a shock to the VFCI in a VAR with the BCA variables

```{r iv-vfci}
#| echo: false
#| fig-align: center
#| fig-width: 6
#| fig-height: 8

dt <- get_var_data(
    make_stationary = F,
    add_cols = c("vf_purged")
  )  |>
  _[, date := as.Date(date)]

iv <- sovereign::VAR(dt, horizon = 1, p = 1, structure = "IV", instrument = "vf_purged", instrumented = "vfci")
iv_irf <- iv |> sovereign::var_irf(horizon = 12)
iv_irf |> sovereign::plot_irf()
```