---
title: "Compare VFCI estimations"
format: pdf
project:
  execute-dir: project
---

- `baseline` is the vfci estimated with only the 4 financial principal components
- `lags` includes two lags of output/consumption growth
- `exlags` includes two lags of output/consumption growth in the regression, but excludes them from estimating the vfci

```{r}
#| include: false
library(data.table)
library(dplyr)
library(purrr)
require(ggplot2)
library(vfciBCHelpers)
```

```{r, echo = F, warning = F}
data <- readRDS("./data/vfci_data.rds")

data <- data |> select("date", contains("vfci"))

p_vfci_data <-
  data |>
  tidyr::pivot_longer(-date) |>
  mutate(fgr = stringr::str_extract(name, "(?<=fgr)[0-9]*")) |>
  mutate(variable = stringr::str_extract(name, "gdpc1|pcecc96")) |>
  mutate(type = stringr::str_extract(name, "(?<=_)lags|(?<=_)exlags")) |>
  mutate(shift = as.numeric(stringr::str_extract(name, "(?<=s)[0-9]*"))) |>
  mutate(shift = ifelse(is.na(shift), 0, shift)) |>
  mutate(type = ifelse(is.na(type), "baseline", type))
```


```{r}
#| output: asis
#| echo: false
for (s in 0:1) {
res <- map_chr(1:30, \(f) {
  knitr::knit_child(text = c(
    "## FGR Horizon: `r f`, Shift: `r s`",
    "```{r, echo = F, warning = F, fig.height = 3, out.width = '100%'}",
    "p_vfci_data |>
  filter(fgr == f) |>
  filter(shift == s) |>
  ggplot(aes(x = date, y = value, color = type)) + 
  geom_line() +
  facet_wrap(vars(variable), ncol = 1, scales = 'free_y') +
  theme_bw(base_size = 8) +
  theme(legend.position = 'top')",
    "```",
    ""), envir = environment(), quiet = TRUE)
})
cat(res, sep = '\n')
}
```
