---
title: "External VFCI, Growth at Risk Figures"
format: pdf
project:
  execute-dir: project
---

```{r}
#| include: false

library(data.table)
library(fevdid)
library(vars)
library(ggplot2)
library(vfciBCHelpers)
```

```{r}
args <- CJ(
  y = c("output", "consumption", "pcecc96", "gdpc1"),
  forward = 1:30,
  sorted = FALSE
)

ext_vfci_list <-
  args |>
  purrr::pmap(
    est_vfci,
    x = c("pc1", "pc2", "pc3", "pc4")
  )

ext_vfci <-
  seq_len(nrow(args)) |>
  purrr::map(~ cbind(ext_vfci_list[[.x]], args[.x, ])) |>
  purrr::list_rbind() |>
  copy()

```


```{r}
#| echo: false
#| warning: false
#| fig-cap: "Actual Values and Residual Estimates"
#| fig-height: 4
#| out-width: 100%


ext_vfci |>
  _[forward == 1] |>
  ggplot(aes(
    x = date
  )) +
  geom_line(aes(y = residuals, color = "residuals")) +
  geom_line(aes(y = mu + residuals, color = "value")) +
  facet_wrap(vars(y), scales = "free_y")

```

```{r}
#| echo: false
#| warning: false
#| fig-cap: "Mean across time"
#| fig-height: 4
#| out-width: 100%

ext_vfci |>
  _[forward == 1] |>
  ggplot(aes(
    x = date,
    y = mu
  )) +
  geom_line() +
  facet_wrap(vars(y), scales = "free_y")

```

```{r}
#| echo: false
#| warning: false
#| fig-cap: "Log Vol (VFCI) across time"
#| fig-height: 4
#| out-width: 100%

ext_vfci |>
  _[forward == 1] |>
  ggplot(aes(
    x = date,
    y = vfci
  )) +
  geom_line() +
  facet_wrap(vars(y), scales = "free_y")

```

```{r}
#| echo: false
#| warning: false
#| fig-cap: "Vol across time"
#| fig-height: 4
#| out-width: 100%

ext_vfci |>
  _[forward == 1] |>
  ggplot(aes(
    x = date,
    y = exp(vfci)
  )) +
  geom_line() +
  facet_wrap(vars(y), scales = "free_y")

```

```{r}
#| echo: false
#| warning: false
#| fig-cap: "Mean and Log Vol (VFCI) Scatter"
#| fig-height: 4
#| out-width: 100%

ext_vfci |>
  _[forward == 1] |>
  ggplot(aes(
    x = vfci,
    y = mu
  )) +
  geom_point() +
  facet_wrap(vars(y), scales = "free")

```

```{r}
#| echo: false
#| warning: false
#| fig-cap: "Mean and Vol Scatter"
#| fig-height: 4
#| out-width: 100%

ext_vfci |>
  _[forward == 1] |>
  ggplot(aes(
    x = exp(vfci),
    y = mu
  )) +
  geom_point() +
  facet_wrap(vars(y), scales = "free")

```

\newpage

```{r}
#| echo: false
#| warning: false

lms <-
  ext_vfci |>
  _[forward == 1] |>
  split(f = ext_vfci[forward == 1, y]) |>
  purrr::map(~ {
    lm(formula = "mu ~ exp(vfci)", data = .x)
  })

lms |>
  purrr::map(summary)

```


```{r}
#| echo: false
#| warning: false
#| fig-cap: "Mean and Vol across time"
#| fig-height: 4
#| out-width: 100%

ext_vfci |>
  _[forward == 1] |>
  ggplot(aes(
    x = date,
    y = mu
  )) +
  geom_line(aes(color = "mean")) +
  geom_line(aes(y = exp(vfci), color = "vol")) +
  facet_wrap(vars(y), scales = "free_y") +
  theme_bw(base_size = 10)

```

```{r}
#| echo: false
#| warning: false
#| fig-cap: "Growth at risk and not at risk (i.e. Mu +- 2 * Vol)"
#| fig-height: 4
#| out-width: 100%

ext_vfci |>
  _[forward == 1] |>
  #_[y == "output"] |>
  ggplot(aes(
    x = date,
    y = mu
  )) +
  geom_line(aes(y = mu - 2 * exp(vfci), color = "at_risk")) +
  geom_line(aes(y = mu + 2 * exp(vfci), color = "not_at_risk")) +
  facet_wrap(vars(y), scales = "free_y") +
  theme_bw(base_size = 10)

```

```{r}
#| echo: false
#| warning: false
#| fig-cap: "Choose Sigma = b (see below)"
#| fig-height: 4
#| out-width: 100%

mean_vol_coefs <-
  lms |>
  purrr::map(summary) |>
  purrr::map_dbl(~ coef(.x)["exp(vfci)", "Estimate"]) |>
  as.data.table(keep.rownames = TRUE) |>
  setnames(c("V1", "V2"), c("y", "coef"))


ext_vfci |>
  _[forward == 1] |>
  merge(mean_vol_coefs, by = "y") |>
  _[, at_risk := mu + coef * exp(vfci)] |>
  _[, not_at_risk := mu - coef * exp(vfci)] |>
  ggplot(aes(
    x = date,
    y = mu
  )) +
  geom_line(aes(y = at_risk, color = "at_risk")) +
  geom_line(aes(y = not_at_risk, color = "not_at_risk")) +
  facet_wrap(vars(y), scales = "free_y") +
  theme_bw(base_size = 10)

```

```{r}

mean_vol_coefs

```

\newpage

#### Heteroskedastic Regressions:
\begin{align}
y_{t+1} &= \beta_0 + \beta R_t + \epsilon_{t+1}
\\
\log(\widehat{\epsilon_{t+1}^2}) &= \gamma_0 + \gamma R_t + \eta_{t+1}
\\
\text{VFCI}_{t+1} &= \widehat{\log(\widehat{\epsilon_{t+1}^2})}
\end{align}

#### Mean - Vol Relationship
\begin{align}
\widehat{y}_{t+1} &= a + b \exp(\widehat{\text{VFCI}}_{t+1}) + c_{t+1}
\end{align}

Now, construct quantiles of $y_{t+1}$,
\begin{align*}
\widehat{y}_{t+1}^{\text{q05}} &= \widehat{y}_{t+1} - 1.96 \cdot \exp(\widehat{\text{VFCI}}_{t+1})
\end{align*}

More generally, let us choose $\sigma$ for some quantile $q$ of the normal distribution.
\begin{align*}
\widehat{y}_{t+1}^{\text{q}} &= \widehat{y}_{t+1} - \sigma \cdot \exp(\widehat{\text{VFCI}}_{t+1})
\end{align*}

Replacing the mean with the mean-vol relationship,
\begin{align*}
\widehat{y}_{t+1}^{\text{q}} &= a + b \exp(\widehat{\text{VFCI}}_{t+1}) + c_{t+1} - \sigma \cdot \exp(\widehat{\text{VFCI}}_{t+1})
\end{align*}

Now choose a $\sigma = b$, and we get
\begin{align}
\widehat{y}_{t+1}^{\text{q}^*} &= a + c_{t+1}
\end{align}
which if the mean-vol relationship holds is homoskedastic noise around level $a$.