---
title: "Uncertainty, the Price of Risk, and the Business Cycle"
subtitle: "Brown Macro Breakfast"
title-note: "^1^The views expressed here are the authors’ and are not necessarily representative of the views of the International Monetary Fund, its Management, or its Executive Directors."
author:
  - name: "Tobias Adrian"
    affiliation: "International Monetary Fund^1^"
  - name: "Matthew DeHaven"
    affiliation: Brown University
  - name: "Fernando Duarte"
    affiliation: Brown University
date: "2024-11-20"
callout-appearance: minimal
editor:
  render-on-save: true
format:
  clean-revealjs:
    self-contained: true
    html-math-method:
        method: mathjax
        url: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    crossrefs-hover: true
    auto-stretch: false
bibliography: references.bib
---

```{r}
#| include: false
library(data.table)
library(tidyfast)
library(vars)
library(fevdid)
library(vfciBCHelpers)
library(purrr)
library(ggplot2)
library(patchwork)
library(gt)
```

```{r}
#| include: false

theme_pres <-
    theme_bw(base_size = 16) +
    theme(
        plot.title = element_text(hjust = 0.5),
        axis.text = element_text(size = 12, margin = margin(0, 0, 0, 0)),
        axis.title = element_text(size = 14),
        plot.caption = element_text(size = 11),
        panel.grid = element_blank(),
        plot.background = element_blank(),
        panel.background = element_blank(),
        strip.background = element_blank(),
        legend.key = element_blank(),
        legend.position = "top",
        legend.background = element_blank(),
        legend.title = element_text(size = 14),
        legend.margin = margin(0, 0, 0, 0),
        legend.text = element_text(size = 10),
        plot.margin = margin(10, 10, 0, 10, "pt")
        )

var_order <- c(
    VFCI = "vfci",
    Unemployment = "unemployment",
    Output = "output",
    `Fed Funds` = "interest",
    Inflation = "inflation",
    Investment = "investment",
    Consumption = "consumption",
    `Hours Worked` = "hours_worked",
    `Labor Share` = "labor_share",
    TFP = "TFP",
    `Labor Prod.` = "productivity"
)

crossplot_order <- c("Output", "Investment", "Consumption", "Hours Worked", "Fed Funds", "Labor Prod.", "Inflation", "TFP", "VFCI")

irf_horizon <- 32

## Colors
bg_line_color <- "gray50"
recession_color <- "gray80"

hv_macro_color <- "steelblue"
hv_fin_color <- "darkorange"
max_share_color <- "firebrick"
```

```{r var_setup}
lags <- 2
bc_freqs <- c( 2 * pi / 32, 2 * pi / 6)

```

```{r load_recession_dates}
rec_dates <- get_recession_dt()
```

```{r}
vfci_dt <- est_vfci(
  y = "output",
  x = c("pc1", "pc2", "pc3", "pc4"),
  forward = 1
)
```


```{r}
#| include: false
fin_cols <- c("pc1", "pc2", "pc3", "pc4")

data <- get_var_data(
  vfci = NULL,
  end_date = "2022-07-01",
  make_stationary = TRUE
)

hv_macro <-
  data |>
  fit_var(lags = 2) |>
  id_linear_het_reg(
    target = "output",
    hetreg_horizon = 12,
    sign = "neg"
  )
```


```{r}
#| include: false
fin_cols <- c("pc1", "pc2", "pc3", "pc4")

data <- get_var_data(
  vfci = NULL,
  end_date = "2022-07-01",
  add_cols = fin_cols,
  make_stationary = TRUE
)

hv_fin <-
  data[, -c(fin_cols), with = FALSE] |>
  fit_var(lags = 2) |>
  id_linear_het_reg(
    target = "output",
    hetreg_horizon = 12,
    x2 = fin_cols,
    extra_data = data[, fin_cols, with = FALSE],
    method = "mriv",
    sign = "neg"
  )
```

```{r}
#| include: false
#| 
data <- get_var_data(
  vfci = NULL,
  end_date = "2022-07-01",
  make_stationary = TRUE
)

mv <- data |>
  fit_var(lags = 2) |>
  id_fevdfd("unemployment", 2 * pi / c(32, 6), sign = "pos")
```


```{r}
#| include: false

var_names <- 
  c(
    var_1 = "MBC",
    var_2 = "Fin",
    var_3 = "Macro"
  )

corr_dt <-
  list(
    mv,
    hv_fin,
    hv_macro
  ) |>
  hs_corr() |>
  _[impulse_x %in% c("Chol_1", "Main") & impulse_y %in% c("Chol_1", "Main")] |>
  _[, .(var_x, var_y, corr)] |>
  _[, var_x := factor(var_x, levels = names(var_names), labels = var_names, ordered = TRUE)] |>
  _[, var_y := factor(var_y, levels = names(var_names), labels = var_names, ordered = TRUE)] |>
  tidyfast::dt_pivot_wider(names_from = var_y, values_from = corr) 
```

```{r}
#| include: false

hv_macro_irf <-
  hv_macro |>
  irf(n.ahead = irf_horizon) |>
  _$irf |>
  as.data.table() |>
  _[impulse == "Chol_1"] |>
  _[, model := "hv_macro"] |>
  _[, response := factor(response, levels = var_order, labels = names(var_order))]

hv_macro_hd <-
  hv_macro |>
  fevdid::hd() |>
  _$hd |>
  as.data.table() |>
  _[impulse == "Chol_1"] |>
  _[response %in% c("unemployment", "output")] |>
  merge(copy(data)[, t := .I - 2][,. (t, date)], by = "t") |>
  _[, model := "hv_macro"] |>
  _[, response := factor(response, levels = var_order, labels = names(var_order))]
```


```{r}
hv_fin_irf <-
  hv_fin |>
  irf(n.ahead = irf_horizon) |>
  _$irf |>
  as.data.table() |>
  _[impulse == "Chol_1"] |>
  _[, model := "hv_fin"] |>
  _[, response := factor(response, levels = var_order, labels = names(var_order))]

hv_fin_hd <-
  hv_fin |>
  fevdid::hd() |>
  _$hd |>
  as.data.table() |>
  _[impulse == "Chol_1"] |>
  _[response %in% c("output", "unemployment") ] |>
  merge(copy(data)[, t := .I - 2][,. (t, date)], by = "t") |>
  _[, model := "hv_fin"] |>
  _[, response := factor(response, levels = var_order, labels = names(var_order))]
```

```{r}
mv_irf <-
  mv |>
  irf(n.ahead = irf_horizon) |>
  _$irf |>
  as.data.table() |>
  _[impulse == "Main"] |>
  _[, model := "mbc"] |>
  _[, response := factor(response, levels = var_order, labels = names(var_order))]

mv_hd <-
  mv |>
  fevdid::hd() |>
  _$hd |>
  as.data.table() |>
  _[impulse == "Main"] |>
  _[response %in% c("output", "unemployment")] |>
  merge(copy(data)[, t := .I - 2][,. (t, date)], by = "t") |>
  _[, model := "mbc"] |>
  _[, response := factor(response, levels = var_order, labels = names(var_order))]
```

## What Causes Business Cycles?

-   Empirically, just one shock captures majority of the variation of aggregate macroeconomic variables.
-   In models, usually a single big shock:
    - TFP, demand, beliefs, confidence, etc.
-   Common propagation can look like a single shock.

. . .

### Why we care
-   Causes determine whether fluctuations are inefficient.
-   Crucial for welfare and policy.


## What Do We Do? {#what-do-we-do}
We construct three shocks:

1. A shock to macroeconomic uncertainty,
    - Which is a shock to the second moment of output growth.
2. A shock to the price of risk,
    - Which is a shock to the compensation financial intermediaries require to hold additional risk,
3. And a Main Business Cycle (MBC) shock,
    - Which is a statistical shock that explains the largest possible forecast error variance of unemployment at business cycle frequencies.

:::{.aside}
[[Related Literature]{.button}](#literature-review)
:::

## What Do We Find?

All three shocks are remarkably similar.

- Produce almost identical impulse response functions.
- Explain a similar pattern of the historical movements in unemployment.
- In particular, the business cycle shock has a correlation of...
    -  `r round(corr_dt[var_x == "MBC", "Macro"], 2)` with the macroeconomic uncertainty shock,
    - and `r round(corr_dt[var_x == "MBC", "Fin"], 2)` with the price of risk shock.

. . .

**Takeaway:**  We provide new evidence linking second moment shocks to the business cycle, helping distinguish between business cycle models.


## VAR Setup

Structural VAR($p$) with $k$ variables in vector $y_t$,
$$
B_0 y_t = B_1 y_{t-1} + . . . + B_p y_{t-p} + w_t
$$
where structural shocks $w_t$ are white noise.

. . .

Estimate reduced-form coefficients $A_i$ and residuals $u_t$,
$$
 y_t = \underbrace{B_0^{-1}B_1}_{A_1} y_{t-1} + \dots + \underbrace{B_0^{-1}B_p}_{A_p} y_{t-p} + \underbrace{B_0^{-1}w_t}_{u_t}
$$

The shock identification problem: $u_t = B_0^{-1} w_t$.

## Construct a Macroeconomic Uncertainty Shock

The VAR implies forecast errors, $FE_{t+h}$, for each variable $y_t$ at horizon $h$.

Let $\sigma_{t+12} = FE_{t+12}^{\text{(output)}}$ be the forecast error three years out for output growth.

Then we can estimate the OLS regression,

$$
\log \sigma_{t+12}^2 = \alpha y_{t} + \varepsilon_t
$$

$\widehat{\alpha} y_t$ is a valid instrument for a shock to output growth uncertainty, at a horizon of three years.


## What is the Price of Risk?

The *market price of risk* is the equilibrium compensation required by investors to hold all aggregate risk
    
- If all risk were tradable in the stock market,
$$
\text{market price of risk} = 
  \frac{
      \text{expected stock returns} - \text{risk-free rate}
    }{
      \text{volatility of stock returns}
    }
$$

. . .

All risk is **not** tradeable on the stock market, so we will instead back out the price of risk from the Euler equation and the SDF.


## Estimating the Price of Risk

The SDF (stochastic discount factor) can be decomposed...

$$
\begin{align}
\text{SDF}_{t+1} = \mathbb{E}_t[\text{SDF}_{t+1}] + \underbrace{\text{Vol}_t[{\text{SDF}_{t+1}}]}_{\lambda_{t}} \epsilon_{t+1}
\end{align}
$$ {#eq-volatility-sdf}

- $\mathbb{E}_t[\text{SDF}_{t+1}]$ is the expected mean,
- $\lambda_{t}$ is the price of risk,
- $\epsilon_{t+1}$ is an unforseen innovation (structural shocks).


## Estimating the Price of Risk

In most macro models, the Euler equation links the SDF to marginal utility:

$$
\text{SDF}_{t+1} = \beta \left(\frac{C_{t+1}}{C_t}\right)^{-\gamma}
$$

. . .

Taking logs and combining with @eq-volatility-sdf, we get,

$$
\gamma \Delta c_{t,t+1} = \beta - \mathbb{E}_t[\text{sdf}_{t+1}] - \lambda_{t} \epsilon_{t+1}
$$ {#eq-consumption-growth}

Projecting the $\text{sdf}_{t+1}$ onto financial returns $R_t$, we can estimate $\lambda_{t}$ through a heteroskedastic regression.


## The Price of Risk Shock in the VAR

A heteroskedastic regression can be implemented by two OLS steps:

1. Estimate the means (the VAR)
2. Estimate the volatility of the residuals:

Taking $\sigma_{t+12}$ as the forecast error for output growth at horizon 12, 

$$
  \log \sigma_{t+12}^2 = \delta R_t + \varepsilon_t
$$

$\widehat{\delta} R_t$ is a measure of the price of risk consistent with the VAR estimation.

We will use this measure as a shock to the price of risk.


## VAR Specification
```{r}
md_start <- as.yearqtr(min(data$date))
md_end <- as.yearqtr(max(data$date))
md_vars <- paste0(var_order[names(var_order) != "VFCI"], sep = ", ", collapse="")
md_vars2 <- substr(md_vars, 1, nchar(md_vars)-2)
```

Time period:

- `r md_start` to `r md_end`, with $p=2$ lags.

Macro Variables ($k = 10$):

- In log differences: output, investment, consumption, hours worked, productivity, TFP
- In levels: unemployment, interest rate, inflation, labor share

Financial Variables (summarized by first 4 principal components):

- S&P 500 returns, S&P 500 volatility, term spread, liquidity spread, credit spread, default spread

## Unemployment and Output 

```{r}
#| echo: false
#| fig-width: 6
#| fig-height: 3.5
#| fig-align: 'center'

list_rbind(list(
  hv_fin_irf,
  hv_macro_irf
)) |>
  _[response %in% c("Output", "Unemployment")] |>
  ggplot(aes(
    x = h,
    y = irf,
    color = model
  )) +
  geom_hline(yintercept = 0, color = bg_line_color) +
  geom_line() +
  facet_wrap(
    vars(response),
    ncol = 4,
    scales = "free_y"
  ) +
  scale_color_manual(
    name = "Shock",
    values = c(
      hv_macro = hv_macro_color,
      hv_fin = hv_fin_color
    ),
      labels = c(
        hv_macro = "Macroeconomic Uncertainty",
        hv_fin = "Financial Risk Compensation"
      )
  ) +
  scale_x_continuous(
    breaks = seq(0, irf_horizon, 8)
  ) +
  labs(
    x = "Horizon (quarters)",
    y = "Impulse Response"
  ) +
  theme_pres +
  theme(
    legend.position = "bottom"
  )
```

## All Impulse Responses

```{r}
#| echo: false
#| fig-width: 10
#| fig-align: 'center'

list_rbind(list(
  hv_fin_irf,
  hv_macro_irf
)) |>
  ggplot(aes(
    x = h,
    y = irf,
    color = model
  )) +
  geom_hline(yintercept = 0, color = bg_line_color) +
  geom_line() +
  facet_wrap(
    vars(response),
    ncol = 4,
    scales = "free_y"
  ) +
  scale_color_manual(
    name = "Shock",
    values = c(
      hv_macro = hv_macro_color,
      hv_fin = hv_fin_color
    ),
      labels = c(
        hv_macro = "Macroeconomic Uncertainty",
        hv_fin = "Financial Risk Compensation"
      )
  ) +
  scale_x_continuous(
    breaks = seq(0, irf_horizon, 8)
  ) +
  labs(
    x = "Horizon (quarters)",
    y = "Impulse Response"
  ) +
  theme_pres +
  theme(
    legend.position = c(0.85, 0.25),
    legend.justification = c(1, 1)
  )
```

## Historical Shock Contributions

```{r}
#| echo: false
#| fig-width: 10
#| fig-align: 'center'

list_rbind(list(
  hv_fin_hd,
  hv_macro_hd
)) |>
  {\(.x)
    ggplot() +
    geom_rect(
          data = rec_dates[start %between% as.Date(c("1962-01-01", "2022-01-01"))],
          aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf),
          fill = recession_color
        ) +
    geom_hline(yintercept = 0, color = bg_line_color) +
    geom_line(
      data = .x,
      aes(
      x = date,
      y = hd,
      color = model
    )) +
    facet_wrap(
      vars(response),
      ncol = 1,
      scales = "free"
    ) +
    scale_color_manual(
      name = "Shock",
      values = c(
        hv_macro = hv_macro_color,
        hv_fin = hv_fin_color
      ),
      labels = c(
        hv_macro = "Macroeconomic Uncertainty",
        hv_fin = "Financial Risk Compensation"
      )
    ) +
    scale_x_date(
      breaks = seq(as.Date("1960-01-01"), as.Date("2020-01-01"), by = "5 years"),
      labels = seq(1960, 2020, 5)
    ) +
    labs(
      x = NULL,
      y = "Historical Contributions"
    ) +
    theme_pres
  }()

```


## Comparison to the Business Cycle Shock

Identify a "main business-cycle" shock $w_t^{(\text{MBC})}$ as in @mbc2020.

1.    Pick a target variable (e.g. unemployment),
2.    Find linear combination of residuals $w_t^{(\text{MBC})} = b^\prime u_t$ that maximizes its contribution to the variance of target variable over business-cycle frequencies.


##  Comparison to the Business Cycle Shock
```{r}
#| echo: false
#| fig-width: 10
#| fig-align: 'center'

list_rbind(list(
  mv_irf,
  hv_fin_irf,
  hv_macro_irf
)) |>
  ggplot(aes(
    x = h,
    y = irf,
    color = model
  )) +
  geom_hline(yintercept = 0, color = bg_line_color) +
  geom_line() +
  facet_wrap(
    vars(response),
    ncol = 4,
    scales = "free_y"
  ) +
  scale_color_manual(
    name = "Shock",
    values = c(
      hv_fin = hv_fin_color,
      hv_macro = hv_macro_color,
      mbc = max_share_color
    ),
      labels = c(
        hv_fin = "Financial Risk Compensation",
        hv_macro = "Macroeconomic Uncertainty",
        mbc = "Main Business Cycle Shock"
      )
  ) +
  scale_x_continuous(
    breaks = seq(0, irf_horizon, 8)
  ) +
  labs(
    x = "Horizon (quarters)",
    y = "Impulse Response"
  ) +
  theme_pres +
  theme(
    legend.position = c(0.85, 0.25),
    legend.justification = c(1, 1)
  )
```

##  Comparison to the Business Cycle Shock
```{r}
#| echo: false
#| fig-width: 10
#| fig-align: 'center'

list_rbind(list(
  mv_hd,
  hv_fin_hd,
  hv_macro_hd
)) |>
  {\(.x)
    ggplot() +
    geom_rect(
          data = rec_dates[start %between% as.Date(c("1962-01-01", "2022-01-01"))],
          aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf),
          fill = recession_color
        ) +
    geom_hline(yintercept = 0, color = bg_line_color) +
    geom_line(
      data = .x,
      aes(
      x = date,
      y = hd,
      color = model
    )) +
    facet_wrap(
      vars(response),
      ncol = 1,
      scales = "free"
    ) +
    scale_color_manual(
      name = "Shock",
      values = c(
        hv_fin = hv_fin_color,
        hv_macro = hv_macro_color,
        mbc = max_share_color
      ),
        labels = c(
          hv_fin = "Financial Risk Compensation",
          hv_macro = "Macroeconomic Uncertainty",
          mbc = "Main Business Cycle Shock"
        )
    ) +
    scale_x_date(
      breaks = seq(as.Date("1960-01-01"), as.Date("2020-01-01"), by = "5 years"),
      labels = seq(1960, 2020, 5)
    ) +
    labs(
      x = NULL,
      y = "Historical Contributions"
    ) +
    theme_pres
  }()
```


## Distinguisihing Business Cycle Models

New Business Cycle Fact:

- Second moments of macroeconomic variables also comove with first moments over the business cycle.

This new fact can help distinguish between business cycle models.

- Most standard business cycle models do not create this comovement.
- A model with financial frictions can create this comovement.


## Conclusion

We construct three shocks:

- macroeconomic uncertainty, price of risk, and a main business cycle shock.

These shocks produce almost identical impulse response functions and explain a similar pattern of historical movements in unemployment.

These new facts can help distinguish between business cycle models.


# Appendix {.appendix visibility="uncounted"}


## References {visibility="uncounted"}

::: {#refs}
:::

## Select Related Literature {#literature-review visibility="uncounted" .smaller}

@mbc2020 --- Introduces the "Main Business Cycle Shock" by showing the same shock explains a large portion of the variance of output, unemployment, consumption, investment and hours worked over business cycle frequencies.  This same shock does not explain much of the variation in inflation or TFP.

@adrian2023 --- Introduces the "Volatility Financial Conditions Index" (VFCI) as the conditional variance of output growth that is explained by financial factors.  The VFCI is the first FCI grounded in theory and does a better job explaining credit spreads than other FCIs.

@brunnermeier2021 --- Identifies a structural VAR with heteroskedastic regimes in order to look at credit expansions and financial stress shocks.

:::{.aside}
[[What Do We Do?]{.button}](#what-do-we-do)
:::