---
title: "Uncertainty, the Price of Risk, and the Business Cycle"
subtitle: "Brown Macro Breakfast"
title-note: "^1^The views expressed here are the authors’ and are not necessarily representative of the views of the International Monetary Fund, its Management, or its Executive Directors."
author:
  - name: "Tobias Adrian"
    affiliation: "International Monetary Fund^1^"
  - name: "Matthew DeHaven"
    affiliation: Brown University
  - name: "Fernando Duarte"
    affiliation: Brown University
date: "2024-11-20"
callout-appearance: minimal
editor:
  render-on-save: true
format:
  clean-revealjs:
    self-contained: true
    html-math-method:
        method: mathjax
        url: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    crossrefs-hover: true
    auto-stretch: false
bibliography: references.bib
---

```{r}
#| include: false
library(data.table)
library(tidyfast)
library(vars)
library(fevdid)
library(vfciBCHelpers)
library(purrr)
library(ggplot2)
library(patchwork)
library(gt)
```

```{r}
#| include: false

theme_pres <-
    theme_bw(base_size = 16) +
    theme(
        plot.title = element_text(hjust = 0.5),
        axis.text = element_text(size = 12, margin = margin(0, 0, 0, 0)),
        axis.title = element_text(size = 14),
        plot.caption = element_text(size = 11),
        panel.grid = element_blank(),
        plot.background = element_blank(),
        panel.background = element_blank(),
        strip.background = element_blank(),
        legend.key = element_blank(),
        legend.position = "top",
        legend.background = element_blank(),
        legend.title = element_text(size = 14),
        legend.margin = margin(0, 0, 0, 0),
        legend.text = element_text(size = 10),
        plot.margin = margin(10, 10, 0, 10, "pt")
        )

var_order <- c(
    VFCI = "vfci",
    Unemployment = "unemployment",
    Output = "output",
    `Fed Funds` = "interest",
    Inflation = "inflation",
    Investment = "investment",
    Consumption = "consumption",
    `Hours Worked` = "hours_worked",
    `Labor Share` = "labor_share",
    TFP = "TFP",
    `Labor Prod.` = "productivity"
)

crossplot_order <- c("Output", "Investment", "Consumption", "Hours Worked", "Fed Funds", "Labor Prod.", "Inflation", "TFP", "VFCI")

irf_horizon <- 32

## Colors
bg_line_color <- "gray50"
recession_color <- "gray80"

hv_macro_color <- "steelblue"
hv_fin_color <- "darkorange"
max_share_color <- "firebrick"
```

```{r var_setup}
lags <- 2
bc_freqs <- c( 2 * pi / 32, 2 * pi / 6)

```

```{r load_recession_dates}
rec_dates <- get_recession_dt()
```

```{r}
vfci_dt <- est_vfci(
  y = "output",
  x = c("pc1", "pc2", "pc3", "pc4"),
  forward = 1
)
```


```{r}
#| include: false
fin_cols <- c("pc1", "pc2", "pc3", "pc4")

data <- get_var_data(
  vfci = NULL,
  end_date = "2022-07-01",
  make_stationary = TRUE
)

hv_macro <-
  data |>
  fit_var(lags = 2) |>
  id_linear_het_reg(
    target = "output",
    hetreg_horizon = 12,
    sign = "neg"
  )
```


```{r}
#| include: false
fin_cols <- c("pc1", "pc2", "pc3", "pc4")

data <- get_var_data(
  vfci = NULL,
  end_date = "2022-07-01",
  add_cols = fin_cols,
  make_stationary = TRUE
)

hv_fin <-
  data[, -c(fin_cols), with = FALSE] |>
  fit_var(lags = 2) |>
  id_linear_het_reg(
    target = "output",
    hetreg_horizon = 12,
    x2 = fin_cols,
    extra_data = data[, fin_cols, with = FALSE],
    method = "mriv",
    sign = "neg"
  )
```

```{r}
#| include: false
#| 
data <- get_var_data(
  vfci = NULL,
  end_date = "2022-07-01",
  make_stationary = TRUE
)

mv <- data |>
  fit_var(lags = 2) |>
  id_fevdfd("unemployment", 2 * pi / c(32, 6), sign = "pos")
```


```{r}
#| include: false

var_names <- 
  c(
    var_1 = "MBC",
    var_2 = "Fin",
    var_3 = "Macro"
  )

corr_dt <-
  list(
    mv,
    hv_fin,
    hv_macro
  ) |>
  hs_corr() |>
  _[impulse_x %in% c("Chol_1", "Main") & impulse_y %in% c("Chol_1", "Main")] |>
  _[, .(var_x, var_y, corr)] |>
  _[, var_x := factor(var_x, levels = names(var_names), labels = var_names, ordered = TRUE)] |>
  _[, var_y := factor(var_y, levels = names(var_names), labels = var_names, ordered = TRUE)] |>
  tidyfast::dt_pivot_wider(names_from = var_y, values_from = corr) 
```

```{r}
#| include: false

hv_macro_irf <-
  hv_macro |>
  irf(n.ahead = irf_horizon) |>
  _$irf |>
  as.data.table() |>
  _[impulse == "Chol_1"] |>
  _[, model := "hv_macro"] |>
  _[, response := factor(response, levels = var_order, labels = names(var_order))]

hv_macro_hd <-
  hv_macro |>
  fevdid::hd() |>
  _$hd |>
  as.data.table() |>
  _[impulse == "Chol_1"] |>
  _[response %in% c("unemployment", "output")] |>
  merge(copy(data)[, t := .I - 2][,. (t, date)], by = "t") |>
  _[, model := "hv_macro"] |>
  _[, response := factor(response, levels = var_order, labels = names(var_order))]
```


```{r}
hv_fin_irf <-
  hv_fin |>
  irf(n.ahead = irf_horizon) |>
  _$irf |>
  as.data.table() |>
  _[impulse == "Chol_1"] |>
  _[, model := "hv_fin"] |>
  _[, response := factor(response, levels = var_order, labels = names(var_order))]

hv_fin_hd <-
  hv_fin |>
  fevdid::hd() |>
  _$hd |>
  as.data.table() |>
  _[impulse == "Chol_1"] |>
  _[response %in% c("output", "unemployment") ] |>
  merge(copy(data)[, t := .I - 2][,. (t, date)], by = "t") |>
  _[, model := "hv_fin"] |>
  _[, response := factor(response, levels = var_order, labels = names(var_order))]
```

```{r}
mv_irf <-
  mv |>
  irf(n.ahead = irf_horizon) |>
  _$irf |>
  as.data.table() |>
  _[impulse == "Main"] |>
  _[, model := "mbc"] |>
  _[, response := factor(response, levels = var_order, labels = names(var_order))]

mv_hd <-
  mv |>
  fevdid::hd() |>
  _$hd |>
  as.data.table() |>
  _[impulse == "Main"] |>
  _[response %in% c("output", "unemployment")] |>
  merge(copy(data)[, t := .I - 2][,. (t, date)], by = "t") |>
  _[, model := "mbc"] |>
  _[, response := factor(response, levels = var_order, labels = names(var_order))]
```

## What Drives Business Cycles?

-   Empirically, a single shock captures majority of business cycle variation in all macro variables
-   In models, usually one big shock:
    - TFP, demand, beliefs, confidence, etc.
-   Common propagation can look like a single shock

. . .

### Why we care
-   Type of shock and propagation determine whether fluctuations are inefficient
-   Crucial for welfare and policy
-   Select models that are consistent with data

## We Construct Three Shocks That Generate Business Cycles {#what-do-we-do}

1. Macroeconomic uncertainty shock
    - uncertainty: second moment of output growth
2. Price of risk shock
    - price of risk: equilibrium compensation to hold aggregate risk
3. Business cycle shock
    - statistically constructed to explain as much business cycle variation as possible

:::{.aside}
[[Related Literature]{.button}](#literature-review)
:::

## We Construct Three Shocks That Generate Business Cycles

Each of the three shocks produces essentially the same:

- impulse response functions for macro variables
- historical contributions to variation in unemployment, output

- Business cycle shock has a correlation of
    -  `r round(corr_dt[var_x == "MBC", "Macro"], 2)` with the macroeconomic 
    uncertainty shock
    -  `r round(corr_dt[var_x == "MBC", "Fin"], 2)` with the price of risk shock

. . .

## Preview: IRF
```{r}
#| echo: false
#| fig-width: 10
#| fig-align: 'center'

list_rbind(list(
  mv_irf,
  hv_fin_irf,
  hv_macro_irf
)) |>
  ggplot(aes(
    x = h,
    y = irf,
    color = model
  )) +
  geom_hline(yintercept = 0, color = bg_line_color) +
  geom_line() +
  facet_wrap(
    vars(response),
    ncol = 4,
    scales = "free_y"
  ) +
  scale_color_manual(
    name = "Shock",
    values = c(
      hv_fin = hv_fin_color,
      hv_macro = hv_macro_color,
      mbc = max_share_color
    ),
      labels = c(
        hv_fin = "Financial Risk Compensation",
        hv_macro = "Macroeconomic Uncertainty",
        mbc = "Main Business Cycle Shock"
      )
  ) +
  scale_x_continuous(
    breaks = seq(0, irf_horizon, 8)
  ) +
  labs(
    x = "Horizon (quarters)",
    y = "Impulse Response"
  ) +
  theme_pres +
  theme(
    legend.position = c(0.85, 0.25),
    legend.justification = c(1, 1)
  )
```

## Preview: Historical Contribution
```{r}
#| echo: false
#| fig-width: 10
#| fig-align: 'center'

list_rbind(list(
  mv_hd,
  hv_fin_hd,
  hv_macro_hd
)) |>
  {\(.x)
    ggplot() +
    geom_rect(
          data = rec_dates[start %between% as.Date(c("1962-01-01", "2022-01-01"))],
          aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf),
          fill = recession_color
        ) +
    geom_hline(yintercept = 0, color = bg_line_color) +
    geom_line(
      data = .x,
      aes(
      x = date,
      y = hd,
      color = model
    )) +
    facet_wrap(
      vars(response),
      ncol = 1,
      scales = "free"
    ) +
    scale_color_manual(
      name = "Shock",
      values = c(
        hv_fin = hv_fin_color,
        hv_macro = hv_macro_color,
        mbc = max_share_color
      ),
        labels = c(
          hv_fin = "Financial Risk Compensation",
          hv_macro = "Macroeconomic Uncertainty",
          mbc = "Main Business Cycle Shock"
        )
    ) +
    scale_x_date(
      breaks = seq(as.Date("1960-01-01"), as.Date("2020-01-01"), by = "5 years"),
      labels = seq(1960, 2020, 5)
    ) +
    labs(
      x = NULL,
      y = "Historical Contributions"
    ) +
    theme_pres
  }()
```


## Evidence for Financial Shocks or Financial Propagation

-   Business cycles tightly linked to financial conditions and macro uncertainty
-   Shocks move first and second moments of macro variables at the same time
-   Either one big financial shock or common propagation of other shocks through 
financial system

### Takeaway
-   Difficult to reconcile with standard models and shocks
-   Suggests models with financial frictions

::: {.notes}
1)    no standard model has a financial shock as the big shock. Models
without financial shocks rarely have strong financial propagation.
2)    time-varying second moments indicate first and second order approximations
are not sufficient (second order approx gives constant price of risk); any
model that is linear(ized) or approx to second order will not work
3)    preference/belief shocks must hit second moment of consumption (or 
whatever variable determines marginal utility)
:::

## VAR Setup

-   Structural VAR($p$) with $k$ variables in vector $y_t$
$$
B_0 y_t = B_1 y_{t-1} + . . . + B_p y_{t-p} + w_t
$$
where structural shocks $w_t$ are white noise

. . .

-   Estimate reduced-form coefficients $A_i$ and residuals $u_t$
$$
 y_t = \underbrace{B_0^{-1}B_1}_{A_1} y_{t-1} + \dots + \underbrace{B_0^{-1}B_p}_{A_p} y_{t-p} + \underbrace{B_0^{-1}w_t}_{u_t}
$$

-   The shock identification problem: $u_t = B_0^{-1} w_t$

## Construct a Macroeconomic Uncertainty Shock

-   Estimate VAR
-   Let $\sigma_{t+12}$ be the twelve-quarter-ahead forecast error for output growth
-   Measure macroeconomic uncertainty by $\mathbb{E}_t\sigma_{t+12}$
-   Run OLS regression

$$
\log \sigma_{t+12}^2 = \alpha y_{t} + \varepsilon_t
$$

-   Fitted value $\widehat{\alpha} y_t$ is a valid instrument for a shock to uncertainty


<!-- ## What is the Price of Risk? -->

<!-- The *market price of risk* is the equilibrium compensation required by investors to hold all aggregate risk -->

<!-- - If all risk were tradable in the stock market, -->
<!-- $$ -->
<!-- \text{market price of risk} =  -->
<!--   \frac{ -->
<!--       \text{expected stock returns} - \text{risk-free rate} -->
<!--     }{ -->
<!--       \text{volatility of stock returns} -->
<!--     } -->
<!-- $$ -->

<!-- . . . -->

<!-- All risk is **not** tradeable on the stock market, so we will instead back out the price of risk from the Euler equation and the SDF. -->


## The Price of Risk

Absence of arbitrage implies the existence of an SDF (stochastic discount factor)

::: {.notes}
Absence of arbitrage is a necessary condition for equilibrium
:::

The SDF can be decomposed as

$$
\begin{align}
\text{SDF}_{t+1} = \mathbb{E}_t[\text{SDF}_{t+1}] + \underbrace{\text{Vol}_t[{\text{SDF}_{t+1}}]}_{\lambda_{t}} \epsilon_{t+1}
\end{align}
$$ {#eq-volatility-sdf}

where

- $1/\mathbb{E}_t[\text{SDF}_{t+1}]$ is the risk-free rate
- $\lambda_{t}$ is the price of risk
- $\epsilon_{t+1}$ is an innovation

::: {.notes}
$\epsilon_{t+1}$ is a single shock. it may be a function of structural shocks.
$\lambda_{t} \epsilon_{t+1}$ is a forecast error
:::

## Estimating the Price of Risk

Assume a representative agent with CRRA utility for simplicity

The FOC of the representative agent is
$$
\text{SDF}_{t+1} = \beta \left(\frac{C_{t+1}}{C_t}\right)^{-\gamma}
$$
Taking logs and combining with @eq-volatility-sdf, we get

$$
\gamma \Delta c_{t,t+1} = \beta - \mathbb{E}_t[\text{sdf}_{t+1}] - \lambda_{t} \epsilon_{t+1}
$$ {#eq-consumption-growth}

Projecting the $\text{sdf}_{t+1}$ onto financial returns $R_t$, we can estimate $\lambda_{t}$ through a heteroskedastic regression.

::: {.notes}
this works for general utility
If there is a representative agent with utility $U_t$, its FOC is 

$$
\text{SDF}_{t+1} = \frac{\nabla U_{t+1}}{\nabla U_{t}}
$$
the gradient of U is marginal utility, written like this to work with any utility
:::

## The Price of Risk Shock in the VAR

A heteroskedastic regression can be implemented by a two-step procedure:

1. Estimate the means (the VAR)
2. Estimate the volatility of the residuals by OLS

$$
  \log \sigma_{t+12}^2 = \delta R_t + \varepsilon_t
$$

Then $\widehat{\delta} R_t$ is a shock to the price of risk consistent with the VAR estimation.

::: {.notes}
Using the VAR automatically takes care of general utility functions; we are writing marginal utility as a function of macro variables and lags, which should capture the determinants of marginal utility for most/all utilities if enough lags are used
:::


## VAR Specification
```{r}
md_start <- as.yearqtr(min(data$date))
md_end <- as.yearqtr(max(data$date))
md_vars <- paste0(var_order[names(var_order) != "VFCI"], sep = ", ", collapse="")
md_vars2 <- substr(md_vars, 1, nchar(md_vars)-2)
```

Time period:

- `r md_start` to `r md_end`, with $p=2$ lags.

Macro Variables ($k = 10$):

- In log differences: output, investment, consumption, hours worked, productivity, TFP
- In levels: unemployment, interest rate, inflation, labor share

Financial Variables (summarized by first 4 principal components):

- S&P 500 returns, S&P 500 volatility, term spread, liquidity spread, credit spread, default spread

## Unemployment and Output 

```{r}
#| echo: false
#| fig-width: 6
#| fig-height: 3.5
#| fig-align: 'center'

list_rbind(list(
  hv_fin_irf,
  hv_macro_irf
)) |>
  _[response %in% c("Output", "Unemployment")] |>
  ggplot(aes(
    x = h,
    y = irf,
    color = model
  )) +
  geom_hline(yintercept = 0, color = bg_line_color) +
  geom_line() +
  facet_wrap(
    vars(response),
    ncol = 4,
    scales = "free_y"
  ) +
  scale_color_manual(
    name = "Shock",
    values = c(
      hv_macro = hv_macro_color,
      hv_fin = hv_fin_color
    ),
      labels = c(
        hv_macro = "Macroeconomic Uncertainty",
        hv_fin = "Financial Risk Compensation"
      )
  ) +
  scale_x_continuous(
    breaks = seq(0, irf_horizon, 8)
  ) +
  labs(
    x = "Horizon (quarters)",
    y = "Impulse Response"
  ) +
  theme_pres +
  theme(
    legend.position = "bottom"
  )
```

## All Impulse Responses

```{r}
#| echo: false
#| fig-width: 10
#| fig-align: 'center'

list_rbind(list(
  hv_fin_irf,
  hv_macro_irf
)) |>
  ggplot(aes(
    x = h,
    y = irf,
    color = model
  )) +
  geom_hline(yintercept = 0, color = bg_line_color) +
  geom_line() +
  facet_wrap(
    vars(response),
    ncol = 4,
    scales = "free_y"
  ) +
  scale_color_manual(
    name = "Shock",
    values = c(
      hv_macro = hv_macro_color,
      hv_fin = hv_fin_color
    ),
      labels = c(
        hv_macro = "Macroeconomic Uncertainty",
        hv_fin = "Financial Risk Compensation"
      )
  ) +
  scale_x_continuous(
    breaks = seq(0, irf_horizon, 8)
  ) +
  labs(
    x = "Horizon (quarters)",
    y = "Impulse Response"
  ) +
  theme_pres +
  theme(
    legend.position = c(0.85, 0.25),
    legend.justification = c(1, 1)
  )
```

## Historical Shock Contributions

```{r}
#| echo: false
#| fig-width: 10
#| fig-align: 'center'

list_rbind(list(
  hv_fin_hd,
  hv_macro_hd
)) |>
  {\(.x)
    ggplot() +
    geom_rect(
          data = rec_dates[start %between% as.Date(c("1962-01-01", "2022-01-01"))],
          aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf),
          fill = recession_color
        ) +
    geom_hline(yintercept = 0, color = bg_line_color) +
    geom_line(
      data = .x,
      aes(
      x = date,
      y = hd,
      color = model
    )) +
    facet_wrap(
      vars(response),
      ncol = 1,
      scales = "free"
    ) +
    scale_color_manual(
      name = "Shock",
      values = c(
        hv_macro = hv_macro_color,
        hv_fin = hv_fin_color
      ),
      labels = c(
        hv_macro = "Macroeconomic Uncertainty",
        hv_fin = "Financial Risk Compensation"
      )
    ) +
    scale_x_date(
      breaks = seq(as.Date("1960-01-01"), as.Date("2020-01-01"), by = "5 years"),
      labels = seq(1960, 2020, 5)
    ) +
    labs(
      x = NULL,
      y = "Historical Contributions"
    ) +
    theme_pres
  }()

```


## Business Cycle Shock

Identify the business-cycle shock $w_t^{(\text{MBC})}$ as in @mbc2020.

1.    Pick a target variable (e.g. unemployment),
2.    Find linear combination of residuals $w_t^{(\text{MBC})} = b^\prime u_t$ that maximizes its contribution to the variance of target variable over business-cycle frequencies.

Using unemployment, output, consumption, etc., produces the same shock

##  Comparison to the Business Cycle Shock
```{r}
#| echo: false
#| fig-width: 10
#| fig-align: 'center'

list_rbind(list(
  mv_irf,
  hv_fin_irf,
  hv_macro_irf
)) |>
  ggplot(aes(
    x = h,
    y = irf,
    color = model
  )) +
  geom_hline(yintercept = 0, color = bg_line_color) +
  geom_line() +
  facet_wrap(
    vars(response),
    ncol = 4,
    scales = "free_y"
  ) +
  scale_color_manual(
    name = "Shock",
    values = c(
      hv_fin = hv_fin_color,
      hv_macro = hv_macro_color,
      mbc = max_share_color
    ),
      labels = c(
        hv_fin = "Financial Risk Compensation",
        hv_macro = "Macroeconomic Uncertainty",
        mbc = "Main Business Cycle Shock"
      )
  ) +
  scale_x_continuous(
    breaks = seq(0, irf_horizon, 8)
  ) +
  labs(
    x = "Horizon (quarters)",
    y = "Impulse Response"
  ) +
  theme_pres +
  theme(
    legend.position = c(0.85, 0.25),
    legend.justification = c(1, 1)
  )
```

##  Comparison to the Business Cycle Shock
```{r}
#| echo: false
#| fig-width: 10
#| fig-align: 'center'

list_rbind(list(
  mv_hd,
  hv_fin_hd,
  hv_macro_hd
)) |>
  {\(.x)
    ggplot() +
    geom_rect(
          data = rec_dates[start %between% as.Date(c("1962-01-01", "2022-01-01"))],
          aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf),
          fill = recession_color
        ) +
    geom_hline(yintercept = 0, color = bg_line_color) +
    geom_line(
      data = .x,
      aes(
      x = date,
      y = hd,
      color = model
    )) +
    facet_wrap(
      vars(response),
      ncol = 1,
      scales = "free"
    ) +
    scale_color_manual(
      name = "Shock",
      values = c(
        hv_fin = hv_fin_color,
        hv_macro = hv_macro_color,
        mbc = max_share_color
      ),
        labels = c(
          hv_fin = "Financial Risk Compensation",
          hv_macro = "Macroeconomic Uncertainty",
          mbc = "Main Business Cycle Shock"
        )
    ) +
    scale_x_date(
      breaks = seq(as.Date("1960-01-01"), as.Date("2020-01-01"), by = "5 years"),
      labels = seq(1960, 2020, 5)
    ) +
    labs(
      x = NULL,
      y = "Historical Contributions"
    ) +
    theme_pres
  }()
```


<!-- ## Distinguisihing Business Cycle Models -->

<!-- New Business Cycle Fact: -->

<!-- - First and second moments of macroeconomic variables driven by same business cycle shock -->

<!-- This new fact can help distinguish between business cycle models -->

<!-- - Most standard business cycle models do not create this comovement -->
<!-- - A model with financial frictions can create this comovement -->


## Conclusion

We construct three shocks:

- macroeconomic uncertainty, price of risk, and a main business cycle shock

These shocks produce almost identical impulse response functions and explain a similar pattern of historical movements in unemployment and output

First and second moments of macroeconomic variables driven by same business cycle shock

These new facts can help distinguish between business cycle models and point to 
models with financial frictions or financial shocks



# Appendix {.appendix visibility="uncounted"}


## References {visibility="uncounted"}

::: {#refs}
:::

## Select Related Literature {#literature-review visibility="uncounted" .smaller}

@mbc2020 --- Introduces the "Main Business Cycle Shock" by showing the same shock explains a large portion of the variance of output, unemployment, consumption, investment and hours worked over business cycle frequencies.  This same shock does not explain much of the variation in inflation or TFP.

@adrian2023 --- Introduces the "Volatility Financial Conditions Index" (VFCI) as the conditional variance of output growth that is explained by financial factors.  The VFCI is the first FCI grounded in theory and does a better job explaining credit spreads than other FCIs.

@brunnermeier2021 --- Identifies a structural VAR with heteroskedastic regimes in order to look at credit expansions and financial stress shocks.

:::{.aside}
[[What Do We Do?]{.button}](#what-do-we-do)
:::