---
title: "The Price of Risk Drives the Business Cycle"
subtitle: "Brown Macro Breakfast"
title-note: "^1^The views expressed here are the authors’ and are not necessarily representative of the views of the International Monetary Fund, its Management, or its Executive Directors."
author:
  - name: "Tobias Adrian"
    affiliation: "International Monetary Fund^1^"
  - name: "Matthew DeHaven"
    affiliation: Brown University
  - name: "Fernando Duarte"
    affiliation: Brown University
date: "2024-11-20"
callout-appearance: minimal
editor:
  render-on-save: true
format:
  clean-revealjs:
    self-contained: true
    html-math-method:
        method: mathjax
        url: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    crossrefs-hover: true
bibliography: references.bib
---

```{r}
#| include: false
library(data.table)
library(tidyfast)
library(vars)
library(fevdid)
library(vfciBCHelpers)
library(purrr)
library(ggplot2)
library(patchwork)
```

```{r}
#| include: false

theme_pres <-
    theme_bw(base_size = 16) +
    theme(
        plot.title = element_text(hjust = 0.5),
        axis.text = element_text(size = 12, margin = margin(0, 0, 0, 0)),
        axis.title = element_text(size = 14),
        plot.caption = element_text(size = 11),
        panel.grid = element_blank(),
        plot.background = element_blank(),
        panel.background = element_blank(),
        strip.background = element_blank(),
        legend.key = element_blank(),
        legend.position = "top",
        legend.background = element_blank(),
        legend.title = element_text(size = 14),
        legend.margin = margin(0, 0, 0, 0),
        legend.text = element_text(size = 10),
        plot.margin = margin(10, 10, 0, 10, "pt")
        )

var_order <- c(
    VFCI = "vfci",
    Unemployment = "unemployment",
    `Fed Funds` = "interest",
    Inflation = "inflation",
    Output = "output",
    Investment = "investment",
    Consumption = "consumption",
    `Hours Worked` = "hours_worked",
    `Labor Share` = "labor_share",
    TFP = "TFP",
    `Labor Prod.` = "productivity"
)

crossplot_order <- c("Output", "Investment", "Consumption", "Hours Worked", "Fed Funds", "Labor Prod.", "Inflation", "TFP", "VFCI")


## Colors
bg_line_color <- "gray50"
recession_color <- "gray80"

cholesky_color <- "steelblue"
hetreg_color <- "darkorange"
max_share_color <- "firebrick"
```

```{r var_setup}
lags <- 2
bc_freqs <- c( 2 * pi / 32, 2 * pi / 6)

```

```{r load_recession_dates}
rec_dates <- get_recession_dt()
```

```{r}
vfci_dt <- est_vfci(
  y = "output",
  x = c("pc1", "pc2", "pc3", "pc4"),
  forward = 1
)
```


```{r}
#| include: false

data <- get_var_data(
  vfci_dt = est_vfci(
    "output",
    c("pc1", "pc2", "pc3", "pc4"),
    forward = 1
  ),
  end_date = "2022-07-01",
  make_stationary = TRUE
)

cv <- data |>
  fit_var(lags = 2) |>
  reorder_var("vfci") |>
  id_ordered_chol()
```

```{r}
#| include: false

data <- get_var_data(
  vfci_dt = est_vfci(
    "output",
    c("pc1", "pc2", "pc3", "pc4"),
    forward = 12
  ),
  end_date = "2022-07-01",
  make_stationary = TRUE
)

cv_f12 <- data |>
  fit_var(lags = 2) |>
  reorder_var("vfci") |>
  id_ordered_chol()
```

```{r}
#| include: false
fin_cols <- c("pc1", "pc2", "pc3", "pc4")

data <- get_var_data(
  vfci = NULL,
  end_date = "2022-07-01",
  add_cols = fin_cols,
  make_stationary = TRUE
)

hv <- data[, -c(fin_cols), with = FALSE] |>
  fit_var(lags = 2) |>
  id_linear_het_reg(
    target = "output",
    hetreg_horizon = 1,
    x2 = fin_cols,
    extra_data = data[, fin_cols, with = FALSE],
    method = "mriv",
    sign = "neg"
  )
```

```{r}
#| include: false
fin_cols <- c("pc1", "pc2", "pc3", "pc4")

data <- get_var_data(
  vfci = NULL,
  end_date = "2022-07-01",
  add_cols = fin_cols,
  make_stationary = TRUE
)

hv_f12 <- data[, -c(fin_cols), with = FALSE] |>
  fit_var(lags = 2) |>
  id_linear_het_reg(
    target = "output",
    hetreg_horizon = 12,
    x2 = fin_cols,
    extra_data = data[, fin_cols, with = FALSE],
    method = "mriv",
    sign = "neg"
  )
```

```{r}
#| include: false
#| 
data <- get_var_data(
  vfci = NULL,
  end_date = "2022-07-01",
  make_stationary = TRUE
)

mv <- data |>
  fit_var(lags = 2) |>
  id_fevdfd("unemployment", 2 * pi / c(32, 6), sign = "pos")
```


## Introduction / Motivation

Time varying volatility / uncertainty / risk aversion is important for macroeconomic dynamics.^[
List relevant papers here.
]

We leverage this heteroskedasticity in the second moment of output growth to identify a VAR.

The IRFs of this VAR match the IRFs of the Business Cycle Shock from @mbc2020.

Our preferred interpretation is that this is driven by the price of risk through frictions in financial intermediation.


## Outline

1. Quick Summary of @adrian2023
    - Measuring the Price of Risk (VFCI)
    - The Price of Risk Moves Linearly with Conditional Output Growth

2. VAR Setup
    - Identification with the VFCI
    - Identification with Heteroskedasticity

3. Comparison to the Business Cycle Shock

4. Implications

## Measuring the Price of Risk

- Math

## The Price of Risk (VFCI)

```{r}
#| echo: false
#| fig-width: 10
#| fig-align: 'center'

vfci_dt |>
  {\(.x)
    ggplot() +
    geom_rect(
      data = rec_dates[start %between% as.Date(c("1962-01-01", "2022-01-01"))],
      aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf),
      fill = recession_color
    ) +
    geom_hline(yintercept = 0, color = bg_line_color) +
    geom_line(
      data = .x,
      aes(
      x = date,
      y = scale(vfci)
    ),
      color = cholesky_color
    ) +
    labs(
      x = NULL,
      y = "VFCI, Normalized to N(0,1)",
      title = "Time Series of the Price of Risk (VFCI)"
    ) +
    theme_pres
  }()
```

## The Price of Risk Moves Linearly with Conditional Output Growth

```{r}
#| echo: false
#| fig-width: 6
#| fig-align: 'center'

vfci_dt |>
  ggplot(aes(
    x = scale(vfci),
    y = mu
  )) +
  geom_point(
    color = cholesky_color
  ) +
  labs(
    x = "VFCI, Normalized to N(0,1)",
    y = "Conditional Output Growth",
    title = "Mean vs. Volatility of Output Growth"
  ) +
  theme_pres
```

## VAR Setup

- Math

## Indentification with the VFCI

- Math



##  Identification with the VFCI
```{r}
#| echo: false
#| fig-width: 10
#| fig-align: 'center'

cv_irf <-
  cv |>
  irf(n.ahead = 40) |>
  _$irf |>
  as.data.table() |>
  _[impulse == "Chol_1"] |>
  _[, model := "cholesky"]

cv_irf |>
  ggplot(aes(
    x = h,
    y = irf
  )) +
  geom_hline(yintercept = 0, color = bg_line_color) +
  geom_line(
    color = cholesky_color
  ) +
  facet_wrap(
    vars(response),
    ncol = 4,
    scales = "free_y"
  ) +
  labs(
    x = "Horizon (quarters)",
    y = "Impulse Response"
  ) +
  theme_pres
```

##  Identification with the VFCI
```{r}
#| echo: false
#| fig-width: 10
#| fig-align: 'center'


cv_hs <-
  cv |>
  hs() |>
  _$hs |>
  as.data.table() |>
  _[impulse == "Chol_1"] |>
  merge(copy(data)[, t := .I - 2][,. (t, date)], by = "t") |>
  _[, model := "cholesky"]

p1 <-
  cv_hs |>
  {\(.x)
    ggplot() +
    geom_rect(
        data = rec_dates[start %between% as.Date(c("1962-01-01", "2022-01-01"))],
        aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf),
        fill = recession_color
      ) +
    geom_hline(yintercept = 0, color = bg_line_color) +
    geom_line(
      data = .x,
      aes(
      x = date,
      y = hs
    ),
      color = cholesky_color
    ) +
    labs(
      x = NULL,
      y = "Historical Shock",
      title = "(a) Historical Shock"
    ) +
    theme_pres
  }()

cv_hd <-
  cv |>
  fevdid::hd() |>
  _$hd |>
  as.data.table() |>
  _[impulse == "Chol_1"] |>
  _[response == "unemployment"] |>
  merge(copy(data)[, t := .I - 2][,. (t, date)], by = "t") |>
  _[, model := "cholesky"]

p2 <-
  cv_hd |>
  {\(.x)
    ggplot() +
    geom_rect(
          data = rec_dates[start %between% as.Date(c("1962-01-01", "2022-01-01"))],
          aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf),
          fill = recession_color
        ) +
    geom_hline(yintercept = 0, color = bg_line_color) +
    geom_line(
      data = .x,
      aes(
      x = date,
      y = hd
    ),
      color = cholesky_color
    ) +
    labs(
      x = NULL,
      y = "Unemployment",
      title = "(b) Historical Shock Contribution to Unemployment"
    ) +
    theme_pres
  }()

p1 / p2
```


## Indentification with Heteroskedasticity

- Math


## Comparing VFCI to Internal VAR VFCI

```{r}
int_vfci <-
  hv$het_reg$dt |>
  _[variable == "output"] |>
  merge(copy(data)[, t := .I - 2][,. (t, date)], by = "t") |>
  _[, .(date, int_vfci = log_var_fitted, int_mu = fitted)]

vfci_dt[, .(date, ext_vfci = vfci, ext_mu = mu)] |>
  merge(int_vfci, by = "date") |>
  ggplot(aes(
    x = date
  )) +
  geom_hline(yintercept = 0, color = bg_line_color) +
  geom_line(aes(y = scale(int_vfci), color = "int_vfci")) +
  geom_line(aes(y = scale(ext_vfci), color = "ext_vfci")) +
  labs(
    x = NULL,
    y = "VFCI, Normalized to N(0,1)",
    title = "Internal VAR VFCI and External VFCI"
  ) +
  scale_x_date(
    breaks = seq(as.Date("1960-01-01"), as.Date("2025-01-01"), by = "5 years"),
    labels = seq(1960, 2025, by = 5)
  ) +
  scale_color_manual(
    values = c(
      ext_vfci = cholesky_color,
      int_vfci = hetreg_color
    ),
    labels = c(
      int_vfci = "Internal VAR VFCI",
      ext_vfci = "External VFCI"
    )
  ) +
  theme_pres
```


##  Identification with Heteroskedasticity
```{r}
#| echo: false
#| fig-width: 10
#| fig-align: 'center'

hv_irf <-
  hv |>
  irf(n.ahead = 40) |>
  _$irf |>
  as.data.table() |>
  _[impulse == "Chol_1"] |>
  _[, model := "heteroskedasticity"]

list_rbind(list(
  hv_irf,
  cv_irf
)) |>
  ggplot(aes(
    x = h,
    y = irf,
    color = model
  )) +
  geom_hline(yintercept = 0, color = bg_line_color) +
  geom_line() +
  facet_wrap(
    vars(response),
    ncol = 4,
    scales = "free_y"
  ) +
  scale_color_manual(
    values = c(
      cholesky = cholesky_color,
      heteroskedasticity = hetreg_color
    ),
      labels = c(
        cholesky = "Cholesky",
        heteroskedasticity = "Heteroskedasticity"
      )
  ) +
  labs(
    x = "Horizon (quarters)",
    y = "Impulse Response"
  ) +
  theme_pres
```

##  Identification with Heteroskedasticity
```{r}
#| echo: false
#| fig-width: 10
#| fig-align: 'center'

hv_hs <-
  hv |>
  hs() |>
  _$hs |>
  as.data.table() |>
  _[impulse == "Chol_1"] |>
  merge(copy(data)[, t := .I - 2][,. (t, date)], by = "t") |>
  _[, model := "heteroskedasticity"]

p1 <-
  list_rbind(list(
    hv_hs,
    cv_hs
  )) |>
  {\(.x)
    ggplot() +
    geom_rect(
        data = rec_dates[start %between% as.Date(c("1962-01-01", "2022-01-01"))],
        aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf),
        fill = recession_color
      ) +
    geom_hline(yintercept = 0, color = bg_line_color) +
    geom_line(
      data = .x,
      aes(
      x = date,
      y = hs,
      color = model
    )) +
    scale_color_manual(
      values = c(
        cholesky = cholesky_color,
        heteroskedasticity = hetreg_color
      ),
      labels = c(
        cholesky = "Cholesky",
        heteroskedasticity = "Heteroskedasticity"
      )
    ) +
    labs(
      x = NULL,
      y = "Historical Shock",
      title = "(a) Historical Shock"
    ) +
    theme_pres
  }()

hv_hd <-
  hv |>
  fevdid::hd() |>
  _$hd |>
  as.data.table() |>
  _[impulse == "Chol_1"] |>
  _[response == "unemployment"] |>
  merge(copy(data)[, t := .I - 2][,. (t, date)], by = "t") |>
  _[, model := "heteroskedasticity"]

p2 <-
  list_rbind(list(
    hv_hd,
    cv_hd
  )) |>
  {\(.x)
    ggplot() +
    geom_rect(
          data = rec_dates[start %between% as.Date(c("1962-01-01", "2022-01-01"))],
          aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf),
          fill = recession_color
        ) +
    geom_hline(yintercept = 0, color = bg_line_color) +
    geom_line(
      data = .x,
      aes(
      x = date,
      y = hd,
      color = model
    )) +
    scale_color_manual(
      values = c(
        cholesky = cholesky_color,
        heteroskedasticity = hetreg_color
      ),
      labels = c(
        cholesky = "Cholesky",
        heteroskedasticity = "Heteroskedasticity"
      )
    ) +
    labs(
      x = NULL,
      y = "Unemployment",
      title = "(b) Historical Shock Contribution to Unemployment"
    ) +
    theme_pres
  }()

p1 / p2
```


## Comparison to the Business Cycle Shock

- Math


##  Comparison to the Business Cycle Shock
```{r}
#| echo: false
#| fig-width: 10
#| fig-align: 'center'

mv_irf <-
  mv |>
  irf(n.ahead = 40) |>
  _$irf |>
  as.data.table() |>
  _[impulse == "Main"] |>
  _[, model := "mbc"]

list_rbind(list(
  cv_irf,
  hv_irf,
  mv_irf
)) |>
  ggplot(aes(
    x = h,
    y = irf,
    color = model
  )) +
  geom_hline(yintercept = 0, color = bg_line_color) +
  geom_line() +
  facet_wrap(
    vars(response),
    ncol = 4,
    scales = "free_y"
  ) +
  scale_color_manual(
    values = c(
      cholesky = cholesky_color,
      heteroskedasticity = hetreg_color,
      mbc = max_share_color
    ),
      labels = c(
        cholesky = "Cholesky",
        heteroskedasticity = "Heteroskedasticity",
        mbc = "Main Business Cycle Shock"
      )
  ) +
  labs(
    x = "Horizon (quarters)",
    y = "Impulse Response"
  ) +
  theme_pres
```

##  Comparison to the Business Cycle Shock
```{r}
#| echo: false
#| fig-width: 10
#| fig-align: 'center'

mv_hs <-
  mv |>
  hs() |>
  _$hs |>
  as.data.table() |>
  _[impulse == "Main"] |>
  merge(copy(data)[, t := .I - 2][,. (t, date)], by = "t") |>
  _[, model := "mbc"]

p1 <-
  list_rbind(list(
    cv_hs,
    hv_hs,
    mv_hs
  )) |>
  {\(.x)
    ggplot() +
    geom_rect(
        data = rec_dates[start %between% as.Date(c("1962-01-01", "2022-01-01"))],
        aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf),
        fill = recession_color
      ) +
    geom_hline(yintercept = 0, color = bg_line_color) +
    geom_line(
      data = .x,
      aes(
      x = date,
      y = hs,
      color = model
    )) +
    scale_color_manual(
      values = c(
        cholesky = cholesky_color,
        heteroskedasticity = hetreg_color,
        mbc = max_share_color
      ),
      labels = c(
        cholesky = "Cholesky",
        heteroskedasticity = "Heteroskedasticity",
        mbc = "Main Business Cycle Shock"
      )
    ) +
    labs(
      x = NULL,
      y = "Historical Shock",
      title = "(a) Historical Shock"
    ) +
    theme_pres
  }()


mv_hd <-
  mv |>
  fevdid::hd() |>
  _$hd |>
  as.data.table() |>
  _[impulse == "Main"] |>
  _[response == "unemployment"] |>
  merge(copy(data)[, t := .I - 2][,. (t, date)], by = "t") |>
  _[, model := "mbc"]


p2 <-
  list_rbind(list(
    cv_hd,
    hv_hd,
    mv_hd
  )) |>
  {\(.x)
    ggplot() +
    geom_rect(
          data = rec_dates[start %between% as.Date(c("1962-01-01", "2022-01-01"))],
          aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf),
          fill = recession_color
        ) +
    geom_hline(yintercept = 0, color = bg_line_color) +
    geom_line(
      data = .x,
      aes(
      x = date,
      y = hd,
      color = model
    )) +
    scale_color_manual(
      values = c(
        cholesky = cholesky_color,
        heteroskedasticity = hetreg_color,
        mbc = max_share_color
      ),
      labels = c(
        cholesky = "Cholesky",
        heteroskedasticity = "Heteroskedasticity",
        mbc = "Main Business Cycle Shock"
      )
    ) +
    labs(
      x = NULL,
      y = "Unemployment",
      title = "(b) Historical Shock Contribution to Unemployment"
    ) +
    theme_pres
  }()

p1 / p2
```


## Using 3-year Horizon for the VFCI

```{r}
#| echo: false
#| fig-width: 10
#| fig-align: 'center'

cv_irf_f12 <-
  cv_f12 |>
  irf(n.ahead = 40) |>
  _$irf |>
  as.data.table() |>
  _[impulse == "Chol_1"] |>
  _[, model := "cholesky"]

hv_irf_f12 <-
  hv_f12 |>
  irf(n.ahead = 40) |>
  _$irf |>
  as.data.table() |>
  _[impulse == "Chol_1"] |>
  _[, model := "heteroskedasticity"]

list_rbind(list(
  hv_irf_f12,
  mv_irf
)) |>
  ggplot(aes(
    x = h,
    y = irf,
    color = model
  )) +
  geom_hline(yintercept = 0, color = bg_line_color) +
  geom_line() +
  facet_wrap(
    vars(response),
    ncol = 4,
    scales = "free_y"
  ) +
  scale_color_manual(
    values = c(
      cholesky = cholesky_color,
      heteroskedasticity = hetreg_color,
      mbc = max_share_color
    ),
      labels = c(
        cholesky = "Cholesky",
        heteroskedasticity = "Heteroskedasticity",
        mbc = "Main Business Cycle Shock"
      )
  ) +
  labs(
    x = "Horizon (quarters)",
    y = "Impulse Response"
  ) +
  theme_pres
```

##  Using 3-year Horizon for the VFCI
```{r}
#| echo: false
#| fig-width: 10
#| fig-align: 'center'

cv_hs_f12 <-
  cv_f12 |>
  hs() |>
  _$hs |>
  as.data.table() |>
  _[impulse == "Chol_1"] |>
  merge(copy(data)[, t := .I - 2][,. (t, date)], by = "t") |>
  _[, model := "cholesky"]

hv_hs_f12 <-
  hv_f12 |>
  hs() |>
  _$hs |>
  as.data.table() |>
  _[impulse == "Chol_1"] |>
  merge(copy(data)[, t := .I - 2][,. (t, date)], by = "t") |>
  _[, model := "heteroskedasticity"]

p1 <-
  list_rbind(list(
    hv_hs_f12,
    mv_hs
  )) |>
  {\(.x)
    ggplot() +
    geom_rect(
        data = rec_dates[start %between% as.Date(c("1962-01-01", "2022-01-01"))],
        aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf),
        fill = recession_color
      ) +
    geom_hline(yintercept = 0, color = bg_line_color) +
    geom_line(
      data = .x,
      aes(
      x = date,
      y = hs,
      color = model
    )) +
    scale_color_manual(
      values = c(
        cholesky = cholesky_color,
        heteroskedasticity = hetreg_color,
        mbc = max_share_color
      ),
      labels = c(
        cholesky = "Cholesky",
        heteroskedasticity = "Heteroskedasticity",
        mbc = "Main Business Cycle Shock"
      )
    ) +
    labs(
      x = NULL,
      y = "Historical Shock",
      title = "(a) Historical Shock"
    ) +
    theme_pres
  }()


cv_hd_f12 <-
  cv_f12 |>
  fevdid::hd() |>
  _$hd |>
  as.data.table() |>
  _[impulse == "Chol_1"] |>
  _[response == "unemployment"] |>
  merge(copy(data)[, t := .I - 2][,. (t, date)], by = "t") |>
  _[, model := "cholesky"]

hv_hd_f12 <-
  hv_f12 |>
  fevdid::hd() |>
  _$hd |>
  as.data.table() |>
  _[impulse == "Chol_1"] |>
  _[response == "unemployment"] |>
  merge(copy(data)[, t := .I - 2][,. (t, date)], by = "t") |>
  _[, model := "heteroskedasticity"]


p2 <-
  list_rbind(list(
    hv_hd_f12,
    mv_hd
  )) |>
  {\(.x)
    ggplot() +
    geom_rect(
          data = rec_dates[start %between% as.Date(c("1962-01-01", "2022-01-01"))],
          aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf),
          fill = recession_color
        ) +
    geom_hline(yintercept = 0, color = bg_line_color) +
    geom_line(
      data = .x,
      aes(
      x = date,
      y = hd,
      color = model
    )) +
    scale_color_manual(
      values = c(
        cholesky = cholesky_color,
        heteroskedasticity = hetreg_color,
        mbc = max_share_color
      ),
      labels = c(
        cholesky = "Cholesky",
        heteroskedasticity = "Heteroskedasticity",
        mbc = "Main Business Cycle Shock"
      )
    ) +
    labs(
      x = NULL,
      y = "Unemployment",
      title = "(b) Historical Shock Contribution to Unemployment"
    ) +
    theme_pres
  }()

p1 / p2
```



## Implications



# Appendix {.appendix visibility="uncounted"}


## References {visibility="uncounted"}

::: {#refs}
:::

