---
title: "The Price of Risk as Driver of Business Cycles"
subtitle: "Brown Economics Faculty Summer Seminar"
title-note: "^1^The views expressed here are the authors’ and are not necessarily representative of the views of the International Monetary Fund, its Management, or its Executive Directors."
author:
  - name: "Tobias Adrian"
    affiliation: "International Monetary Fund^1^"
  - name: "Matthew DeHaven"
    affiliation: Brown University
  - name: "Fernando Duarte"
    affiliation: Brown University
date: "2024-07-03"
callout-appearance: minimal
editor:
  render-on-save: true
format:
  clean-revealjs:
    self-contained: true
    html-math-method: katex
    crossrefs-hover: true
bibliography: references.bib
---

```{r}
#| include: false
library(data.table)
library(tidyfast)
library(vars)
library(svars)
library(fevdid)
library(ggplot2)
library(patchwork)
library(vfciBCHelpers)
```

```{r}
#| include: false

theme_pres <-
    theme_bw(base_size = 16) +
    theme(
        plot.title = element_text(hjust = 0.5),
        axis.text = element_text(size = 12, margin = margin(0, 0, 0, 0)),
        axis.title = element_text(size = 14),
        plot.caption = element_text(size = 11),
        panel.grid = element_blank(),
        plot.background = element_blank(),
        panel.background = element_blank(),
        strip.background = element_blank(),
        legend.key=element_rect(fill = alpha("white", 0)),
        legend.position = "top",
        legend.background = element_blank(),
        legend.title = element_text(size = 14),
        legend.margin = margin(0, 0, 0, 0),
        legend.text = element_text(size = 10),
        plot.margin = margin(10, 10, 1, 10, "pt")
        )

var_order <- c(
    VFCI = "price of risk",
    Unemployment = "unemployment",
    `Fed Funds` = "interest",
    Inflation = "inflation",
    Output = "output",
    Investment = "investment",
    Consumption = "consumption",
    `Hours Worked` = "hours worked",
    `Labor Share` = "labor share",
    TFP = "TFP",
    `Labor Prod.` = "productivity"
)

crossplot_order <- c("Output", "Investment", "Consumption", "Hours Worked", "Fed Funds", "Labor Prod.", "Inflation", "TFP", "VFCI")


## Colors
bg_line_color <- "gray50"
recession_color <- "gray80"

hetreg_color <- "steelblue"
max_share_color <- "darkorange"
```

```{r var_setup}
lags <- 2
bc_freqs <- c( 2 * pi / 32, 2 * pi / 6)

vfci_data <- get_var_data(cols = c(vfci = "vfci_fgr1gdpc1", vfci_10 = "vfci_fgr10gdpc1", mu = "mu_fgr1gdpc1"))

data <- get_var_data(vfci = "vfci_fgr1gdpc1")
data[, t := .I - lags]

v <- fit_var(data[, -c("t", "vfci")], lags = lags)
v_vfci <- fit_var(data[, -"t"], lags = lags)

mv <- id_fevdfd(v, "unemployment", bc_freqs, sign = "neg")
mvv <- id_fevdfd(v_vfci, "unemployment", bc_freqs, sign = "neg")
vv <- id_fevdfd(v_vfci, "vfci", bc_freqs, sign = "pos")
hv <- id_linear_het_reg(v, "consumption", het_reg_lags = 0:1, hetreg_method = "twostep", sign = "pos")
cv <- fevdid::id_ordered_chol(reorder_var(v_vfci, "vfci"))
```

```{r load_recession_dates}
rec_dates <- get_recession_dt()
```

```{r}
pr_data <-
  hv$het_reg$dt |>
  merge(data[, .(t, date)], by = "t") |>
  _[variable == "unemployment"]  
```
## Definitions

-   The *market price of risk* is the equilibrium compensation
required by investors to hold all aggregate risk
    -   If all risk were tradable in the stock market,
$$
\text{market price of risk} = 
  \frac{
      \text{expected stock returns} - \text{risk-free rate}
    }{
      \text{volatility of stock returns}
    }
$$
-   The *business cycle* is the comovement among macroeconomic 
variables over the
1 to 10 year "short run"
    -   Patterns of comovement are qualitatively the same 
    across different cycles

## Main Result: Price of Risk "drives" Business Cycles
-   Empirically, two new patterns
    1.   Price of risk shocks essentially the same as
    "business cycle shock"
    2.   For business cycle variables $y_t$,
    
    $$
      \mathbb{E}[y_{t+1} \mid \text{price of risk}_t] = 
        a + b Vol[y_{t+1} \mid \text{price of risk}_t]
    $$
-   Theoretically,
    -   Empirical patterns inconsistent with mainstream models
    -   But consistent with a model of financial frictions

## Price of Risk and Business Cycle Shocks Match Up
```{r}
#| echo: false
#| 

hd_hv_data <- hd(hv)$hd |>
  as.data.table() |>
  merge(data[, .(t, date)], by = "t") |>
  _[impulse == "Chol_1"] |>
  _[response == "unemployment"]
  
hd_mv_data <- hd(mv)$hd |>
  as.data.table() |>
  merge(data[, .(t, date)], by = "t") |>
  _[impulse == "Main"] |>
  _[response == "unemployment"]

hd_data <- rbind(
  hd_hv_data[, method := "hetreg"],
  hd_mv_data[, method := "max_share"]
)

ggplot() +
  geom_rect(
    data = rec_dates[start %between% as.Date(c("1965-01-01", "2017-01-01"))],
    aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf),
    fill = recession_color
  ) +
  geom_hline(yintercept = 0, color = bg_line_color) + 
  geom_line(
    data = hd_data,
    aes(
      x = date,
      y = hd,
      color = method
    )
  ) +
  labs(
    x = NULL,
    y = "Two-year moving average",
    title = NULL,
    caption = "Gray shaded areas indicate NBER recessions"
  ) +
  scale_x_date(
    breaks = seq(as.Date("1960-01-01"), as.Date("2020-01-01"), by = "10 years"),
    labels = seq(1960, 2020, 10),
    limits = as.Date(c("1960-01-01", "2021-01-01")),
    expand = c(0,0)
  ) + 
  scale_color_manual(
    values = c(hetreg_color, max_share_color),
    labels = c(hetreg = "Price of risk shock", max_share = "Business cycle shock"),
    name = NULL
  ) +
  theme_pres
```
## Impulse Response Functions Also Coincide
```{r}
mv_list <- list(
    max_share_vfci = vv,
    max_share_u = mvv
)

irf_df <- rbindlist(lapply(seq_along(mv_list), function(i) {
  irf <- irf(mv_list[[i]], n.ahead = 40)$irf |> setDT()
  irf$model <- names(mv_list)[[i]]
  rbind(irf[ impulse == "Chol_1" ], irf[ impulse == "Main" ])
}))

irf_df$model <- forcats::fct_rev(as.ordered(irf_df$model))

irf_df |>
  ggplot(aes(
    x = h,
    y = irf,
    color = model
  )) +
  geom_hline(yintercept = 0) +
  geom_line() +
  facet_wrap(
    vars(response),
    nrow = 3,
    scales = "free_y"
  ) +
  labs(
    x = "Quarters",
    y = "IRF",
    title = NULL,
    caption = NULL
  )  +
  scale_color_manual(
    values = c(max_share_vfci = hetreg_color, max_share_u = max_share_color),
    labels = c(max_share_vfci = "Price of risk shock", max_share_u = "Business cycle shock"),
    name = NULL
  ) +
  theme_bw(base_size = 8) +
  theme(legend.position = "top") +
  theme_pres +
  theme(axis.text = element_text(size = 10, margin = margin(0, 0, 0, 0))) 
```
## Linear Relation for Conditional Means and Volatilities
```{r}
var_means <-
  fitted(v) |>
  as.data.table() |> 
  _[, t := .I] |>
  tidyfast::dt_pivot_longer(-t, names_to = "variable", values_to = "var_fitted")

mean_vol_data <- 
  hv$het_reg$dt |> 
  merge(var_means, by = c("t", "variable"))

mean_vol_data |>
  _[variable == "unemployment"] |>
  ggplot(aes(y = var_fitted, x = scale(log_var_fitted) )) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_point() +
  theme_pres +
  labs(
    x = "Fitted Log-Volatility",
    y = "Fitted Mean",
    title = "Unemployment"
  ) +
  theme(aspect.ratio = 1)
```
## Business cycles

### Causes
-   Empirically, just one shock captures 50-75\% of variation of many variables
-   In models, usually a single big shock: TFP, beliefs, confidence, etc.
-   Common propagation can look like a single shock

### Why we care
-   Causes determine whether fluctuations are inefficient
-   Crucial for welfare and policy

::: {.notes}
The combination of expansions and recessions, the ebb and flow of economic activity, is called the business cycle.
In many ways, the term “business cycle” is misleading. “Cycle” seems to imply that there is some regularity in the timing and duration of upswings and downswings in economic activity. Most economists, however, do not think there is, better to say
“short-run economic fluctuations” 
:::

## Empirical Framework: Vector Autoregression (VAR)

Structural VAR($p$) with $k$ variables in vector $y_t$,
$$
B_0 y_t = B_1 y_{t-1} + . . . + B_p y_{t-p} + w_t
$$
where structural shocks $w_t$ are white noise

. . .

Estimate reduced-form coefficients $A_i$ and residuals $u_t$
$$
 y_t = \underbrace{B_0^{-1}B_1}_{A_1} y_{t-1} + \dots + \underbrace{B_0^{-1}B_p}_{A_p} y_{t-p} + \underbrace{B_0^{-1}w_t}_{u_t}
$$

## VAR Specification is Standard

```{r}
md_start <- as.yearqtr(min(data$date))
md_end <- as.yearqtr(max(data$date))
md_vars <- paste0(var_order[names(var_order) != "VFCI"], sep = ", ", collapse="")
md_vars2 <- substr(md_vars, 1, nchar(md_vars)-2)
```

-   `r md_start` to `r md_end` with $p=2$ lags
-   $k=10$ macro variables: `r md_vars2`
-   Or, $k=11$ with price of risk
    -   Use "volatility financial conditions index"
    <!-- -   Results robust to using other broad-based financial conditions indices  -->
    <!-- -   But not robust to using single asset class -->
    <!-- (e.g., just stocks, just bonds) -->

## Volatility Financial Conditions Index (VFCI)
:::: {.columns style='display: flex !important; height: 80%;'}

::: {.column width="25%" style='display: flex; justify-content: left; align-items: top;; font-size: 20pt'}
-   VFCI is the volatility of consumption spanned by financial assets
-   VFCI is a linear combination of stock returns, Treasury term and liquidity 
    spreads, corporate credit and default spreads
-   Measures price of risk of representative agent in large class of models
-   Can be interpreted as tightness in financial conditions
:::

::: {style='display: flex; justify-content: right; align-items: center;'}
```{r}
ggplot() +
  geom_rect(
    data = rec_dates[start %between% as.Date(c("1965-01-01", "2017-01-01"))],
    aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf),
    fill = recession_color
  ) +
  geom_hline(yintercept = 0, color = bg_line_color) + 
  geom_line(data = vfci_data, aes(x = date, y = scale(vfci))) +  
  scale_x_date(
    breaks = seq(as.Date("1960-01-01"), as.Date("2020-01-01"), by = "10 years"),
    labels = seq(1960, 2020, 10),
    limits = as.Date(c("1960-01-01", "2021-01-01")),
    expand = c(0,0)
  ) +
  labs(
    x = NULL,
    y = NULL,
    subtitle = "Index with zero mean and unit variance",
    title = "Volatility Financial Conditions Index",
    caption = "Gray shaded areas indicate NBER recessions"
  ) +
  theme_pres
```
:::

::::
## Usual Identification Problem
-   Compare VAR in structural and reduced form
$$
\begin{align*}
B_0 y_t &= B_1 y_{t-1} + . . . + B_p y_{t-p} + w_t \\
 y_t &= \underbrace{B_0^{-1}B_1}_{A_1} y_{t-1} + \dots + \underbrace{B_0^{-1}B_p}_{A_p} y_{t-p} + \underbrace{B_0^{-1}w_t}_{u_t}
 \end{align*}
$$
-   Structural shocks $w_t$ related to reduced-form residuals $u_t$ by
$$
w_t = B_0 u_t
$$
but $B_0$ not identified

## A Main Business Cycle Shock
-   Identify a "main business-cycle" shock $w_t^{(\text{MBC})}$ as in @mbc2020
    1.    Pick a target variable (e.g., unemployment)
    2.    Find linear combination of residuals
    $$
      w_t^{(\text{MBC})} = b^T u_t
    $$
    that maximizes
    contribution of MBC shock to the volatility of target variable
    over business-cycle frequencies

## Same Business Cycle Shock for Different Targets
-   MBC shock explains 50-75% of volatility of unemployment, investment, output,
    hours worked
-   Targeting unemployment, output, hours worked, consumption, or investment
    identifies same shock
-   Shock looks like demand shock without inflation
-   @mbc2020 propose a "confidence" shock (higher-order belief shocks) as 
    explanation
    
## Price of Risk Shocks
-   We construct three price of risk shocks
    1.    Max-share: same as MBC shock but target price of risk
    2.    Timing: only shock that affects price of risk contemporaneously
    3.    Heteroskedasticity: shock to the volatility of consumption within
    the VAR
-   All three look essentially the same as main business cycle shock
<!-- -   Correlation with MBC shock $0.6-0.75$ -->

## Estimation of Volatility of Consumption    
-   Assume conditional volatility of consumption residuals is linear in
VAR variables
-   Two-step estimation
    -   Estimate VAR and define $\sigma_t := \hat{u}_t^{(\text{cons})}$
    -   Run OLS regression
    $$
    \log \sigma_t^2 = c^T y_t + \vartheta_t
    $$
-   Use $\hat{c}^T y_t$

## IRF of Max-Share Price of Risk Shock
```{r}
mv_list <- list(
    max_share_vfci = vv,
    max_share_u = mvv
)

irf_df <- rbindlist(lapply(seq_along(mv_list), function(i) {
  irf <- irf(mv_list[[i]], n.ahead = 40)$irf |> setDT()
  irf$model <- names(mv_list)[[i]]
  rbind(irf[ impulse == "Chol_1" ], irf[ impulse == "Main" ])
}))

irf_df$model <- forcats::fct_rev(as.ordered(irf_df$model))

irf_df |>
  ggplot(aes(
    x = h,
    y = irf,
    color = model
  )) +
  geom_hline(yintercept = 0) +
  geom_line() +
  facet_wrap(
    vars(response),
    nrow = 3,
    scales = "free_y"
  ) +
  labs(
    x = "Quarters",
    y = "IRF",
    title = NULL,
    caption = NULL
  )  +
  scale_color_manual(
    values = c(max_share_vfci = hetreg_color, max_share_u = max_share_color),
    labels = c(max_share_vfci = "Price of risk shock", max_share_u = "Business cycle shock"),
    name = NULL
  ) +
  theme_bw(base_size = 8) +
  theme(legend.position = "top") +
  theme_pres +
  theme(axis.text = element_text(size = 10, margin = margin(0, 0, 0, 0))) 
```


## IRF of Timing Price of Risk Shock
```{r}
mv_list <- list(
    chol_vfci = cv,
    max_share_u = mvv
)

irf_df <- rbindlist(lapply(seq_along(mv_list), function(i) {
  irf <- irf(mv_list[[i]], n.ahead = 40)$irf |> setDT()
  irf$model <- names(mv_list)[[i]]
  rbind(irf[ impulse == "Chol_1" ], irf[ impulse == "Main" ])
}))

irf_df$model <- forcats::fct_rev(as.ordered(irf_df$model))

irf_df |>
  ggplot(aes(
    x = h,
    y = irf,
    color = model
  )) +
  geom_hline(yintercept = 0) +
  geom_line() +
  facet_wrap(
    vars(response),
    nrow = 3,
    scales = "free_y"
  ) +
  labs(
    x = "Quarters",
    y = "IRF",
    title = NULL,
    caption = NULL
  )  +
  scale_color_manual(
    values = c(chol_vfci = hetreg_color, max_share_u = max_share_color),
    labels = c(chol_vfci = "Price of risk shock", max_share_u = "Business cycle shock"),
    name = NULL
  ) +
  theme_bw(base_size = 8) +
  theme(legend.position = "top") +
  theme_pres +
  theme(axis.text = element_text(size = 10, margin = margin(0, 0, 0, 0))) 
```


## IRF of Heteroskedasticity Price of Risk Shock
```{r}
mv_list <- list(
    het_reg_vfci = hv,
    max_share_u = mvv
)

irf_df <- rbindlist(lapply(seq_along(mv_list), function(i) {
  irf <- irf(mv_list[[i]], n.ahead = 40)$irf |> setDT()
  irf$model <- names(mv_list)[[i]]
  rbind(irf[ impulse == "Chol_1" ], irf[ impulse == "Main" ])
}))

irf_df$model <- forcats::fct_rev(as.ordered(irf_df$model))

irf_df |>
  ggplot(aes(
    x = h,
    y = irf,
    color = model
  )) +
  geom_hline(yintercept = 0) +
  geom_line() +
  facet_wrap(
    vars(response),
    nrow = 3,
    scales = "free_y"
  ) +
  labs(
    x = "Quarters",
    y = "IRF",
    title = NULL,
    caption = NULL
  )  +
  scale_color_manual(
    values = c(het_reg_vfci = hetreg_color, max_share_u = max_share_color),
    labels = c(het_reg_vfci = "Price of risk shock", max_share_u = "Business cycle shock"),
    name = NULL
  ) +
  theme_bw(base_size = 8) +
  theme(legend.position = "top") +
  theme_pres +
  theme(axis.text = element_text(size = 10, margin = margin(0, 0, 0, 0))) 
```
 
## Mean and Vol Linked by Price of Risk
```{r}
#| fig-align: center
vfci_data |>
  ggplot(aes(
    x = vfci,
    y = mu
  )) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_point() +
  theme_pres + 
  theme(aspect.ratio = 1) +
  labs(
    y = "Conditional Mean",
    x = "Condtional Volatility\n(Consumption)",
    title = NULL
  )
```

## Mean and Vol Linked by Price of Risk
```{r}
var_means <-
  fitted(v) |>
  as.data.table() |> 
  _[, t := .I] |>
  tidyfast::dt_pivot_longer(-t, names_to = "variable", values_to = "var_fitted")

mean_vol_data <- 
  hv$het_reg$dt |> 
  merge(var_means, by = c("t", "variable"))

mean_vol_data |>
  _[variable == "unemployment"] |>
  ggplot(aes(y = var_fitted, x = scale(log_var_fitted) )) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_point() +
  theme_pres +
  labs(
    y = "Conditional Mean",
    x = "Condtional Volatility\n(Unemployment)",
    title = NULL
  ) +
  theme(aspect.ratio = 1)
```
## Conclusion

### A New Business Cycle Fact

We show a new feature of business cycles

-   the price of risk comoves with the mean and vol of macro varaibles during the cycle
-   not consistent with real business cycle (RBC) or New Keynesian models

## Conclusion

### Financial Conditions Propagate the Business Cycle

There exists a general equilibrium structural model that is consistent with

-   the price of risk comoving with the business cycle
-   the empirical mean-vol relationship of macro variables


In the model, financial sector aggregates, amplifies, and propagates
disparate shocks

-   Driven by a financial friction: the value-at-risk constraint
-   Not a single shock but a story of strong propagation

# Appendix {.appendix visibility="uncounted"}

## References

::: {#refs}
:::