---
title: "BCA Replication: Bayesian VAR IRFs"
author: Matthew DeHaven
date: July 31, 2023
output: 
    pdf_document
---
```{r setup, include = FALSE}
require(here)
require(data.table)
require(dplyr)
require(ggplot2)
```

# What is being replicated
The goal is to replicate Figure 1 of ``Business Cycle Anatomy'', copied below.

```{r fig20, echo = F, fig.cap="Impulse Response Functions to the MBC Shock", out.width = "100%"}
knitr::include_graphics(here("./data-raw/figs-raw/bca-replication/figure1.pdf"))
```

The black line shows the median Bayesian posterior distribution of the IRF at that horizon.
The shaded area shows the 16th to 84th percentile of that distribution.

# Replication
```{r loadData, include = F}
df <- fread(here("./data/replicated_bca_bayesian_VAR_IRF.csv"))
df[, variable := factor(variable, levels = c(
    "unemployment", "output", "hours_worked", "investment", "consumption", "TFP", "productivity", "labor_share", "inflation", "interest"
), ordered = TRUE)]
df[model == "bayesian_fd", model := "Original BCA, Frequency Domain"]
df[model == "bayesian_td4", model := "Original BCA, Time Domain, Horizon 4"]
df[model == "bayesian_td632", model := "Original BCA, Time Domain, Horizon 6:32 (act. 32)"]
```

```{r plotRepBayFD, echo = F, warning = F, fig.cap = "Replication of Bayesian VAR targetting Frequency Domain", fig.height = 2.5, out.width = "100%"}
df[model %in% c("Replication, Frequency Domain", "Original BCA, Frequency Domain")] |>
    ggplot(aes(
        x = horizon,
        color = model,
        fill = model
    )) +
    geom_hline(yintercept = 0) +
    geom_ribbon(aes(
        ymin = lower,
        ymax = upper
        ),
        alpha = 0.3,
        color = NA
        ) +
    geom_line(aes(
        y = median
    )) +
    labs(y = NULL, x = "Horizon") +
    scale_color_manual(values = c("firebrick", "steelblue")) +
    scale_fill_manual(values = c("firebrick", "steelblue")) +
    facet_wrap(
        vars(variable),
        scales = "free_y",
        nrow = 2
        ) +
    theme_bw(base_size = 8) +
    theme(legend.position = "bottom", legend.margin = margin(-10, 0, 0, 0, "pt"), strip.background = element_blank(), panel.grid.minor = element_blank())
```

\newpage

# Time Domain Targetting

In addition, the original paper compares targetting the time domain to the frequency domain.

They argue that targetting a horizon of 4 is the most similar to the frequency domain result from above.
The below plot compares a replication of a Bayesian VAR targetting a horizon of 4.

```{r plotRepBayTD4, echo = F, warning = F, fig.cap = "Replication of Bayesian VAR targetting Time Domain, Horizon 4", fig.height = 2.5, out.width = "100%"}
df[model %in% c("Replication, Time Domain, Horizon 4", "Original BCA, Time Domain, Horizon 4")] |>
    ggplot(aes(
        x = horizon,
        color = model,
        fill = model
    )) +
    geom_hline(yintercept = 0) +
    geom_ribbon(aes(
        ymin = lower,
        ymax = upper
        ),
        alpha = 0.3,
        color = NA
        ) +
    geom_line(aes(
        y = median
    )) +
    labs(y = NULL, x = "Horizon") +
    scale_color_manual(values = c("firebrick", "steelblue")) +
    scale_fill_manual(values = c("firebrick", "steelblue")) +
    facet_wrap(
        vars(variable),
        scales = "free_y",
        nrow = 2
        ) +
    theme_bw(base_size = 8) +
    theme(legend.position = "bottom", legend.margin = margin(-10, 0, 0, 0, "pt"), strip.background = element_blank(), panel.grid.minor = element_blank())
```

The authors argue that targetting a horizon of 6 to 32 (to match the frequencies of 2 pi / 6 tp 2 pi /32) would be naive and does not give the closest matched IRF.
However, their code incorrectly targets only the 32nd horizon, not the full range.
Thus in comparing the replication below, I target the 32nd horizon.

```{r plotRepBayTD32, echo = F, warning = F, fig.cap = "Replication of Bayesian VAR targetting Time Domain, Horizon 32", fig.height = 2.5, out.width = "100%"}
df[model %in% c("Replication, Time Domain, Horizon 32", "Original BCA, Time Domain, Horizon 6:32 (act. 32)")] |>
    ggplot(aes(
        x = horizon,
        color = model,
        fill = model
    )) +
    geom_hline(yintercept = 0) +
    geom_ribbon(aes(
        ymin = lower,
        ymax = upper
        ),
        alpha = 0.3,
        color = NA
        ) +
    geom_line(aes(
        y = median
    )) +
    labs(y = NULL, x = "Horizon") +
    scale_color_manual(values = c("firebrick", "steelblue")) +
    scale_fill_manual(values = c("firebrick", "steelblue")) +
    facet_wrap(
        vars(variable),
        scales = "free_y",
        nrow = 2
        ) +
    theme_bw(base_size = 8) +
    theme(legend.position = "bottom", legend.margin = margin(-10, 0, 0, 0, "pt"), strip.background = element_blank(), panel.grid.minor = element_blank())
```