---
title: "BCA Replication: Classical VAR IRFs, Bootsstrapped"
author: Matthew DeHaven
date: July 25, 2023
output: 
    pdf_document
---
```{r setup, include = FALSE}
require(here)
require(data.table)
require(dplyr)
require(ggplot2)
```

# Replication of the Bootstrapped VARs

```{r loadData, include = F}
comb_df <- fread(here("./data/replicated_bca_classical_VAR_IRF_boot.csv")) |>
    mutate(model = ifelse(model == "classical_fd", "BCA Original", model))
```

Using a resampled bootstrap method (drawing from VAR errors with replacement), the bootstrapped IRFs look very close to the original BCA IRFs.

```{r figBootRes, echo = F, warning = F, fig.cap="Comparison of Replicated IRF to Original, Resample Bootstrap", fig.height = 2.5, out.width = "100%"}
comb_df |>
    filter(model %in% c("Bootstrap Replication Resample", "BCA Original")) |>
    ggplot(aes(
        x = h,
        color = model,
        fill = model
    )) +
    geom_ribbon(aes(ymin = lower, ymax = upper), color = NA, alpha = 0.3) +
    geom_line(aes(y = median)) +
    facet_wrap(vars(variable), nrow = 2, scales = "free_y") +
    scale_color_manual(values = c("firebrick", "steelblue")) +
    scale_fill_manual(values = c("firebrick", "steelblue")) +
    scale_x_continuous(limits = c(0, 20)) +
    labs(x = "Horizon", y = NULL) +
    theme_gray(base_size = 8) +
    theme(legend.position = "bottom", strip.background = element_blank(), panel.grid.minor = element_blank())
```


The original BCA paper also bias-adjusts their bootstraps, through a method called bootstrap-after-bootstrap.
Implementing this makes the replicated IRFs even close to the original BCA IRFs.

```{r figBootResBA, echo = F, warning = F, fig.cap="Comparison of Replicated IRF to Original, Resample Bootstrap, Bias adjusted", fig.height = 2.5, out.width = "100%"}
comb_df |>
    filter(model %in% c("Bootstrap Replication Resample Bias Adjust", "BCA Original")) |>
    ggplot(aes(
        x = h,
        color = model,
        fill = model
    )) +
    geom_ribbon(aes(ymin = lower, ymax = upper), color = NA, alpha = 0.3) +
    geom_line(aes(y = median)) +
    facet_wrap(vars(variable), nrow = 2, scales = "free_y") +
    scale_color_manual(values = c("firebrick", "steelblue")) +
    scale_fill_manual(values = c("firebrick", "steelblue")) +
    scale_x_continuous(limits = c(0, 20)) +
    labs(x = "Horizon", y = NULL) +
    theme_gray(base_size = 8) +
    theme(legend.position = "bottom", strip.background = element_blank(), panel.grid.minor = element_blank())
```

### Comparison to Wild Bootsstrap

Using the Wild bootstrap with the gaussian distribution (multiply the errors by a random vector drawn from N(0,1)), the impulse responses are very different from the original BCA results.

```{r figBootWildG, echo = F, warning = F, fig.cap="Comparison of Replicated IRF to Original, Wild Bootstrap, Gaussian, nboot = 1000", fig.height = 2.5, out.width = "100%"}
comb_df |>
    filter(model %in% c("Bootstrap Replication Wild Gaussian", "BCA Original")) |>
    ggplot(aes(
        x = h,
        color = model,
        fill = model
    )) +
    geom_ribbon(aes(ymin = lower, ymax = upper), color = NA, alpha = 0.3) +
    geom_line(aes(y = median)) +
    facet_wrap(vars(variable), nrow = 2, scales = "free_y") +
    scale_color_manual(values = c("firebrick", "steelblue")) +
    scale_fill_manual(values = c("firebrick", "steelblue")) +
    scale_x_continuous(limits = c(0, 20)) +
    labs(x = "Horizon", y = NULL) +
    theme_gray(base_size = 8) +
    theme(legend.position = "bottom", strip.background = element_blank(), panel.grid.minor = element_blank())
```

This remains true even if the number of bootstrap iterations are doubled. 
There is almost no different between 1000 and 2000 iterations.

```{r figBootWildG2000, echo = F, warning = F, fig.cap="Comparison of Replicated IRF to Original, Wild Bootstrap, Gaussian, nboot = 2000", fig.height = 2.5, out.width = "100%"}
comb_df |>
    filter(model %in% c("Bootstrap Replication Wild Gaussian 2000", "BCA Original")) |>
    ggplot(aes(
        x = h,
        color = model,
        fill = model
    )) +
    geom_ribbon(aes(ymin = lower, ymax = upper), color = NA, alpha = 0.3) +
    geom_line(aes(y = median)) +
    facet_wrap(vars(variable), nrow = 2, scales = "free_y") +
    scale_color_manual(values = c("firebrick", "steelblue")) +
    scale_fill_manual(values = c("firebrick", "steelblue")) +
    scale_x_continuous(limits = c(0, 20)) +
    labs(x = "Horizon", y = NULL) +
    theme_gray(base_size = 8) +
    theme(legend.position = "bottom", strip.background = element_blank(), panel.grid.minor = element_blank())
```

\newpage
Another distribution can be used in the Wild bootstrap method, "Rademacher" simply multiplies the VAR errors by a random vector of -1s and 1s.

```{r figBootWildR, echo = F, warning = F, fig.cap="Comparison of Replicated IRF to Original, Wild Bootstrap, Rademacher", fig.height = 2.5, out.width = "100%"}
comb_df |>
    filter(model %in% c("Bootstrap Replication Wild Rademacher", "BCA Original")) |>
    ggplot(aes(
        x = h,
        color = model,
        fill = model
    )) +
    geom_ribbon(aes(ymin = lower, ymax = upper), color = NA, alpha = 0.3) +
    geom_line(aes(y = median)) +
    facet_wrap(vars(variable), nrow = 2, scales = "free_y") +
    scale_color_manual(values = c("firebrick", "steelblue")) +
    scale_fill_manual(values = c("firebrick", "steelblue")) +
    scale_x_continuous(limits = c(0, 20)) +
    labs(x = "Horizon", y = NULL) +
    theme_gray(base_size = 8) +
    theme(legend.position = "bottom", strip.background = element_blank(), panel.grid.minor = element_blank())
```

In order to deal with this we can choose to set the sign of the identified VARs by the sign of the cummulative IRF of the target variable out to `n` quarters.
Effectively, by setting the sign based off of impact, we do this out to `n = 1`.
If instead we set it out to `n = 10`, then we allow for negative IRFs on impact but that become strongly positive shortly thereafter.
These bootstraps look much better.

```{r figBootWildGsh, echo = F, warning = F, fig.cap="Comparison of Replicated IRF to Original, Wild Bootstrap, Gaussian, Sign Horizon", fig.height = 2.5, out.width = "100%"}
comb_df |>
    filter(model %in% c("Bootstrap Replication Wild Gaussian, Sign Horizon", "BCA Original")) |>
    ggplot(aes(
        x = h,
        color = model,
        fill = model
    )) +
    geom_ribbon(aes(ymin = lower, ymax = upper), color = NA, alpha = 0.3) +
    geom_line(aes(y = median)) +
    facet_wrap(vars(variable), nrow = 2, scales = "free_y") +
    scale_color_manual(values = c("firebrick", "steelblue")) +
    scale_fill_manual(values = c("firebrick", "steelblue")) +
    scale_x_continuous(limits = c(0, 20)) +
    labs(x = "Horizon", y = NULL) +
    theme_gray(base_size = 8) +
    theme(legend.position = "bottom", strip.background = element_blank(), panel.grid.minor = element_blank())
```