---
title: "BCA Replication: Classical VAR IRFs, Bootsstrapped"
author: Matthew DeHaven
date: July 25, 2023
output: 
    pdf_document
---
```{r setup, include = FALSE}
require(here)
require(data.table)
require(dplyr)
require(ggplot2)
```

# Replication of the Bootstrapped VARs

```{r loadData, include = F}
comb_df <- fread(here("./data/replicated_bca_classical_VAR_IRF_boot.csv")) |>
    mutate(
        response = factor(response, levels = c("unemployment", "output", "hours_worked", "investment", "consumption", "TFP", "productivity", "labor_share", "inflation", "interest"), ordered = T)
    )
```

Using a resampled bootstrap method (drawing from VAR errors with replacement), the bootstrapped IRFs look very close to the original BCA IRFs.

The original BCA paper also bias-adjusts their bootstraps, through a method called bootstrap-after-bootstrap.
Implementing this makes the replicated IRFs almost exact to the original BCA IRFs.

```{r figBootResBA, echo = F, warning = F, fig.cap="Comparison of Replicated IRF to Original, Resample Bootstrap, Bias adjusted", fig.height = 3, out.width = "100%"}
comb_df |>
    ggplot(aes(
        x = h,
        color = version,
        fill = version
    )) +
    geom_hline(yintercept = 0) +
    geom_ribbon(aes(ymin = lower, ymax = upper), color = NA, alpha = 0.3) +
    geom_line(aes(y = median)) +
    facet_wrap(vars(response), nrow = 2, scales = "free_y") +
    scale_color_manual(values = c("firebrick", "steelblue")) +
    scale_fill_manual(values = c("firebrick", "steelblue")) +
    scale_x_continuous(limits = c(0, 20)) +
    labs(x = "Horizon", y = NULL) +
    theme_bw(base_size = 8) +
    theme(legend.position = "bottom", strip.background = element_blank(), panel.grid.minor = element_blank())
```
