---
title: "Mean - Vol Relationship"
output: 
    pdf_document
---

```{r setup, include = FALSE}
require(here)
require(data.table)
require(tidyfast)
require(dplyr)
require(ggplot2)
require(patchwork)
require(vars)
require(svars)
require(fevdid)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```


\section{VAR}

A reduced form VAR of some number of variables, $k$, and  some number of lags, $p$, is written as,
\begin{align}
y_t = A_1 y_{t-1} + ... + A_{p} y_{t-p} + u_t
\end{align}

where $A_i$ is a $kxk$ matrix of coefficients and,
\begin{align}
  y_t = \begin{bmatrix}\text{output}_t\\\text{unemployment}_t\\\text{inflation}_t\\...\\\text{vfci}_t\end{bmatrix}, \hspace{1cm} u_t = \begin{bmatrix}u^\text{output}_t\\u^\text{unemployment}_t\\u^\text{inflation}_t\\...\\u^\text{vfci}_t\end{bmatrix}
\end{align}

Each $u_t$ are the reduced form residuals for the accompanying data series.

With this, we can estimate the predicted values, $\widehat{y}_t$, and the residuals, $\widehat{u}_t$.

\subsubsection{Volatility}

We can then define the log-variance as 
\begin{align}
  \text{Var}_t \equiv \log(\widehat{u}_t^2)
\end{align}

This can be estimated with the regression,
\begin{align}
  \text{Var}_t = \alpha_1 y_{t-1} + . . . + \alpha_{p} y_{t-p} + \epsilon_t
\end{align}

Then the VFCI is defined as,
\begin{align}
  \text{VFCI}_t \equiv \widehat{\text{Var}_t}
\end{align}

The VFCI series is then rescaled to N(0,1).

\subsection{External VFCI}
We will compare to the externally estimated VFCI on forward GDP growth (forwarded 1 quarter) and financial principal components.

\newpage



```{r, include = F}
all_data <- fread(here("./data/vfciBC_data.csv"))
X <- all_data[,.(date, output, investment, consumption, hours_worked, unemployment, labor_share, interest, inflation, productivity, TFP)]
X <- X[date <= as.Date("2017-01-01")]

lags <- 2
bc_freqs <- c(2 * pi / 32, 2 * pi / 6)

v <- VAR(X[, -"date"], p = lags, type = "const")

X[, t := seq_len(nrow(X)) - lags]
```

```{r}
data <- copy(X)

data <- data |>
  dt_pivot_longer(-c(date, t), names_to = "variable", values_to = "original")

resid_data <-
  residuals(v) |>
  as.data.table() |>
  _[, t := .I] |>
  dt_pivot_longer(-t, names_to = "variable", values_to = "residuals")

fitted_data <-
  fitted(v) |>
  as.data.table() |>
  _[, t := .I] |>
  dt_pivot_longer(-t, names_to = "variable", values_to = "fitted")

data <- data |>
  merge(resid_data, by = c("t", "variable")) |>
  merge(fitted_data, by = c("t", "variable"))

data[, log_var := log(residuals^2)]
```

```{r}
reg_data <- data |>
  merge(
    dt_pivot_wider(data[,.(t, variable, original)], names_from = variable, values_from = original),
    by = c("t"),
    allow.cartesian = TRUE
  )

fitted_log_var_data <-
  rbindlist(lapply(unique(reg_data$variable), function(n) {
    het_model <- lm(formula = "log_var ~ 
    lag(output) + lag(investment) + lag(consumption) + lag(hours_worked) + lag(unemployment) + 
    lag(labor_share) + lag(interest) + lag(inflation) + lag(productivity) + lag(TFP) +
    lag(output, 2) + lag(investment, 2) + lag(consumption, 2) + lag(hours_worked, 2) + lag(unemployment, 2) + 
    lag(labor_share, 2) + lag(interest, 2) + lag(inflation, 2) + lag(productivity, 2) + lag(TFP, 2)
    ", data = reg_data[variable == n])
    fitted(het_model) |>
      as.data.table() |>
      setnames("V1", "fitted_log_var") |>
      _[, variable := n] |>
      _[, t := .I] |>
      _[, resid_log_var := residuals(het_model)]
  }))

data <- data |>
  merge(fitted_log_var_data, by = c("t", "variable"))
```


```{r}
data[, var_vfci := scale(fitted_log_var), by = variable]
#data[, var_vfci := scale(sqrt(exp(fitted_log_var))), by = variable]

## Flip variables with reverse mean-vol relationship
data[variable == "unemployment", var_vfci := -1 * var_vfci]


## Merge on vfci data
ext_vfci_data <- all_data[, .(date, vfci_ext = vfci_fgr1gdpc1)]
ext_vfci_data[, vfci_ext := scale(vfci_ext)]

data <- data |>
  merge(ext_vfci_data, by = "date")

```


```{r, include = TRUE, fig.height = 9, out.width = "100%"}
data |>
  ggplot(aes(
    x = date,
    y = var_vfci,
    color = variable
  )) +
  geom_line(aes(color = "vfci_var")) +
  geom_line(aes(y = vfci_ext, color = "vfci_ext")) +
  facet_wrap(
    vars(variable),
    ncol = 2,
    scales = "free"
  ) +
  scale_color_manual(
    values = c(
      vfci_var = "firebrick",
      vfci_ext = "black"
    ),
    labels = c(
      vfci_var = "VFCI from VAR (one for each variable)",
      vfci_ext = "External VFCI"
    )
  ) +
  theme_bw(base_size = 8) + 
  theme(legend.position = "top")
```


\newpage

```{r, include = TRUE, fig.height = 9, out.width = "100%"}
corrs <- data[, .(corr = cor(var_vfci, vfci_ext)), by = variable]

data |>
  ggplot(aes(
    x = vfci_ext,
    y = var_vfci
  )) +
  geom_hline(yintercept = 0, color = "gray50") +
  geom_vline(xintercept = 0, color = "gray50") +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, fullrange = TRUE) +
  geom_label(data = corrs, aes(x = - 4, y = 3.5, label = paste0("Corr: ", round(corr, 2))), hjust = 0) +
  facet_wrap(
    vars(variable),
    ncol = 3
  ) +
  scale_x_continuous(limits = c(-4, 4), expand = c(0,0)) +
  scale_y_continuous(limits = c(-4, 4), expand = c(0,0)) +
  theme_bw(base_size = 8) +
  coord_fixed()
```