---
title: "Mean - Vol Relationship"
output: 
    pdf_document
---

```{r setup, include = FALSE}
require(here)
require(data.table)
require(tidyfast)
require(dplyr)
require(ggplot2)
require(patchwork)
require(vars)
require(svars)
require(fevdid)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```


\section{VAR}

A reduced form VAR of some number of variables, $k$, and  some number of lags, $p$, is written as,
\begin{align}
y_t = A_1 y_{t-1} + ... + A_{p} y_{t-p} + u_t
\end{align}

where $A_i$ is a $kxk$ matrix of coefficients and,
\begin{align}
  y_t = \begin{bmatrix}\text{output}_t\\\text{unemployment}_t\\\text{inflation}_t\\...\\\text{vfci}_t\end{bmatrix}, \hspace{1cm} u_t = \begin{bmatrix}u^\text{output}_t\\u^\text{unemployment}_t\\u^\text{inflation}_t\\...\\u^\text{vfci}_t\end{bmatrix}
\end{align}

Each $u_t$ are the reduced form residuals for the accompanying data series.

With this, we can estimate the predicted values, $\widehat{y}_t$, and the residuals, $\widehat{u}_t$.

\subsubsection{Volatility}

We can then define the log-variance as 
\begin{align}
  \text{Var}_t \equiv \log(\widehat{u}_t^2)
\end{align}

This can be estimated with the regression,
\begin{align}
  \text{Var}_t = \alpha_1 y_{t-1} + . . . + \alpha_{p} y_{t-p} + \epsilon_t
\end{align}

Then the predicted (conditional) volatility is
\begin{align}
  \widehat{\text{Vol}}_t = \left[ \exp(\widehat{\text{Var}_t}) \right]^{\frac{1}{2}}
\end{align}

\subsubsection{Mean-Vol Relationship}

The mean-vol relationship is then the relationship between

(1) the predicted means:

\begin{align}
  \widehat{y}_{t}
\end{align}

(2) the predicted volatility:

\begin{align}
  \widehat{\text{Vol}}_t
\end{align}

\textbf{Note:} There are $k$ mean-vol relationships, one for each variable in the VAR.

\newpage

\section{Fitting the VAR}

(1) Fit the VAR to the raw (original) data, find the fitted values and the residuals.

```{r, include = F}
X <- fread(here("./data/vfciBC_data.csv"))
X <- X[,.(date, output, investment, consumption, hours_worked, unemployment, labor_share, interest, inflation, productivity, TFP, vfci = vfci_fgr10gdpc1)]
X <- X[date <= as.Date("2017-01-01")]

lags <- 2
bc_freqs <- c(2 * pi / 32, 2 * pi / 6)

v <- VAR(X[, -"date"], p = lags, type = "const")

X[, t := seq_len(nrow(X)) - lags]
```

```{r}
data <- copy(X)

data <- data |>
  dt_pivot_longer(-c(date, t), names_to = "variable", values_to = "original")

resid_data <-
  residuals(v) |>
  as.data.table() |>
  _[, t := .I] |>
  dt_pivot_longer(-t, names_to = "variable", values_to = "residuals")

fitted_data <-
  fitted(v) |>
  as.data.table() |>
  _[, t := .I] |>
  dt_pivot_longer(-t, names_to = "variable", values_to = "fitted")

data <- data |>
  merge(resid_data, by = c("t", "variable")) |>
  merge(fitted_data, by = c("t", "variable"))

data[, log_var := log(residuals^2)]
```

```{r, include = TRUE, fig.height = 8, out.width = "100%"}
p1 <- data |>
  ggplot(aes(
    x = date
  )) +
  geom_line(aes(y = original, color = "original")) +
  geom_line(aes(y = fitted, color = "fitted")) +
  facet_grid(rows = vars(variable), scales = "free") +
 theme_bw(base_size = 8) +
 theme(legend.position = "top")

p2 <- data |>
  ggplot(aes(
    x = date
  )) +
  geom_line(aes(y = residuals, color = "residuals")) +
  facet_grid(rows = vars(variable), scales = "free") +
 theme_bw(base_size = 8) +
 theme(legend.position = "top")

p1 | p2
```

```{r}
reg_data <- data |>
  merge(
    dt_pivot_wider(data[,.(t, variable, original)], names_from = variable, values_from = original),
    by = c("t"),
    allow.cartesian = TRUE
  )

fitted_log_var_data <-
  rbindlist(lapply(unique(reg_data$variable), function(n) {
    het_model <- lm(formula = "log_var ~ 
    lag(output) + lag(investment) + lag(consumption) + lag(hours_worked) + lag(unemployment) + 
    lag(labor_share) + lag(interest) + lag(inflation) + lag(productivity) + lag(TFP) + lag(vfci) +
    lag(output, 2) + lag(investment, 2) + lag(consumption, 2) + lag(hours_worked, 2) + lag(unemployment, 2) + 
    lag(labor_share, 2) + lag(interest, 2) + lag(inflation, 2) + lag(productivity, 2) + lag(TFP, 2) + lag(vfci, 2)
    ", data = reg_data[variable == n])
    fitted(het_model) |>
      as.data.table() |>
      setnames("V1", "fitted_log_var") |>
      _[, variable := n] |>
      _[, t := .I] |>
      _[, resid_log_var := residuals(het_model)]
  }))

data <- data |>
  merge(fitted_log_var_data, by = c("t", "variable"))
```

\newpage

\subsection{Fitting the heteroskedastic variance regression}

(2) Construct the log-variance for each residual. Fit a regression on the same variables as the VAR, find the fitted value.
```{r, include = TRUE, fig.height = 8, out.width = "100%"}
p1 <- data |>
  ggplot(aes(
    x = date
  )) +
  geom_line(aes(y = log_var, color = "original log_var")) +
  geom_line(aes(y = fitted_log_var, color = "fitted log_var")) +
  facet_grid(rows = vars(variable), scales = "free") +
 theme_bw(base_size = 8) +
 theme(legend.position = "top")

p2 <- data |>
  ggplot(aes(
    x = date
  )) +
  geom_line(aes(y = resid_log_var, color = "residuals log_var")) +
  facet_grid(rows = vars(variable), scales = "free") +
 theme_bw(base_size = 8) +
 theme(legend.position = "top")

p1 | p2
```
\newpage

\subsection{Nonstationary Series}

(3) Some of the values in $\widehat{y}_t$ are non-stationary (they have a trend), so we take differences before presenting them in the mean-vol and mean-var figures.

The differences variables are: output, consumption, investment, productivity, and TFP


```{r, include = TRUE, fig.height = 8, out.width = "100%"}
diff_vars <- c("output", "consumption", "investment", "productivity", "TFP")
data[, fitted_adj := fitted]
data[variable %in% diff_vars, fitted_adj := fitted - shift(fitted, n = 1, type = "lag"), by = "variable"]

p1 <- data |>
  ggplot(aes(
    x = date,
    y = fitted
  )) +
  geom_line() +
  facet_grid(rows = vars(variable), scales = "free") +
  labs(subtitle = "Fitted Series") +
  theme_bw(base_size = 8) +
  theme(legend.position = "top")

p2 <- data |>
  ggplot(aes(
    x = date,
    y = fitted_adj
  )) +
  geom_line() +
  facet_grid(rows = vars(variable), scales = "free") +
  labs(subtitle = "Adjusted Fitted Series") +
  theme_bw(base_size = 8) +
  theme(legend.position = "top")

p1 | p2
```

\newpage

\subsection{Mean-Vol Relationship}

(4) Chart the relationship between the fitted means and the fitted volatilities.

```{r, include = TRUE, fig.height = 8, out.width = "100%"}
data |>
  ggplot(aes(
    x = sqrt(exp(fitted_log_var)),
    y = fitted_adj
  )) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(
    vars(variable),
    ncol = 3,
    scales = "free"
  ) +
  theme_bw(base_size = 8) +
  theme(legend.position = "top")
```

\newpage

\subsection{Mean-Var Relationship}

(5) Chart the relationship between the fitted means and the fitted log-variances.

```{r, include = TRUE, fig.height = 8, out.width = "100%"}
data |>
  ggplot(aes(
    x = fitted_log_var,
    y = fitted_adj
  )) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(
    vars(variable),
    ncol = 3,
    scales = "free"
  ) +
  theme_bw(base_size = 8) +
  theme(legend.position = "top")
```





\newpage

\section{Using 10-variable VAR}

We can repeat the above analysis with just 10 variables in the VAR (dropping vfci).

```{r, include = F}
X <- fread(here("./data/vfciBC_data.csv"))
X <- X[,.(date, output, investment, consumption, hours_worked, unemployment, labor_share, interest, inflation, productivity, TFP)]
X <- X[date <= as.Date("2017-01-01")]

lags <- 2
bc_freqs <- c(2 * pi / 32, 2 * pi / 6)

v <- VAR(X[, -"date"], p = lags, type = "const")

X[, t := seq_len(nrow(X)) - lags]
```

```{r}
data <- copy(X)

data <- data |>
  dt_pivot_longer(-c(date, t), names_to = "variable", values_to = "original")

resid_data <-
  residuals(v) |>
  as.data.table() |>
  _[, t := .I] |>
  dt_pivot_longer(-t, names_to = "variable", values_to = "residuals")

fitted_data <-
  fitted(v) |>
  as.data.table() |>
  _[, t := .I] |>
  dt_pivot_longer(-t, names_to = "variable", values_to = "fitted")

data <- data |>
  merge(resid_data, by = c("t", "variable")) |>
  merge(fitted_data, by = c("t", "variable"))

data[, log_var := log(residuals^2)]
```


```{r}
reg_data <- data |>
  merge(
    dt_pivot_wider(data[,.(t, variable, original)], names_from = variable, values_from = original),
    by = c("t"),
    allow.cartesian = TRUE
  )

fitted_log_var_data <-
  rbindlist(lapply(unique(reg_data$variable), function(n) {
    het_model <- lm(formula = "log_var ~ 
    lag(output) + lag(investment) + lag(consumption) + lag(hours_worked) + lag(unemployment) + 
    lag(labor_share) + lag(interest) + lag(inflation) + lag(productivity) + lag(TFP) +
    lag(output, 2) + lag(investment, 2) + lag(consumption, 2) + lag(hours_worked, 2) + lag(unemployment, 2) + 
    lag(labor_share, 2) + lag(interest, 2) + lag(inflation, 2) + lag(productivity, 2) + lag(TFP, 2) 
    ", data = reg_data[variable == n])
    fitted(het_model) |>
      as.data.table() |>
      setnames("V1", "fitted_log_var") |>
      _[, variable := n] |>
      _[, t := .I] |>
      _[, resid_log_var := residuals(het_model)]
  }))

data <- data |>
  merge(fitted_log_var_data, by = c("t", "variable"))
```

```{r, include = TRUE, fig.height = 8, out.width = "100%"}
diff_vars <- c("output", "consumption", "investment", "productivity", "TFP")
data[, fitted_adj := fitted]
data[variable %in% diff_vars, fitted_adj := fitted - shift(fitted, n = 1, type = "lag"), by = "variable"]
```

\subsection{Mean-Vol Relationship - 10 variable VAR}

```{r, include = TRUE, fig.height = 8, out.width = "100%"}
data |>
  ggplot(aes(
    x = sqrt(exp(fitted_log_var)),
    y = fitted_adj
  )) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(
    vars(variable),
    ncol = 3,
    scales = "free"
  ) +
  theme_bw(base_size = 8) +
  theme(legend.position = "top")
```

\newpage

\subsection{Mean-Var Relationship - 10 variable VAR}

```{r, include = TRUE, fig.height = 8, out.width = "100%"}
data |>
  ggplot(aes(
    x = fitted_log_var,
    y = fitted_adj
  )) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(
    vars(variable),
    ncol = 3,
    scales = "free"
  ) +
  theme_bw(base_size = 8) +
  theme(legend.position = "top")
```








\newpage

\section{Estimating the VAR in differences}

We can estimate the VAR in differences rather than taking differences of the non-stationary variables.

```{r, include = F}
X <- fread(here("./data/vfciBC_data.csv"))
X <- X[,.(date, output, investment, consumption, hours_worked, unemployment, labor_share, interest, inflation, productivity, TFP, vfci = vfci_fgr10gdpc1)]
diff_vars <- c("output", "consumption", "investment", "productivity", "TFP")
X[, output := output - shift(output, 1, type = "lag")]
X[, consumption := consumption - shift(consumption, 1, type = "lag")]
X[, investment := investment - shift(investment, 1, type = "lag")]
X[, productivity := productivity - shift(productivity, 1, type = "lag")]
X[, TFP := TFP - shift(TFP, 1, type = "lag")]
X <- X[!is.na(output)]
X <- X[date <= as.Date("2017-01-01")]

lags <- 2
bc_freqs <- c(2 * pi / 32, 2 * pi / 6)

v <- VAR(X[, -"date"], p = lags, type = "const")

X[, t := seq_len(nrow(X)) - lags]
```

```{r}
data <- copy(X)

data <- data |>
  dt_pivot_longer(-c(date, t), names_to = "variable", values_to = "original")

resid_data <-
  residuals(v) |>
  as.data.table() |>
  _[, t := .I] |>
  dt_pivot_longer(-t, names_to = "variable", values_to = "residuals")

fitted_data <-
  fitted(v) |>
  as.data.table() |>
  _[, t := .I] |>
  dt_pivot_longer(-t, names_to = "variable", values_to = "fitted")

data <- data |>
  merge(resid_data, by = c("t", "variable")) |>
  merge(fitted_data, by = c("t", "variable"))

data[, log_var := log(residuals^2)]
```

```{r, include = TRUE, fig.height = 8, out.width = "100%"}
p1 <- data |>
  ggplot(aes(
    x = date
  )) +
  geom_line(aes(y = original, color = "original")) +
  geom_line(aes(y = fitted, color = "fitted")) +
  facet_grid(rows = vars(variable), scales = "free") +
 theme_bw(base_size = 8) +
 theme(legend.position = "top")

p2 <- data |>
  ggplot(aes(
    x = date
  )) +
  geom_line(aes(y = residuals, color = "residuals")) +
  facet_grid(rows = vars(variable), scales = "free") +
 theme_bw(base_size = 8) +
 theme(legend.position = "top")

p1 | p2
```

```{r}
reg_data <- data |>
  merge(
    dt_pivot_wider(data[,.(t, variable, original)], names_from = variable, values_from = original),
    by = c("t"),
    allow.cartesian = TRUE
  )

fitted_log_var_data <-
  rbindlist(lapply(unique(reg_data$variable), function(n) {
    het_model <- lm(formula = "log_var ~ 
    lag(output) + lag(investment) + lag(consumption) + lag(hours_worked) + lag(unemployment) + 
    lag(labor_share) + lag(interest) + lag(inflation) + lag(productivity) + lag(TFP) + lag(vfci) +
    lag(output, 2) + lag(investment, 2) + lag(consumption, 2) + lag(hours_worked, 2) + lag(unemployment, 2) + 
    lag(labor_share, 2) + lag(interest, 2) + lag(inflation, 2) + lag(productivity, 2) + lag(TFP, 2) + lag(vfci, 2)
    ", data = reg_data[variable == n])
    fitted(het_model) |>
      as.data.table() |>
      setnames("V1", "fitted_log_var") |>
      _[, variable := n] |>
      _[, t := .I] |>
      _[, resid_log_var := residuals(het_model)]
  }))

data <- data |>
  merge(fitted_log_var_data, by = c("t", "variable"))
```

\newpage

\subsection{Mean-Vol Relationship - VAR estimated in differences}

```{r, include = TRUE, fig.height = 8, out.width = "100%"}
data |>
  ggplot(aes(
    x = sqrt(exp(fitted_log_var)),
    y = fitted
  )) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(
    vars(variable),
    ncol = 3,
    scales = "free"
  ) +
  theme_bw(base_size = 8) +
  theme(legend.position = "top")
```

\newpage

\subsection{Mean-Var Relationship - VAR estimated in differences}

```{r, include = TRUE, fig.height = 8, out.width = "100%"}
data |>
  ggplot(aes(
    x = fitted_log_var,
    y = fitted
  )) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(
    vars(variable),
    ncol = 3,
    scales = "free"
  ) +
  theme_bw(base_size = 8) +
  theme(legend.position = "top")
```










\newpage

\section{Business Cycle Shock Mean-Vol}

This section attempts to construct the same mean-vol relationship for just the identified business cycle shock.

This requires a change in the definition of the "mean" values.

We would expect (and we find) a null result here, as the business cycle shock should be homoskedastic.

\subsubsection{Max Share ID}

Using the max share identification method, we identify one structural shock, $w^{\text{BC}}_t$, which drives the business cycle.

\begin{align}
u_t = P Q^* w_t
\end{align}

where $P$ is the choleskey of $\Sigma_u$ and $Q^*$ is the identified rotation matrix that returns $w^{\text{BC}}_t$ in the first column.

Then we can write,
\begin{align}
y_t = A_1 y_{t-1} + ... + A_{p} y_{t-p} + P Q^* \begin{bmatrix}w^{\text{BC}}_t \\ w^2_t \\ w^3_t \\ . . . \\ w^k_t \end{bmatrix}
\end{align}

\subsubsection{Business Cycle Contribution}

Using the identified business cycle shock, we can construct the time series for $y_t^{\text{BC}}$ which are the contributions to each variable of the VAR driven by the identified shock.

\begin{align}
  \widehat{y^{\text{BC}}}_t = \widehat{\begin{bmatrix}\text{output}^{\text{BC}}_t\\\text{unemployment}^{\text{BC}}_t\\\text{inflation}^{\text{BC}}_t\\...\\\text{vfci}^{\text{BC}}_t\end{bmatrix}}
\end{align}

\subsubsection{Business Cycle Volatility}

The log-variance for the business cycle can be defined as:
\begin{align}
\text{Var}^\text{BC}_t = log(\left[\widehat{w}^{\text{BC}}_t\right]^2)
\end{align}

This can also be modeled using the variables of the VAR:
\begin{align}
\text{Var}^\text{BC}_t = \alpha_1 y_{t-1} + ... + \alpha_{p} y_{t-p} + \epsilon_t
\end{align}

where $\alpha_i$ is a $1xk$ vector of coefficients.

We can then estimate $\widehat{\text{Var}}^\text{BC}_t$ as a time series.

This can be done for each model that identifies a structural shock.
In our case, we can compare the model targetting unemployment and the one targetting vfci.

```{r, include = F}
X <- fread(here("./data/vfciBC_data.csv"))
X <- X[,.(date, output, investment, consumption, hours_worked, unemployment, labor_share, interest, inflation, productivity, TFP, vfci = vfci_fgr10gdpc1)]
X <- X[date <= as.Date("2017-01-01")]

lags <- 2
bc_freqs <- c(2 * pi / 32, 2 * pi / 6)

v <- VAR(X[, -"date"], p = lags, type = "const")

X[, t := seq_len(nrow(X)) - lags]
```

```{r make_shocks, include = F}
models <- list(
  mv_u = id_fevdfd(v, "unemployment", bc_freqs),
  mv_vfci = id_fevdfd(v, "vfci", bc_freqs, sign = "neg")
)

contributions <- rbindlist(lapply(names(models), function(n) {
  hd(models[[n]])$hd |>
    as.data.table() |>
    _[, model := n]
}))

shocks <- lapply(names(models), function(n) {
  hs(models[[n]])$hs |>
    as.data.table() |>
    _[, model := n] |>
    _[, log_var := log(hs^2)]
})

het_data <- lapply(shocks, function(s) {
  s |>
  merge(X, by = "t")
})

het_models <- lapply(het_data, function(h) {
  lm(formula = "log_var ~ 
  lag(output) + lag(investment) + lag(consumption) + lag(hours_worked) + lag(unemployment) + 
  lag(labor_share) + lag(interest) + lag(inflation) + lag(productivity) + lag(TFP) + lag(vfci) +
  lag(output, 2) + lag(investment, 2) + lag(consumption, 2) + lag(hours_worked, 2) + lag(unemployment, 2) + 
  lag(labor_share, 2) + lag(interest, 2) + lag(inflation, 2) + lag(productivity, 2) + lag(TFP, 2) + lag(vfci, 2)
  ", data = h)
})

pred_log_var <- rbindlist(lapply(seq_along(het_data), function(i) {
  het_data[[i]][, predicted_log_var := predict(het_models[[i]])]
}))


## Merge mean and vol data
mv_data <- merge(
  contributions,
  pred_log_var,
  by = c("model", "t", "impulse"),
  allow.cartesian = TRUE
)
```

\newpage 

Here are the contributions of the Business Shock, $\widehat{y_t^{\text{BC}}}$, to a subset of the variables, compared to the total unexplained varaition in those series.

```{r, include = TRUE, fig.height = 4, out.width = "100%"}
mv_data |>
 _[response %in% c("output", "unemployment", "interest", "inflation", "vfci")] |>
 _[impulse == "Main"] |>
 ggplot(aes(
   x = date
 )) +
 geom_line(aes(y = total, color = "Total Residual")) +
 geom_line(aes(y = hd, color = "Contribution from BC Shock")) +
 facet_grid(
  cols = vars(model),
  rows = vars(response),
  scales = "free"
 ) +
 theme_bw(base_size = 8) +
 theme(legend.position = "top")
```

And here is the log-variance of the business cycle shock, $\text{Var}^\text{BC}_t$, and the predicted value, $\widehat{\text{Var}}^\text{BC}_t$:

```{r, include = TRUE, fig.height = 3, out.width = "100%"}
mv_data |>
  _[impulse == "Main"] |>
  ggplot(aes(
    x = date
  )) +
  geom_line(aes(y = log_var, color = "log_var")) +
  geom_line(aes(y = predicted_log_var, color = "predicted log_var")) +
  facet_grid(
    cols = vars(model),
    scales = "free"
  ) +
 theme_bw(base_size = 8) +
 theme(legend.position = "top")
```

\newpage 

The correlation between the two sets of predicted values are not particularly high.

This is just showing a small subsample of the variables in the VAR.

```{r, include = TRUE, fig.height = 8, fig.width = 4, out.width = "100%", fig.align = "center"}
mv_data |>
  _[response %in% c("unemployment", "output", "interest", "inflation", "vfci")] |>
  _[impulse == "Main"] |>
  ggplot(aes(
    x = sqrt(exp(log_var)),
    y = hd
  )) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(
    rows = vars(response),
    cols = vars(model),
    scales = "free"
  ) +
  labs(y = "Contribution from BC Shock to Means") +
 theme_bw(base_size = 8) +
 theme(legend.position = "top")
```

