---
title: "VFCI Business Cycle VAR IRFs"
output: 
    pdf_document
---
```{r setup, include = FALSE}
require(here)
require(data.table)
require(ggplot2)
```

```{r loadData, include = F}
df <- fread(here("./data/vfciBC_classical_VAR_IRF.csv"))
vfci <- fread(here("./data/vfciBC_data.csv"))
```

## Adding VFCI to VAR
All VARs are identified by targetting unemployment in the business cycle frequency.

Comparing here three VARs:

  - original: 10 variables, BCA sample
  - bca_variables: 10 variables, vfci sample
  - vfci: 11 variables, vfci sample

The original BCA paper uses data from 1955 to 2017.
The new vfci data uses a sample from `r min(vfci$date)` to 2017.

Adding in the vfci series does not change much for the 10 variables already in the VAR.
The change in the sample does have an impact, but all of the impulse response are still similar to before.

```{r figVARubc, echo = F, warning = F, fig.height = 3, out.width = "100%"}
df[target == "u_bc"] |>
    ggplot(aes(
        x = h,
        y = irf,
        color = version
    )) +
    geom_hline(yintercept = 0) +
    geom_line() +
    facet_wrap(vars(response), nrow = 2, scales = "free_y") +
    labs(x = "Horizon", y = NULL) +
    theme_bw(base_size = 8) +
    theme(
        legend.position = "bottom",
        strip.background = element_blank(),
        panel.grid.minor = element_blank()
        )
```


## Targetting VFCI Instead

```{r figVARvfci, echo = F, warning = F, fig.height = 3, out.width = "100%"}
df[version == "vfci"] |>
    ggplot(aes(
        x = h,
        y = irf,
        color = target
    )) +
    geom_hline(yintercept = 0) +
    geom_line() +
    facet_wrap(vars(response), nrow = 2, scales = "free_y") +
    labs(x = "Horizon", y = NULL) +
    theme_bw(base_size = 8) +
    theme(
        legend.position = "bottom",
        strip.background = element_blank(),
        panel.grid.minor = element_blank()
        )
```