---
title: "VFCI Business Cycle VAR IRFs"
output: 
    pdf_document
---
```{r setup, include = FALSE}
require(here)
require(data.table)
require(ggplot2)
```

```{r loadData, include = F}
df <- fread(here("./data/classical_vfcibc_VAR_IRF.csv"))
df[, label := paste(target, period_l, ":", period_h)]
rmse <- fread(here("./data/classical_vfcibc_VAR_IRF_rmse.csv"))
vfci <- fread(here("./data/vfciBC_data.csv"))
bca_irf <- fread(here("./data/replicated_bca_classical_VAR_IRF.csv"))
```

```{r, echo = F, fig.height = 3, out.width = "100%"}
ggplot() +
    geom_hline(yintercept = 0) +
    geom_line(
        data = df[target == "unemployment" & sign == "pos" & period_l == 6 & period_h == 32],
        aes(
            x = h,
            y = irf,
            color = "+ vfci (11 variables)"
    )) +
    geom_line(
        data = bca_irf[version == "Replication"],
        aes(
            x = h,
            y = irf,
            color = "BCA (10 variables)"
    )) +
    facet_wrap(
        vars(response),
        nrow = 2,
        scales = "free_y"
    ) +
    theme_bw(base_size = 8) +
    theme(
        legend.position = "bottom"
    )
  

```

## All iterations

```{r figVARubc, echo = F, warning = F, fig.height = 7, out.width = "80%"}
df |>
    dplyr::filter((target == "vfci" & sign == "neg") | (target == "unemployment" & u_irfsign == "neg")) |>
    ggplot(aes(
        x = h,
        y = irf,
        color = period_l,
        group = label
    )) +
    geom_hline(yintercept = 0) +
    geom_line() +
    facet_grid(rows = vars(response), cols = vars(target), scales = "free_y") +
    scale_color_viridis_c() +
    theme_bw(base_size = 8) +
    theme(legend.position = "right")
```

\newpage

Showing the IRFS of targetting unemployment and VFCI at the business cycle frequencies.
```{r}
rmse[u_sign == "pos" & u_period_l == 6 & u_period_h == 32][vfci_sign == "neg" & vfci_period_l == 6 & vfci_period_h == 32]
```

```{r figBC, echo = F, warning = F, fig.height = 3, out.width = "100%"}
df |>
    dplyr::filter((target == "vfci" & sign == "neg" & period_l == 6 & period_h == 32) | (target == "unemployment" & sign == "pos" & period_l == 6 & period_h == 32)) |>
    ggplot(aes(
        x = h,
        y = irf,
        color = label
    )) +
    geom_hline(yintercept = 0) +
    geom_line() +
    facet_wrap(vars(response), nrow = 2, scales = "free_y") +
    theme_bw(base_size = 8) +
    theme(legend.position = "bottom")
```

Picking the VFCI Iteration that most closely matches the Main Business Cycle shock across all impulse responses (target: unemployment, 6 - 32 q).

```{r}
rmse[u_sign == "pos" & u_period_l == 6 & u_period_h == 32][rmse == min(rmse)]
```

```{r figMatchMBCall, echo = F, warning = F, fig.height = 3, out.width = "100%"}

df |>
    dplyr::filter((target == "vfci" & sign == "neg" & period_l == 28 & period_h == 76) | (target == "unemployment" & sign == "pos" & period_l == 6 & period_h == 32)) |>
    ggplot(aes(
        x = h,
        y = irf,
        color = label
    )) +
    geom_hline(yintercept = 0) +
    geom_line() +
    facet_wrap(vars(response), nrow = 2, scales = "free_y") +
    theme_bw(base_size = 8) +
    theme(legend.position = "bottom")
```


\newpage

Picking the two iterations of VFCI and Unemployment that are most closely similar across all responses.

```{r}
rmse[rmse == min(rmse)]
```

```{r figMatchUVFCI, echo = F, warning = F, fig.height = 3, out.width = "100%"}
df |>
    dplyr::filter((target == "vfci" & sign == "neg" & period_l == 24 & period_h == 30) | (target == "unemployment" & sign == "pos" & period_l == 12 & period_h == 38)) |>
    ggplot(aes(
        x = h,
        y = irf,
        color = label
    )) +
    geom_hline(yintercept = 0) +
    geom_line() +
    facet_wrap(vars(response), nrow = 2, scales = "free_y") +
    theme_bw(base_size = 8) +
    theme(legend.position = "bottom")
```


Picking the two iterations of VFCI and unemployment that are the most closely similar across all responses and have the same target period.

```{r}
rmse[u_period_l == vfci_period_l & u_period_h == vfci_period_h][rmse == min(rmse)]
```

```{r figMatchUVFCI_sameperiod, echo = F, warning = F, fig.height = 3, out.width = "100%"}
df |>
    dplyr::filter((target == "vfci" & sign == "neg" & period_l == 20 & period_h == 32) | (target == "unemployment" & sign == "pos" & period_l == 20 & period_h == 32)) |>
    ggplot(aes(
        x = h,
        y = irf,
        color = label
    )) +
    geom_hline(yintercept = 0) +
    geom_line() +
    facet_wrap(vars(response), nrow = 2, scales = "free_y") +
    theme_bw(base_size = 8) +
    theme(legend.position = "bottom")
```

\newpage

What does targetting the other half of the business cycle requency look like?

```{r figOtherBC, echo = F, warning = F, fig.height = 3, out.width = "100%"}
df |>
    dplyr::filter((target == "vfci" & sign == "neg" & period_l == 6 & period_h == 20) | (target == "unemployment" & sign == "pos" & period_l == 6 & period_h == 20)) |>
    ggplot(aes(
        x = h,
        y = irf,
        color = label
    )) +
    geom_hline(yintercept = 0) +
    geom_line() +
    facet_wrap(vars(response), nrow = 2, scales = "free_y") +
    theme_bw(base_size = 8) +
    theme(legend.position = "bottom")
```

\newpage

# Targetting Unemployment IRF

Picking the VFCI Iteration that most closely matches the Main Business Cycle shock for just the unemployment impulse responses (target: unemployment, 6 - 32 q). 

```{r}
rmse[u_sign == "pos" & u_period_l == 6 & u_period_h == 32][u_rmse == min(u_rmse)]
```

```{r, echo = F, warning = F, fig.height = 3, out.width = "100%"}
df |>
    dplyr::filter((target == "vfci" & sign == "neg" & period_l == 30 & period_h == 80) | (target == "unemployment" & sign == "pos" & period_l == 6 & period_h == 32)) |>
    ggplot(aes(
        x = h,
        y = irf,
        color = label
    )) +
    geom_hline(yintercept = 0) +
    geom_line() +
    facet_wrap(vars(response), nrow = 2, scales = "free_y") +
    theme_bw(base_size = 8) +
    theme(legend.position = "bottom")
```


Picking the two iterations of VFCI and Unemployment that are most closely similar for the unemployment irf. 

```{r}
rmse[u_rmse == min(u_rmse)]
```

```{r, echo = F, warning = F, fig.height = 3, out.width = "100%"}
df |>
    dplyr::filter((target == "vfci" & sign == "neg" & period_l == 22 & period_h == 28) | (target == "unemployment" & sign == "pos" & period_l == 22 & period_h == 30)) |>
    ggplot(aes(
        x = h,
        y = irf,
        color = label
    )) +
    geom_hline(yintercept = 0) +
    geom_line() +
    facet_wrap(vars(response), nrow = 2, scales = "free_y") +
    theme_bw(base_size = 8) +
    theme(legend.position = "bottom")
```

\newpage

Picking the two iterations of VFCI and unemployment that are the most closely similar for unemployment and have the same target period.  Exact same frequency as before.

```{r}
rmse[u_period_l == vfci_period_l & u_period_h == vfci_period_h][u_rmse == min(u_rmse)]
```

```{r, echo = F, warning = F, fig.height = 3, out.width = "100%"}
df |>
    dplyr::filter((target == "vfci" & sign == "neg" & period_l == 22 & period_h == 30) | (target == "unemployment" & sign == "pos" & period_l == 22 & period_h == 30)) |>
    ggplot(aes(
        x = h,
        y = irf,
        color = label
    )) +
    geom_hline(yintercept = 0) +
    geom_line() +
    facet_wrap(vars(response), nrow = 2, scales = "free_y") +
    theme_bw(base_size = 8) +
    theme(legend.position = "bottom")
```

\newpage

# Splitting the Business Cycle frequency

Hold the high period target constant at 32 quarters, let the low end vary, compare targetting VFCI or unemployment.
The two IRFs noticeably converge at 22q for the low period.
Note that the VFCI IRFs don't change much, but the unemployment targetting one does.

```{r, echo = F, warning = F, fig.height = 8, out.width = "100%"}
df |>
    dplyr::filter((target == "vfci" & sign == "neg" & period_l >= 6 & period_h == 32) | (target == "unemployment" & u_irfsign == "neg" & period_l >= 6 & period_h == 32)) |>
    ggplot(aes(
        x = h,
        y = irf,
        color = target
    )) +
    geom_hline(yintercept = 0) +
    geom_line() +
    facet_grid(rows = vars(response), cols = vars(period_l), scales = "free_y") +
    theme_bw(base_size = 8) +
    theme(legend.position = "bottom")
```


\newpage

Hold the low period target constant at 6 quarters, let the high end vary, compare targetting VFCI or unemployment.

```{r, echo = F, warning = F, fig.height = 8, out.width = "100%"}
df |>
    dplyr::filter((target == "vfci" & sign == "neg" & period_l == 6 & period_h <= 32) | (target == "unemployment" & u_irfsign == "neg" & period_l == 6 & period_h <= 32)) |>
    ggplot(aes(
        x = h,
        y = irf,
        color = target
    )) +
    geom_hline(yintercept = 0) +
    geom_line() +
    facet_grid(rows = vars(response), cols = vars(period_h), scales = "free_y") +
    theme_bw(base_size = 8) +
    theme(legend.position = "bottom")
```
